---
title: "Functions and Algebra of Functions"
subtitle: "Session 01"  
author: 
  - "Alejandro Ucan"
date: '`r Sys.Date()`'
output:
  xaringan::moon_reader:
    css: xaringan-themer.css
    nature:
      slideNumberFormat: "%current%"
      highlightStyle: github
      highlightLines: true
      ratio: 16:9
      countIncrementalSlides: true
---

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
knitr::opts_chunk$set(
  fig.width=9, fig.height=3.5, fig.retina=3,
  out.width = "100%",
  cache = FALSE,
  echo = FALSE,
  message = FALSE, 
  warning = FALSE,
  hiline = TRUE
)
```

```{r xaringan-themer, include=FALSE, warning=FALSE}
library(xaringanthemer)
style_duo_accent(
  primary_color = "#000080",
  secondary_color = "#f2f3f4",
  inverse_header_color = "#00147e"
)
```

```{r xaringan-editable, echo=FALSE}
xaringanExtra::use_editable(expires = 1)
xaringanExtra::use_scribble()
```

```{r xaringan-logo, echo=FALSE}
xaringanExtra::use_logo(
  image_url = "https://github.com/alxcn/TecLogoEIC/blob/9562a53875418e749a296c85808a19c85fc4be74/IngenieriaCiencias_Horizontal_RGB.png",
  position = xaringanExtra::css_position(top = "2em", right = "2em")
)
```
 Session's Goals

 * Introduce set relations.
 * Define a function.
 * Describe the parts of a function.
 * Describe the algebra between functions.

---
 # Motivation

Communly, in life, we have different quantities that vary with respect to others. For example: <br/><br/>
  * The price depends on the offer and demand.
  * The blood sugar depends on the food we eat.
  * The resistence of a concret mix depends on the amount of water we use.

Those dependencies are relations between quantities. In this session, we will introduce the mathematical concept that allows us to describe those relations: **functions**.

---
# Set Relations

> **Definition:** Given two sets $D$ and $C,$ a relation $R$ from $D$ to $C$ is just a way to pair elements of $D$ with elements of $C.$ <br/><br/>

For example: <br/><br/>
 * If $D$ is the set of people, and $C$ is the set of positive numbers, we can define a relation $R$ that pairs each person with their age.
 * If $D$ is the set of students, and $C$ is the set of courses, we can define a relation $R$ that pairs each student with the courses they are taking.
 * If $D$ is the set of real numbers, and $C$ is the set of real numbers, we can define a relation $R$ that pairs each number with its square.

---
## Functions

> **Definition:** A function $f$ from a set $D$ to a set $C$ is a relation that pairs each element of $D$ with exactly one element of $C.$ <br/>
The set $D$ is called the **domain** of $f,$ and the set $C$ is called the **codomain** of $f.$ <br/><br/>

For example: <br/><br/>
 * The function that pairs each person with their age is a function from the set of people to the set of positive numbers.
 * The function that pairs each student with the grade they got in a course is a function from the set of students to the set of grades.
 * The function that pairs each number with its square is a function from the set of real numbers to the set of real numbers.

---
### Not all Relations are Functions:

Some counterexamples of functions are: <br/><br/>
 * The relation that pairs each person with their favorite color is not a function because a person can have more than one favorite color.
 * The relation that pairs each student with the courses they are taking is not a function because a student can take more than one course.
 * The relation that pairs each number with its square root is not a function because a number can have more than one square root.

---
# Functions in real life contexts:

* The pesos that we get for a dollar is a function of the exchange rate.
* A plant produced oxygen depends on the amount of light it receives.
* The time it takes to get to a place depends on the speed we drive.
* The value in a stock market depends on the offer and demand.
---
## Modelling using functions: 

In general, we can model a quantity that depends on another using a function. For example: <br/><br/>

* The distance traveled by a particle (with constant speed) depends on the time, and can be modeled by $d(t) = v \cdot t,$ where $d(t)$ is the distance traveled at time $t,$ and $v$ is the speed of the particle. <br/><br/>

* The compound interest for an investment depends on the time, and can be modeled by $A(t) = P \cdot (1 + r)^t,$ where $A(t)$ is the amount of money at time $t,$ $P$ is the initial investment, $r$ is the interest rate, and $t$ is the time. <br/><br/>

---
## Modelling using functions:

* The temperature of a body depends on the time, and can be modeled by $T(t) = T_0 + (T_1 - T_0) \cdot e^{-k \cdot t},$ where $T(t)$ is the temperature at time $t,$ $T_0$ is the initial temperature, $T_1$ is the final temperature, and $k$ is a constant. <br/><br/>

* The position of a pendullum depends on the time, and can be modeled by $x(t) = A \cdot \cos(\omega \cdot t),$ where $x(t)$ is the position at time $t,$ $A$ is the amplitude, and $\omega$ is the angular velocity. <br/><br/>

---
## Types of Functions:

There are different types of functions, depending on the domain and codomain. Some examples are: <br/><br/>

* *Polynomial Functions:* Function given by a polynomial expression of some degree. Has the form $f(x) = a_n \cdot x^n + a_{n-1} \cdot x^{n-1} + \ldots + a_1 \cdot x + a_0.$ Whose domain is the set of real numbers, and codomain is the set of real numbers. <br/><br/>

* *Rational Functions:* Function given by a ratio of two polynomial functions. Has the form $f(x) = \frac{p(x)}{q(x)},$ where $p(x)$ and $q(x)$ are polynomial functions. Whose domain is the set of real numbers except for those where the denominator is zero, and codomain is the set of real numbers. <br/><br/>

---

## Types of Functions:

* *Root Functions:* Function given rational power of the variable. Has the form $f(x) = \sqrt[n]{x},$ where $n$ is positive integer. Whose domain is depends on the value of $n,$ and codomain is the set of real numbers. For odd $n,$ the domain is the set of real numbers, and for even $n,$ the domain is the set of non-negative real numbers. <br/><br/>

* *Trigonometric Functions:* Function given by a trigonometric expression. Has the form $f(x) = \sin(x),$ $f(x) = \cos(x),$ $f(x) = \tan(x),$ etc. Whose domain is the set of real numbers, and codomain is the set of real numbers. <br/><br/>

---
# Functions using other functions:

Given two functions $f$ and $g,$ we can define new functions using $f$ and $g.$ Some examples are: <br/><br/>

* *Sum of functions:* Given $f$ and $g,$ we can define a new function $h$ by $h(x) = f(x) + g(x).$ <br/><br/>
* *Product of functions:* Given $f$ and $g,$ we can define a new function $h$ by $h(x) = f(x) \cdot g(x).$ <br/><br/>
* *Quotient of functions:* Given $f$ and $g,$ we can define a new function $h$ by $h(x) = \frac{f(x)}{g(x)}$ for 4$g(x) \neq 0.$ <br/><br/>
