---
title: "Exponential Models"
subtitle: "Session 07"
author: 
  - "Alejandro Ucan"
date: '`r Sys.Date()`'
output:
  xaringan::moon_reader:
    css: xaringan-themer.css
    nature:
      slideNumberFormat: "%current%"
      highlightStyle: github
      highlightLines: true
      ratio: 16:9
      countIncrementalSlides: true
---

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
knitr::opts_chunk$set(
  fig.width=9, fig.height=3.5, fig.retina=3,
  out.width = "100%",
  cache = FALSE,
  echo = FALSE,
  message = FALSE, 
  warning = FALSE,
  hiline = TRUE
)
```

```{r xaringan-themer, include=FALSE, warning=FALSE}
library(xaringanthemer)
style_duo_accent(
  primary_color = "#000080",
  secondary_color = "#f2f3f4",
  inverse_header_color = "#00147e"
)
```

```{r xaringan-editable, echo=FALSE}
xaringanExtra::use_editable(expires = 1)
xaringanExtra::use_scribble()
```

```{r xaringan-logo, echo=FALSE}
xaringanExtra::use_logo(
  image_url = "https://github.com/alxcn/TecLogoEIC/blob/9562a53875418e749a296c85808a19c85fc4be74/IngenieriaCiencias_Horizontal_RGB.png",
  position = xaringanExtra::css_position(top = "2em", right = "2em")
)
```
# Goals:

* Describe cases where the growth/decay is exponential. <br/><br/>
* Describe the criterium to decide if a model is exponential. <br/><br/>
* Describe the process to find an exponential model. <br/><br/>
* Use the exponential regression to predict values. <br/><br/>

---
# Exponential behavior

> Sometimes our data shows a behavior where it grows or decays everytime by the same factor. This growth/decay of our data cannot be controlled by us, it is just the nature of the data. <br/><br/> When this happens we say that our data has an **exponential behavior**. <br/><br/> In this case, we can use an **exponential model** to describe our data.

--
<br/><br/>
#### Example 

Consider the following data: <br/><br/>
| x | y |
|:-----------:|:------------:|
| 1   | 2   |
| 2   | 4   |
| 3   | 8   |
| 4   | 16  |
| 5   |  ?  |

---
# The exponential model:

> __Definition:__ The exponential function is a function of the form $f(x)=ab^x$, where $a$ and $b$ are constants and $b>0$. The relation between the dependent and independent variables is that the dependent one grows/decays by a factor of $b$ everytime the independent variable increases by one unit.

--
```{r, echo=FALSE, message=FALSE, warning=FALSE}
y<-runif(50,2,8)
plot(y,exp(y))
```
---
# Criterium

> Assume that we have some real world data, and we want to know if it has an exponential behavior. <br/><br/> We can use the following criterium to decide if our data has an exponential behavior: <br/><br/>
If three observations (with a unit of distance), then the ratio of the dependent variables is the same.

---
#### Example

Consider the following data: <br/><br/>
| $x$ | $y$ |
|:-----------:|:------------:|
| 1   | 5   |
| 2   | 10  |
| 3   | 20  |
| 4   | 40  |
| $n$ |  ?   |

---
# Finding the exponential model

> To find the exponential model we need to find the values of $a$ and $b$ in the function $f(x)=ab^x$. <br/><br/> To do this we need to use the following system of equations: <br/><br/>
$$\begin{cases} f(x_1)=ab^{x_1} \\ f(x_2)=ab^{x_2} \end{cases}$$ 

---
#### Example 

> The population of a city in 2010 was 100,000 people. If the rate of change of the population is 2.5% per year, find the exponential model that describes the population of the city.

--
<br/><br/>
#### Solution

We know that the population of the city in 2010 was 100,000 people, so we have the point $(0,100,000)$. <br/><br/> We also know that the rate of change of the population is 2.5% per year, so we have the point $(1,102,500)$. <br/><br/> Therefore, we have the following system of equations: <br/><br/>
$$\begin{cases} f(0)=ab^0=100,000 \\ f(1)=ab^1=102,500 \end{cases}$$

---
#### Example  

> The INEGI reported that in the month June 2019. The INPC (National Consumer Price Index) registered a montly increase of 0.06%. Assuming that in the following months, this index follows the same behavior. Find the exponential model that describes the INPC if at June this index was 100.00.

--
<br/><br/>
#### Solution

We know that the INPC in June 2019 was 100.00, so we have the point $(0,100.00)$. We also know that the INPC in July 2019 was 100.06, so we have the point $(1,100.06)$. Therefore, we have the following system of equations: <br/><br/>

$$\begin{cases} f(0)=ab^0=100.00 \\ f(1)=ab^1=100.06 \end{cases}$$

---
## Applications to business

> The compound interest is an example of an exponential model. <br/><br/> The compound interest is the interest that is added to the principal of a deposit or loan so that the added interest also earns interest from then on. <br/><br/> The formula to calculate the compound interest is: <br/><br/>
$$A=P\left(1+\frac{r}{n}\right)^{nt}$$ where $A$ is the amount of money accumulated after $t$ years, $P$ is the principal, $r$ is the annual interest rate, and $n$ is the number of times the interest is compounded per year.

---
#### Example

> A person deposits $\$1000$ in a bank account that pays 5% annual interest, compounded monthly. Find the amount of money accumulated after 10 years.

--
<br/><br/>
#### Solution

We know that the principal is $\$1000$, the annual interest rate is 5%, and the number of times the interest is compounded per year is 12. Therefore, we have the following information: <br/><br/>
$$\begin{cases} P=1000 \\ r=0.05 \\ n=12 \\ t=10 \end{cases}$$

---
#### Example

> A person deposits $\$1500$ in a bank account that pays 11.6% annual interest, compounded three-monthly. Find the amount of money accumulated after 5 years.

---
## Base e exponential model

> Suppose that you invest $\$1$ in a bank account that pays 100% annual interest, compounded $n$ times a year. <br/><br/> The amount of money accumulated after $t$ years is given by the formula: <br/><br/> $$A=\left(1+\frac{1}{n}\right)^{nt}$$

<br/><br/>
What happend if we increase the number of compounded times a year? Well, the number $(1+\frac{1}{n})$ will be closer to a number $e$.

---
## Base e exponential model

> The number $e$ is a number that is approximately equal to $2.7182818284590452353602874713527$. <br/><br/> The number $e$ is a very important number in mathematics, because it is the base of the natural logarithm. <br/><br/> The number $e$ is also the base of the exponential model that is used in many applications.

