---
title: "Triple Integrals"
subtitle: "Session 12"
author: 
  - "Alejandro Ucan"
date: '`r Sys.Date()`'
output:
  xaringan::moon_reader:
    css: xaringan-themer.css
    nature:
      slideNumberFormat: "%current%"
      highlightStyle: github
      highlightLines: true
      ratio: 16:9
      countIncrementalSlides: true
---

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
knitr::opts_chunk$set(
  fig.show =('asis'),
  dev='png',
	message = FALSE,
	warning = FALSE,
	cache = FALSE,
	hiline = TRUE,
	out.width = "90%"
)
```

```{r xaringan-themer, include=FALSE, warning=FALSE}
library(xaringanthemer)
style_duo_accent(
  primary_color = "#000080",
  secondary_color = "#f2f3f4",
  inverse_header_color = "#00147e"
)
```

```{r xaringan-editable, echo=FALSE}
xaringanExtra::use_editable(expires = 1)
xaringanExtra::use_scribble()
```
# Goals for the session:

 * Know the volume differential. <br/><br/>
 * Define the integral in $\mathbb{R}^3$. <br/><br/>
 * Calculate triple integrals in elementary regions.


---
# Triple Integral:

> The notion of triple integral is a generalization of the concept of integral to three-dimensional space. Using this, we can calculate the "volume" of a solid of four dimensions. The naive idea comes from the volume of a box, in fourth dimension, the box will have 4 different sides, and the volume will be the product of the four sides. Therefore in our case, $$f(x,y,z)dx dy dz$$ is the volume of a box-base with sides $dx, dy$ and $dz$.

---
# How to compute triple integrals?

Fubini's theorem is valid for triple integrals, so to calculate it, we just need to perform an iterated integration.

$${\int\!\int\!\int}_B fdV=\int_a^b \int_c^d \int_e^f f(x,y,z)dzdydx.$$

Recall that in the case of numerical limits, the change does not affect the result.

---
# Triple integral in elementary regions:

> **Definition:** an *elementary* region in $\mathbb{R}^3$ is a region in which one of the variables is bounded between two functions of the other two variables.

<br/><br/>

**Example:** The *ball* of radius one in $\mathbb{R}^3$ is an elementary region described by the following inequalities: $$-1\leq x\leq 1,$$
$$-\sqrt{1-x^2} \leq y\leq \sqrt{1-x^2},$$ 
$$ -\sqrt{1-x^2-y^2} \leq z \leq \sqrt{1-x^2-y^2}.$$

---
# Other Examples:

Other examples of elementary regions are:

* Consider the following cylinder: $$-1\leq x \leq 1$$ $$-\sqrt{1-x^2}\leq y \leq \sqrt{1-x^2}$$ $$ -2\leq z \leq 2.$$

* Consider the following prism: 
$$0 \leq x \leq 2$$ $$0\leq y \leq 2$$ $$0\leq z\leq 5-2x-2y$$

---
# Example 1:

> Find the volume of the solid enclosed by the function $f(x,y,z)=z$ over the region bounded by $x=0,\,y=0,\, z=0$ y $x+y+z=1$
--

$$R: 0\leq x\leq 1,\, 0\leq y \leq 1-x,\, 0\leq z\leq 1-x-y.$$

$$\int\int\int_R z dV= \int_0^1\int_0^{1-x}\int_0^{1-x-y} z dzdydx$$
---
# Example 2:

> Find the volume of the solid enclosed by the function $f(x,y,z)=z$ over the region bounde by the paraboloid $y=x^2+z^2$ and $y=4.$
--

$$E: -2\leq x\leq 2,\, -\sqrt{4-x^2}\leq z\leq \sqrt{4-x^2},\, x^2+z^2\leq y\leq 4$$

$$\int\int\int_E \sqrt{x^2+z^2}dV= \int_{-2}^2\int_{-\sqrt{4-x^2}}^{\sqrt{4-x^2}}\int_{x^2+z^2}^4  z dydzdx$$
---
What if we try with the following limits:

$$0\leq y\leq4,\, -\sqrt{y}\leq x \leq \sqrt{y},\, -\sqrt{y-x^2}\leq z \leq \sqrt{y-x^2}$$
---
#### Exercise 1:

Compute the following triple integrals:
  * $\int_0^1\int_0^z\int_0^{x+z}6xzdydxdz$
  * $\int_0^1\int_0^z\int_0^y ze^{-y^2}dxdydz$
  * $\int_0^{\sqrt{\pi}}\int_0^x\int_0^{xz}x^2\sin(y)dydzdx$


---
#### Exercise 2:

Draw the region $E$ and compute the integral of $f(x,y,z)=6.$ The region is given under the plane $z=1+x+y$ and above $z=0$ bounded by the curves $y=\sqrt{x},\, y=0$ y $x=1.$


