---
title: "Eigenvalues and Eigenvectors"
subtitle: "Sesion 05"
author: 
  - "Alejandro Ucan"
date: '`r Sys.Date()`'
output:
  xaringan::moon_reader:
    css: xaringan-themer.css
    nature:
      slideNumberFormat: "%current%"
      highlightStyle: github
      highlightLines: true
      ratio: 16:9
      countIncrementalSlides: true
---

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
knitr::opts_chunk$set(
  fig.width=9, fig.height=3.5, fig.retina=3,
  out.width = "100%",
  cache = FALSE,
  echo = FALSE,
  message = FALSE, 
  warning = FALSE,
  hiline = TRUE
)
```

```{r xaringan-themer, include=FALSE, warning=FALSE}
library(xaringanthemer)
style_duo_accent(
  primary_color = "#000080",
  secondary_color = "#f2f3f4",
  inverse_header_color = "#00147e"
)
```

```{r xaringan-editable, echo=FALSE}
xaringanExtra::use_editable(expires = 1)
xaringanExtra::use_scribble()
```

# Goals of the Session

* Define the concepts of Eigenvectors and Eigenvalues. <br/><br/>
* Calculate eigenvectors and eigenvalues. <br/><br/>

---
# Motivation

An assembly line with three workstations (A, B, C) has its production dynamics modeled by a linear transformation given by the following matrix $$T=\left(\begin{array}{ccc} 0.6 & 0.2 & 0.1 \\
    0.3 & 0.7 & 0.2 \\
    0.1 & 0.1 & 0.7 \\
\end{array}\right)$$ representing the relations between the workstations. Is there a steady production vector? <br/><br/> That is, is there a vector $v$ such that $$T(v)=v$$?

 
---
# Eigenvectors and Eigenvalues

> **Definition:** Given a linear transformation $T:V\to W$, a vector $v\in V$ is called an *eigenvector* of $T$ if $v\neq 0$ and there exists a scalar $\lambda$ such that $T(v)=\lambda v$. The scalar $\lambda$ is called an *eigenvalue* of $T$ associated to the eigenvector $v$.

---
## Properties of Eigenvectors and Eigenvalues

1. The pair of eigenvector-eigenvalue acts as a "dynamical system", i.e., the eigenvector is a direction in which the transformation acts by stretching or compressing by the factor of the eigenvalue. <br/><br/>
2. If $v$ is an eigenvector of $T$ with eigenvalue $\lambda$, then $T(T(v))=\lambda T(v)=\lambda^2 v$. <br/><br/>
3. If all eigenvalues are real and unique, then the eigenvectors are linearly independent, therefore a **basis** of the vector space. 

---
##### Example 1:

Let $T(x,y)=(2x,-y)$, then $T(1,0)=(2,0)$ and $T(0,1)=(0,-1)$. Therefore, $T$ has eigenvectors $(1,0)$ and $(0,1)$ with eigenvalues $\lambda_1=2$ and $\lambda_2=-1$ respectively.


---
## How to compute eigenvectors and eigenvalues?

Note that if $v$ is an eigenvector of $T$ with eigenvalue $\lambda$ and $M$ is a matrix that represents $T$ then it must satisfy that $$Mv=\lambda v$$ and therefore $$(M-\lambda I)v=0$$ where $I$ is the identity matrix and this is for every eigenvector. <br/><br/> Therefore, the system of equations must have infinite solutions and this only happens if $$\det(M-\lambda I)=0.$$

---
## The Characteristic Polynomial

> **Definition:** Given a square matrix $M$, the _characteristic polynomial_ of $M$ is the polynomial in variable $\lambda$ obtained from $$\det(M-\lambda I).$$

--
<br/><br/>

> **Claim:** If $\lambda_j$ is a root of the characteristic polynomial of $M$ then $\lambda_j$ is an eigenvalue.

---
##### Example 3:

> Compute the characteristic polynomial and its roots of the transformation $T(x,y)=(3x-2y,x+2y)$.

--
<br/><br/>

##### Solution

$$\det\left(\begin{array}{cc} 3-\lambda & 2 \\ 1 & 2-\lambda \end{array}\right)=(3-\lambda)(2-\lambda)-4=\lambda^2-5\lambda+4$$

and their solutions are: $$\lambda_{1,2}=\frac{5\pm\sqrt{5^2-4(4)}}{2}$$ where $$\lambda_1=\frac{5-\sqrt{9}}{2}\quad \lambda_2=\frac{5+\sqrt{9}}{2}$$

---
##### Example 4:

> Compute the characteristic polynomial and its roots of the transformation $T(x,y,z)=(0,x+2y-7z,-2x-3y-2z)$.

--
<br/><br/>

##### Solution:
$$\det\left(\begin{array}{ccc} 0-\lambda & 0 & 0 \\ 1 & 2-\lambda & -7 \\ -2 & -3 & -2-\lambda \end{array}\right)$$. <br/><br/> Solving the determinant we obtain $$\lambda^3-25\lambda$$ whose roots are $$\lambda_1=0,\quad \lambda_2=5,\quad \lambda_3=-5$$

---
## Find the eigenvectors

> **Claim:** If $\lambda_j$ is an eigenvalue of the transformation $T$, then any $v$ that is in the kernel of $T-\lambda_j I$ is an eigenvector of $T$ with eigenvalue $\lambda_j$. 

--
##### Example 5:

> We know that $1$ is an eigenvalue of the transformation $T(x,y)=(3x-2y,x+2y)$. Find an eigenvector associated to this eigenvalue.

--
##### Solution:

Matrix's kernel $T-1I$ is $$\left(\begin{array}{cc} 2 & 2 \\ 1 & 1 \end{array}\right)$$ and therefore, the vector of the form $(1,-1)$ is an eigenvector associated to $1.$

---
##### Example 6:

> We know that $5$ is an eigenvalue of the transformation $T(x,y,z)=(0,x+2y-7z,-2x-3y-2z)$. Find an eigenvector associated to this eigenvalue.

--
<br/><br/>

##### Solution

The kernel of the matrix $T-5I$ is $$\left(\begin{array}{ccc} -5 & 0 & 0 \\ 1 & -3 & -7 \\ -2 & -3 & -7 \end{array}\right)$$ and therefore, the vector of the form $(0,7,3)$ is an eigenvector associated to $5.$

---
### Applied Eigenvalues and Eigenvectors

An assembly line with three workstations (A, B, C) has its production dynamics modeled by a linear transformation given by the following matrix $$T=\left(\begin{array}{ccc} 0.6 & 0.2 & 0.1 \\
    0.3 & 0.7 & 0.2 \\
    0.1 & 0.1 & 0.7 \\
\end{array}\right)$$ representing the relations between the workstations. Is there a steady production vector? <br/><br/> That is, is there a vector $v$ such that $$T(v)=v$$?

---
### Applied Eigenvalues and Eigenvectors

A vibration control system for an industrial robot arm is modeled by the linear transformation given by the matrix $$A=\left(\begin{array}{ccc} 4 & 1 & 0 \\
    0 & 3 & -1 \\
    0 & 2 & 0 \\
\end{array}\right)$$, where the transformation matrix A represents the system's dynamics. Find if there is directions in which the system grows or shrinks exponentially, and if so, find the eigenvalues and eigenvectors associated to those directions.
