---
title: "Higher Order Linear Differential Equations with constant coefficients"
subtitle: "Session 03"  
author: 
  - "Alejandro Ucan"
date: '`r Sys.Date()`'
output:
  xaringan::moon_reader:
    css: xaringan-themer.css
    nature:
      slideNumberFormat: "%current%"
      highlightStyle: github
      highlightLines: true
      ratio: 16:9
      countIncrementalSlides: true
---

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
knitr::opts_chunk$set(
  fig.width=9, fig.height=3.5, fig.retina=3,
  out.width = "80%",
  cache = FALSE,
  echo = FALSE,
  message = FALSE, 
  warning = FALSE,
  hiline = TRUE
)
```

```{r xaringan-themer, include=FALSE, warning=FALSE}
library(xaringanthemer)
style_duo_accent(
  primary_color = "#000080",
  secondary_color = "#f2f3f4",
  inverse_header_color = "#00147e"
)
```

```{r xaringan-editable, echo=FALSE}
xaringanExtra::use_editable(expires = 1)
xaringanExtra::use_scribble()
```

```{r xaringan-logo, echo=FALSE}
xaringanExtra::use_logo(
  image_url = "https://github.com/alxcn/TecLogoEIC/blob/9562a53875418e749a296c85808a19c85fc4be74/IngenieriaCiencias_Horizontal_RGB.png",
  position = xaringanExtra::css_position(top = "2em", right = "2em")
)
```
# Goals

* Describe the general form of a linear homogeneous differential equation with constant coefficients and the IVP. <br/><br/>
* Determine the general solution for a linear homogeneous differential equation with constant coefficients. <br/><br/>
* Deduce the method to solve linear homogeneous differential equations with constant coefficients. <br/><br/>
* Apply the method to solve linear homogeneous differential equations with constant coefficients.

---
# Linear homogeneous differential equations with constant coefficients.

> __Definition:__ the generic form of a linear homogeneous differential equation of order $n$ is: $$a_n(x)y^{(n)}+a_{n-1}(x)y^{(n-1)}+\cdots+a_2(x)y''+a_1(x)y'+a_0(x)y=g(x).$$

To define an IVP we need to add the initial conditions: $$y(x_0)=y_0,\, y'(x_0)=y_1,\,y''(x_0)=y_2,\cdots,\,y^{(n-1)}(x_0)=y_{n-1}.$$

---
#### Example 1:

> For the $3y'''+7y''-7y'-3y=0$ we have that $y(x)=ce^{x}$ and $y(x)=ce^{-x/3}+ke^{-3x}$ are solutions. <br/><br/>

--
<br/><br/>

> The function $y=3e^{2x}+e^{-2x}-3x$ is the solution for the IVP $y''-4y=12x$ with initial conditions $y(0)=4$ and $y'(0)=1.$ 

---
# First Order LODE with cc.

> The general form of a first order LODE with cc is $ay'+by=g(x),$ lets try to solve the homogeneous case. <br/><br/>
  * The solution have to be of the form $y=e^{mx}.$ <br/><br/>
  * Substituting in the LODE we have that $$ame^{mx}+be^{mx}=0=e^{mx}(am+b),$$ 
  * From the fact that $e^{mx}\neq 0$ we have that $$am+b=0\Rightarrow m=-b/a.$$
  * Therefore the solution is $y(x)=ce^{-bx/a}.$
---
# Second Order LODE with cc.

> Lets take the second order homogeneous LODE with cc. <br/><br/>
  * The solution have to be of the form $y=e^{mx}.$ <br/><br/>
  * Substituting in the LODE we obtain that our __auxiliary polynomial__ is $$am^2+bm+c=0.$$ <br/><br/>
  * We have that there two roots to this polynomio (they might be egual): $m_1$ and $m_2.$ <br/><br/>
  * Remember that if:
    * If $\Delta=b^2-4ac>0$ then the roots are real and different.
    * If $\Delta=0$ then the roots are real and equal.
    * If $\Delta<0$ then the roots are complex conjugates.
    
---
## First Case: Different real roots:

> In the case the roots are real and different, $m_1$ and $m_2,$ then the general solution is $$y=c_1 e^{m_1 x}+c_2e^{m_2 x}.$$

--
<br/>
#### Example 1: 

Find the general solution to $2y''-5y'-3y=0,$ 

--
<br/>
#### Solution:
* The auxiliary polynomio is $2m^2-5m-3=0$ <br/><br/>
* The roots are $m_1=3$ and $m_2=-1/2$ <br/><br/>
* Therefore, the general solution is: $$y=c_1e^{3x}+c_2e^{-x/2}.$$
---
## Second Case: Equal real roots:

> In the case the roots are real and equal, $m_1=m_2=m,$ then the general solution is $$y=c_1 e^{mx}+c_2xe^{mx}.$$

--
<br/>
#### Example 2:
Find the general solution to $y''-10y'+25y=0.$

--
<br/>
#### Solution:
* The auxiliary polynomio is $m^2-10m+25=0$ <br/><br/>
* The roots are $m_1=m_2=5$ <br/><br/>
* Therefore, the general solution is: $$y=c_1e^{5x}+c_2xe^{5x}.$$

---
## Third Case: Complex conjugate roots:
> In the case the roots are complex conjugates, $m_1=a+bi$ and $m_2=a-bi,$ then the general solution is $$y=c_1 e^{a}\cos(bx)+c_2e^{ax}\sin(bx).$$

--
<br/>
#### Example 3:
Find the general solution to $y''+4y'+7y=0.$

--
<br/>
#### Solution:
* The auxiliary polynomio is $m^2+4m+7=0$ <br/><br/>
* The roots are $m_1=-2+\sqrt{3}i$ and $m_2=-2-\sqrt{3}i.$ <br/><br/>
* Therefore, the general solution is: $$y=c_1e^{-2x}\cos(\sqrt{3}x)+c_2e^{-2x}\sin(\sqrt{3}x).$$
---
#### Example 4:

Find the solution of $4y''+4y'+17y=0$ if $y(0)=-1$ y $y'(0)=2.$

--
<br/>
#### Solution:
* The auxiliary polynomio is $4m^2+4m+17=0$ <br/><br/>
* The roots are $m_1=-\frac{1}{2}-2i$ and $m_2=-\frac{1}{2}+2i.$ <br/><br/>
* Therefore, the general solution is: $$y=c_1e^{-\frac{1}{2}x}\cos(2x)+c_2e^{-\frac{1}{2}x}\sin(2x).$$
* Applying the initial condition we have the following system of equations: $$\begin{cases} c_1=-1 \\ -\frac{1}{2}c_1+2c_2=2 \end{cases}$$
whose solution are $c_1=-1$ and $c_2=1.$ Therefore the solution is $$y=-e^{-\frac{1}{2}x}\cos(2x)+e^{-\frac{1}{2}x}\sin(2x).$$

