---
title: "Cualitative Analysis and Applications"
subtitle: "Session 06"
author: 
  - "Alejandro Ucan"
date: '`r Sys.Date()`'
output:
  xaringan::moon_reader:
    css: xaringan-themer.css
    nature:
      slideNumberFormat: "%current%"
      highlightStyle: github
      highlightLines: true
      ratio: 16:9
      countIncrementalSlides: true
---

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
knitr::opts_chunk$set(
  fig.width=9, fig.height=3.5, fig.retina=3,
  out.width = "100%",
  cache = FALSE,
  echo = FALSE,
  message = FALSE, 
  warning = FALSE,
  hiline = TRUE
)
```

```{r xaringan-themer, include=FALSE, warning=FALSE}
library(xaringanthemer)
style_duo_accent(
  primary_color = "#000080",
  secondary_color = "#f2f3f4",
  inverse_header_color = "#00147e"
)
```

```{r xaringan-editable, echo=FALSE}
xaringanExtra::use_editable(expires = 1)
xaringanExtra::use_scribble()
```

```{r xaringan-logo, echo=FALSE}
xaringanExtra::use_logo(
  image_url = "https://github.com/alxcn/TecLogoEIC/blob/9562a53875418e749a296c85808a19c85fc4be74/IngenieriaCiencias_Horizontal_RGB.png",
  position = xaringanExtra::css_position(top = "2em", right = "2em")
)
```
# Goals

* Deduction of the Tangent line and relation with the derivative. <br/><br/>
* Deduction of the Normal line and relation with the derivative. <br/><br/>
* Monotonicity of a function. <br/><br/>
* Concavity of a function. <br/><br/>

---
# Motivation

Assume that we are following a circular path really fast, and suddenly we make an stop. What will happen to us? What path will be following by the law of physics?

Let us translate this problem for us. If our position is given by a function $d(t),$ then our mean velocity is given by the function $$v(t)=\frac{v(t)-v(t_0)}{t-t_0}.$$

---
# Motivation

Now, if we reduce the time interval to zero, then we have the instantaneous velocity 

$$v(t)=\lim_{t\to t_0}\frac{v(t)-v(t_0)}{t-t_0}.$$


---
# Tangent Line

> **Definition:** Given a function $f$ and a point $P(a,f(a))$ in the graph of $f$, the tangent line to the graph of $f$ at the point $P$ is the line that passes through $P$ and has slope $f'(a)$.

-- 

#### Remarks

1. The tangent line to the graph of $f$ at the point $P(a,f(a))$ is given by the equation $y=f(a)+f'(a)(x-a).$
2. The tangent line to the graph of $f$ at the point $P(a,f(a))$ is the best linear approximation to the graph of $f$ near the point $P.$
3. The tangent line is the first order Taylor polynomial of $f$ at the point $P.$

---
# Examples

1. Find the equation of the tangent line to the graph of $f(x)=x^2$ at the point $P(1,1).$ <br/><br/>

--

#### Solution

We have that $f'(x)=2x,$ then $f'(1)=2.$ Therefore, the equation of the tangent line is $y=1+2(x-1)=2x-1.$ <br/><br/>

---
# Examples

2. Without calculator find the approximate value of $\sqrt{16.1}.$

--
#### Solution

We have that $f(x)=\sqrt{x}$ and $f'(x)=\frac{1}{2\sqrt{x}}.$ Then, $f'(16)=\frac{1}{2\sqrt{16}}=\frac{1}{8}.$ Therefore, the equation of the tangent line is $y=4+\frac{1}{8}(x-16)=\frac{1}{8}x+3.$ Then, $\sqrt{16.1}\approx \frac{1}{8}(16.1)+3=4.0125.$

---
# Normal line

> **Definition:** Given a function $f$ and a point $P(a,f(a))$ in the graph of $f$, the normal line to the graph of $f$ at the point $P$ is the line that passes through $P$ and is perpendicular to the tangent line to the graph of $f$ at the point $P.$

---
# Cualitative Analysis 

## Monotonicity

> **Definition:** A function $f$ is said to be increasing in an interval $I$ if for all $x_1,x_2$ in $I$ such that $x_1<x_2,$ then $f(x_1)<f(x_2).$ A function $f$ is said to be decreasing in an interval $I$ if for all $x_1,x_2$ in $I$ such that $x_1<x_2,$ then $f(x_1)>f(x_2).$

-- 
#### Derivatives and Monotonicity

1. If $f'(x)>0$ for all $x$ in an interval $I,$ then $f$ is increasing in $I.$
2. If $f'(x)<0$ for all $x$ in an interval $I,$ then $f$ is decreasing in $I.$

---
#### Examples

Describe the Monotonicity of the following functions.

1. $f(x)=3x+$
2. $f(x)=x^3-x^2+x-1$
3. $f(x)=\cos(x).$

---
# Cualitative Analysis

## Concavity

> **Definition:** A function $f$ is said to be concave up in an interval $I$ if for all $x_1,x_2$ in $I$ such that $x_1<x_2,$ then the graph of $f$ lies above the tangent line to the graph of $f$ at the point $(x_1,f(x_1)).$ A function $f$ is said to be concave down in an interval $I$ if for all $x_1,x_2$ in $I$ such that $x_1<x_2,$ then the graph of $f$ lies below the tangent line to the graph of $f$ at the point $(x_1,f(x_1)).$

--
#### Derivatives and Concavity

1. If $f''(x)>0$ for all $x$ in an interval $I,$ then $f$ is concave up in $I.$
2. If $f''(x)<0$ for all $x$ in an interval $I,$ then $f$ is concave down in $I.$

---

#### Examples

Describe the Concavity of the following functions.

1. $f(x)=3x^2$
2. $f(x)=x^3-x^2+x-1$
3. $f(x)=\cos(x).$

---
### Examples

Describe the behavior of the following functions:

1. $f(x)=x^3-x^2+x-1$
2. $f(x)=\sqrt{3x+1}$
3. $f(x)=x^3-3x^2+3x-1$

