---
title: "Applied Laplace Transform"
subtitle: "Session 10"
author: 
  - "Alejandro Ucan"
date: '`r Sys.Date()`'
output:
  xaringan::moon_reader:
    css: xaringan-themer.css
    nature:
      slideNumberFormat: "%current%"
      highlightStyle: github
      highlightLines: true
      ratio: 16:9
      countIncrementalSlides: true
---

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
knitr::opts_chunk$set(
  fig.width=9, fig.height=3.5, fig.retina=3,
  out.width = "100%",
  cache = FALSE,
  echo = FALSE,
  message = FALSE, 
  warning = FALSE,
  hiline = TRUE
)
```

```{r xaringan-themer, include=FALSE, warning=FALSE}
library(xaringanthemer)
style_duo_accent(
  primary_color = "#000080",
  secondary_color = "#f2f3f4",
  inverse_header_color = "#00147e"
)
```

```{r xaringan-editable, echo=FALSE}
xaringanExtra::use_editable(expires = 1)
xaringanExtra::use_scribble()
```

```{r xaringan-logo, echo=FALSE}
xaringanExtra::use_logo(
  image_url = "https://github.com/alxcn/TecLogoEIC/blob/9562a53875418e749a296c85808a19c85fc4be74/IngenieriaCiencias_Horizontal_RGB.png",
  position = xaringanExtra::css_position(top = "2em", right = "2em")
)
```
# Goals


* Introduce the inverse of a Laplace transform. <br/><br/>
* Introduce the relationship between the Laplace transform and the derivative of a function. <br/><br/>
* Solve initial value problems using Laplace transforms. <br/><br/>

---
# Most commun Laplace transforms:

The next list has the Laplace transforms for the most commun functions

  * $\mathcal{L}\left\{t^n\right\}=\frac{n!}{s^{n+1}}$ 
  * $\mathcal{L}\left\{e^{at}\right\}=\frac{1}{s-a}$
  * $\mathcal{L}\left\{\sin(kt)\right\}=\frac{k}{s^2+k^2}$
  * $\mathcal{L}\left\{\cos(kt)\right\}=\frac{s}{s^2+k^2}$
  * $\mathcal{L}\left\{\sinh(kt)\right\}=\frac{k}{s^2-k^2}$
  * $\mathcal{L}\left\{\cos(kt)\right\}=\frac{s}{s^2-k^2}$
  
---
# Inverse Laplace Transform 

> Suppose that we know the Laplace transform of an unknown function, how to determine the function that produces the Laplace transform? <br/><br/>
> __Definition:__ The inverse Laplace transform of $F(s)$ is the function $f(t)$ such that $$\mathcal{L}\{f(t)\}=F(s).$$
  
--
<br/><br/>  
  * $F(s)=\frac{1}{s^5}$
  * $F(s)=\frac{1}{s^2+7}$
  * $F(s)=\frac{s^2+6s+9}{(s-1)(s-2)(s+4)}$
  
---
# Laplace transform of a Derivative:

> __Theorem:__ If $\mathcal{L}\{f(t)\}=F(s)$ exists, then $$\mathcal{L}\{f'(t)\}=sF(s)-f(0)$$ $$\mathcal{L}\{f^{(n)}(t)\}=s^nF(s)-s^{n-1}f(0)-s^{n-2}f'(0)-\cdots- f^{(n-1)}(0).$$

--
<br/><br/>
### Example: 

Find the solution of the initial value problem $y'+3y=13\sin(2t)\quad y(0)=6.$

---
#### Solution:

1. We apply Laplace transform in both sides of the equation:
$$\mathcal{L}(y')+3\mathcal{L}(y)=\mathcal{13\sin(2t)}$$
2. We use derivative theorem to express the transform of the derivative.
$$s^2 \mathcal{L}(y)-sy(0)=\frac{26}{s^2+4}$$
3. We solve for $\mathcal{L}(y)$ $$(s^2-6s)\mathcal{L}(y)=\frac{26}{s^2+4}\Rightarrow \mathcal{L}(y)=\frac{26}{s(s-6)(s^2+4)}$$

---
#### Solution:

4. We use partial fractions to express $\mathcal{L}(y)$ as a sum of simpler fractions 
$$\mathcal{L}(y)=\frac{13 (3 s - 2)}{40 (s^2 + 4)} + \frac{13}{120 (s - 6)} - \frac{13}{12 s}$$

5. Then we use the table of Laplace transforms to find the inverse transform. 

$$y(t)=-\frac{13}{12}+\frac{13}{120}e^{6t}+\frac{13}{40}\left(3\cos(2t)-\frac{1}{2}\sin(2t)\right)$$

---
### Example

Find the solution of the following initial value problem $$y''-3y'+2y=e^{-4t},\quad y(0)=1,\quad y'(0)=5.$$


---
### Example

Find the solution of the following initial value problem $$2y'''+3y''-3y'-2y=e^{-t},\, y(0)=0,\,y'(0)=0,\,y''(0)=1.$$

---
