---
title: "Lagrange Multipliers"
subtitle: "Session 08"  
author: 
  - "Alejandro Ucan"
date: '`r Sys.Date()`'
output:
  xaringan::moon_reader:
    css: xaringan-themer.css
    nature:
      slideNumberFormat: "%current%"
      highlightStyle: github
      highlightLines: true
      ratio: 16:9
      countIncrementalSlides: true
---

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
knitr::opts_chunk$set(
  fig.width=9, fig.height=3.5, fig.retina=3,
  out.width = "100%",
  cache = FALSE,
  echo = FALSE,
  message = FALSE, 
  warning = FALSE,
  hiline = TRUE
)
```

```{r xaringan-themer, include=FALSE, warning=FALSE}
library(xaringanthemer)
style_duo_accent(
  primary_color = "#000080",
  secondary_color = "#f2f3f4",
  inverse_header_color = "#00147e"
)
```

```{r xaringan-editable, echo=FALSE}
xaringanExtra::use_editable(expires = 1)
xaringanExtra::use_scribble()
```
# Goals for the session:

* Understand the necessary and sufficient conditions for the existence of extreme values in problems with restriction. <br/><br/>
* Find these values using the Lagrange multiplier method.
 
---
# Restrictions?


> A rectangular box without a lid is made from $12m^2$ of cardboard. Find the maximum volume of the box.
<br/>
--

Note that we have a restriction on the variables and it is that they satisfy $$A_{sup}=2xz+2zy+xy=12.$$

---
# Intuitively

We want to find the extreme values of a function $f(x,y)$ such that the variables satisfy a restriction that can be expressed as a level curve of a function $g(x,y).$ <br/><br/>
We want to find the $(x,y)$ such that $$f(x,y)\mbox{ has an extreme value}$$ $$g(x,y)=k.$$

---
## Visually


<iframe src="https://www.geogebra.org/classic/uebb3j98?embed" width="1200" height="450" allowfullscreen style="border: 1px solid #e4e4e4;border-radius: 4px;" frameborder="0"></iframe>

---
## The multiplier


> **Definition:** Given the conditions of the previous drawing, we will say that $\lambda$ is a **Lagrange multiplier** if it satisfies that $$\nabla f(x_0,y_0,z_0)=\lambda \nabla g(x_0,y_0,z_0).$$

---
# The Lagrange multiplier method

> To find the extreme values of a function $f(x,y,z)$ such that the variables satisfy a restriction that can be expressed as a level curve of a function $g(x,y,z).$ We can use the following steps:

  * Find the values $x,y,z$ and $\lambda$ such that $$\nabla f(x,y,z)=\lambda \nabla g(x,y,z).$$
  * Evaluate $f$ at these points, the largest of these numbers is the maximum and the smallest the minimum.
  
---
# Example 1:

> Find the extreme values of the function $f(x,y)=5x^2+6y^2-xy$ under the restriction $x+2y=24.$

--
1. Compute the gradient of the objective function: $$\nabla f= (10x-y,12y-x)$$
1. Compute the gradient of the restriction: $$\nabla g =(1,2)$$
1. Find the values of $x,y$ and $\lambda$ such that $\nabla f(x,y)=\lambda \nabla g(x,y).$
 $$10x-y=\lambda$$ $$12y-x=\lambda$$ $$x+2y=24$$
1. Solve the system of equations.
---



# Example 1: 

5.Explain my results $$y=-\frac{-288}{25},\quad x=\frac{576}{5}.$$


---
# Example 2:


> The Cobb-Douglas production function for a production is given by $f(x,y)=100x^{3/4}y^{1/4}.$ Where $x$ are the units of production (at 150mxn per unit) and $y$ are the units of capital (at 250mxn per unit). If our production and capital units are restricted to 50000mxn. Find the maximum production for this manufacturing.
---

# Example 3:

>  BioTech is a manufacturer of prothesis and has developed a model of earnings that depends on the number of prothesis ($x$) that sells per month (in thousands) and the number of hours per month that invests in advertising ($y$). $$f(x,y)=48x+96y-x^2-2xy-9y^2.$$ Our budget for production of prothesis and advertising is given by $20x+4y=216.$ Find the number of prothesis and hours of advertising that maximize my earnings.

---
# Example 4:

> Find the extreme values for the function $f(x,y,z)=2x^2+y^2+3z^2$ under the restriction $2x-3y-4z=49.$
