---
title: "Generating Sets and Dimension"
subtitle: "Session 02" 
author: 
  - "Alejandro Ucan"
date: '`r Sys.Date()`'
output:
  xaringan::moon_reader:
    css: xaringan-themer.css
    nature:
      slideNumberFormat: "%current%"
      highlightStyle: github
      highlightLines: true
      ratio: 16:9
      countIncrementalSlides: true
---

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
knitr::opts_chunk$set(
  fig.width = 9, fig.height = 3.5, fig.retina = 3,
  out.width = "100%",
  cache = FALSE,
  echo = FALSE,
  message = FALSE, 
  warning = FALSE,
  hiline = TRUE
)
```

```{r xaringan-themer, include=FALSE, warning=FALSE}
library(xaringanthemer)
style_duo_accent(
  primary_color = "#000080",
  secondary_color = "#f2f3f4",
  inverse_header_color = "#00147e"
)
```

```{r xaringan-editable, echo=FALSE}
xaringanExtra::use_editable(expires = 1)
xaringanExtra::use_scribble()
```

```{r xaringan-logo, echo=FALSE}
xaringanExtra::use_logo(
  image_url = "https://github.com/alxcn/TecLogoEIC/blob/9562a53875418e749a296c85808a19c85fc4be74/IngenieriaCiencias_Horizontal_RGB.png",
  position = xaringanExtra::css_position(top = "2em", right = "2em")
)
```
# Goals of the Session:

* Introduce linear combinations. <br/><br/>
* Distinguish between linear dependence and independence. <br/><br/>
* Introduce generated spaces. <br/><br/>
* Define the concept of dimension. <br/><br/>


---
# Linear Combinations.

> **Definition:** Let $v\in V$ be a vector. If $u_1,\cdots,u_n\in V$ are vectors, we say that $v$ is a _linear combination_ of $u_1,\cdots,u_n$ if there exist scalars $a_1,\cdots,a_n$ such that $$v=a_1u_1+\cdots+a_nu_n.$$ <br/><br/>These scalars $a_1,\cdots,a_n$ are called _coefficients_ of the linear combination.


##### Example 1
In $\mathbb{R}^2,$ the vector $(3,4)$ has the following linear combinations:
1. $(3,4)=1(3,0)+4(0,1)$
1. $(3,4)=2(1,2)+1(1,0)$
1. $(3,4)=1(1,2)+2(1,1)$

--
##### Example 2:

In other vector spaces we can find linear combinations like these:
1. In $M(2\times 2, \mathbb{R}),$ the vector $\begin{pmatrix}1 & 0 \\ 0 & 1\end{pmatrix}$ has the linear combination $\begin{pmatrix}1 & 0 \\ 0 & 1\end{pmatrix}=1\begin{pmatrix}1 & 0 \\ 0 & 0\end{pmatrix}+1\begin{pmatrix}0 & 0 \\ 0 & 1\end{pmatrix}.$ <br/><br/>
1. In $P_2(\mathbb{R}),$ the polynomial $p(x)=x^2$ has the linear combination $x^2=1(x^2)+0(x)+0(1).$


---
## Determine a L.C.

> Let $v=(1,1,1)$ and $S=\{(1,2,3),(0,1,2),(-1,0,1)\}.$ Find a l.c. of $v$ in terms of the vectors in $S.$ <br/><br/>

--

$$v=a_1(1,2,3)+a_2(0,1,2)+a_3(-1,0,1)$$ <br/>
$$v=(a_1-a_3,2a_1+a_2,3a_1+2a_2+a_3)$$ <br/>
Using Gauss-Jordan in the matrix $$\left(\begin{array}{ccc|c}1 & 0 & -1 & 1 \\ 2& 1 & 0 & 1 \\ 3 & 2 & 1 & 1\end{array}\right)$$ we get $a_1=2, \,a_2=-3,$ and $a_3=1.$ Therefore, $$v=2(1,2,3)-3(0,1,2)+(-1,0,1).$$


---
## Generating Sets

> **Definition:** Let $S=\{u_1,\cdots,u_n\}$ be a set of vectors in $V.$ The _space generated_ by $S,$ denoted by $\langle S\rangle,$ is the set of all linear combinations of $u_1,\cdots,u_n.$ <br/><br/>That is, $$\langle S\rangle=\{a_1u_1+\cdots+a_nu_n\mid a_1,\cdots,a_n\in\mathbb{R}\}.$$ <br/><br/> $S$ is known as a generating set.
--

##### Example

The set $S=\{(1,0,0),(0,1,0),(0,0,1)\}$ is a generating set of $\mathbb{R}^3$ and $S=\{(1,1),(1,-1)\}$ is a generating set of $\mathbb{R}^2.$ The set $S=\{1,x,x^2\}$ is a generating set of $P_2(\mathbb{R}).$

---
##### Counterexamples

> The set $S=\{(1,2),(-2,-4)\}$ does not generate $\mathbb{R}^2.$ We cannot find a l.c. of $S$ for $(0,1).$<br/><br/><br/><br/>


> The set $S=\{(1,2,3),(0,1,2),(-1,0,1)\}$ does not generate $\mathbb{R}^3.$ This is because the vector $(1,-2,2)$ is not a l.c. of $S.$

---
## Generated Vector Subspaces

> **Theorem:** Let $S=\{u_1,\cdots,u_n\}$ be a set of vectors in $V.$ Then $\langle S\rangle$ is a vector subspace of $V$ and is the smallest vector subspace containing $S.$

---
# Linear Dependence and Independence

> **Definition:** Let $S=\{u_1,\cdots,u_n\}$ be a set of vectors in $V.$ We say that $S$ is _linearly dependent_ if there exists a linear combination of $u_1,\cdots,u_n$ that is equal to the zero vector and at least one of the coefficients of the linear combination is different from zero. <br/><br/>If $S$ is not linearly dependent, we say that $S$ is _linearly independent_.

--
##### Example

The set $S=\{(1,1),(1,-1)\}$ is l.i. but $S=\{(1,2,3),(0,1,2),(-1,0,1)\}$ is not.

---
## Linear Dependence and Independence

> **Theorem:** Let $S=\{u_1,\cdots,u_n\}$ be a set of vectors in $V.$ Then $S$ is linearly dependent if and only if at least one of the vectors in $S$ is a linear combination of the other vectors in $S.$ Even more, if we can find a squeare matrix with the vectors in $S$ and the determinant is non-zero, then $S$ is linearly dependent. 


---
# Dimension

> **Definition:**  We say that $S$ is a basis for the vector space $V,$ if $S$ is l.i. and generates $V.$
--
##### Example

For example, $\{(0,1),(1,0)\},$ $\{(1,1),(1,0)\}$ and $\{(1,1),(-1,1)\}$ are bases of $\mathbb{R}^2.$ And $\{(1,1,0),(0,1,0),(0,1,1,)$ is a basis of $\mathbb{R}^3.$

---
## Dimension


> **Theorem:** If $S$ is a basis of $V$ with $n-$elements. Then any subset of vectors with $n+1$ elements is linearly dependent. <br/><br/>Moreover, if $S$ is a basis of $V$ with $n-$elements. Then any subset of vectors with $n-1$ elements does not generate $V.$

--

> **Definition:** If $V$ has a basis with $n-$elements, then we say that $V$ has _dimension_ $n.$

---
# Activity


1. Let $S=\{(1,1,1),(1,2,3),(1,3,6)\}.$ Prove that $S$ is l.i. and generates $\mathbb{R}^3.$
1. Find a base for the line $L=\{(x,3x)\in\mathbb{R}^2:x\in\mathbb{R}\}.$ What dimension should it have?
1. Find a base for the vector space $P_3(\mathbb{R}).$ What dimension should it have?
1. If $S=\{(1,1,1),(1,0,1),(1,a,1)\}$ is l.i., what value should $a$ have?
