---
title: "Gaussian Reduction"
subtitle: "Session 03"  
author: 
  - "Alejandro Ucan"
date: '`r Sys.Date()`'
output:
  xaringan::moon_reader:
    css: xaringan-themer.css
    nature:
      slideNumberFormat: "%current%"
      highlightStyle: github
      highlightLines: true
      ratio: 16:9
      countIncrementalSlides: true
---

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
knitr::opts_chunk$set(
  fig.width=9, fig.height=3.5, fig.retina=3,
  out.width = "80%",
  cache = FALSE,
  echo = FALSE,
  message = FALSE, 
  warning = FALSE,
  hiline = TRUE
)
```

```{r xaringan-themer, include=FALSE, warning=FALSE}
library(xaringanthemer)
style_duo_accent(
  primary_color = "#000080",
  secondary_color = "#f2f3f4",
  inverse_header_color = "#00147e"
)
```

```{r xaringan-editable, echo=FALSE}
xaringanExtra::use_editable(expires = 1)
xaringanExtra::use_scribble()
```

```{r xaringan-logo, echo=FALSE}
xaringanExtra::use_logo(
  image_url = "https://github.com/alxcn/TecLogoEIC/blob/9562a53875418e749a296c85808a19c85fc4be74/IngenieriaCiencias_Horizontal_RGB.png",
  position = xaringanExtra::css_position(top = "2em", right = "2em")
)
```
# Goals

* Introduce the Gaussian elimination algorithm.
* Solve systems of linear equations (SLE) using the Gaussian elimination method.
* Rank of a Matrix and Theorem of Rouché–Capelli.

---
# Motivation

> Previously, we learned how to solve Systems of Linear Equations (SLE) using the reduction method, but easily we encountered the following problems: <br/><br/>
1. *Handling with lots of symbols:* leads to mistakes. <br/>
2. *Inneficient:* for large systems, it is not practical to solve them by hand. <br/>

How can we do more efficiently this process?

---
# Main Idea:

> The symbols of the variables are not necessary to solve the system, therefore we can use the matrix representation of the system to apply the reduction method. <br/>
This process is called **Gaussian elimination**. <br/>


---
# Elemental Operations


* **Row Interchange**: Two rows are swapped: $R_i \leftrightarrow R_j$.


* **Row Scaling:** A row is multiplied by a non-zero scalar: $kR_i \rightarrow R_i$.


* **Row Addition:** A multiple of one row is added to another row: $R_i + kR_j \rightarrow R_i$.


---
# The key Idea:

Find equivalent systems of equations by applying the Elemental Operations to the augmented matrix of the system, until we reach a *reduced row echelon form*.

  ---
### The method:

  1. **Step 1**: Select the first pivot, to do this, we locate a non-zero entry in the first column and swap rows (if necessary), placing the row with the pivot into the first row.


  $$\begin{bmatrix} 0 & 5 & 0 \\ 4 & 0 & 2 \\ 2 & -1 & 3 \end{bmatrix} \rightarrow_{R \rightarrow R } \begin{bmatrix} \, & \, & \, \\ \, & \, & \, \\ \, & \, & \, \end{bmatrix}$$

---

2. **Step 2**: Make the first pivot equal to 1 by dividing the row by the pivot value.

$$\begin{bmatrix} 4 & 0 & 2 \\ 0 & 5 & 0 \\ 2 & -1 & 3 \end{bmatrix} \rightarrow \begin{bmatrix} \, & \, & \, \\ \, & \, & \, \\ \, & \, & \, \end{bmatrix}$$

---


3. **Step 3**: Make zero the elements below the pivot: by subtracting multiples of the pivot row from the rows below.

$$\begin{bmatrix} 1 & 0 & 1/2 \\ 0 & 5 & 0 \\ 2 & -1 & 3 \end{bmatrix} \rightarrow \begin{bmatrix} \, & \, & \, \\ \, & \, & \, \\ \, & \, & \, \end{bmatrix}$$

---
4. **Step 4**: Once you have made all the entries below the first pivot zero, move to the next column and repeat steps 1 to 3. Repeat this process until you have a upper triangular matrix.

---
# Applying the method:


### Example 1:

Find the solution to: $$\begin{cases} 2x -y = 3 \\ 3x +4z = 18\\ x + 5y + 6z = 25 \end{cases}.$$


---
### Practice:


1. Find the solution to: $$\begin{cases} 2x -y = 3 \\ 3x +4z = 18\\ x + 5y + 6z = 25 \end{cases}.$$


2. Find the solution to: $$\begin{cases} 8x +8z = 24 \\ x +z = 3\\ y-z = 16 \end{cases}.$$

3. Find the solution to: $$\begin{cases} 8x +8z = 24 \\ x +z = 2\\ y-z = 16 \end{cases}.$$

---
# Rank of a Matrix:


> **Definition:** The rank of a matrix $A$, denoted by $rk(A)$, is the number of non-zero rows in the row echelon form of the matrix.

--


##### Examples

Compute the rank of the following matrices:

1. $\begin{bmatrix} 1 & 0 & 1 & 3 \\ -1 & 1 & 0 & 2 \\ 0 & 1 & -1 & 4 \end{bmatrix}$

2. $\begin{bmatrix} 8 & 0 & 8 & 24 \\ 1 & 0 & 1 & 3 \\ 0 & 1 & -1 & 16 \end{bmatrix}$

3. $\begin{bmatrix} 8 & 0 & 8 & 24 \\ 1 & 0 & 1 & 2 \\ 0 & 1 & -1 & 16 \end{bmatrix}$

---
# Theorem of Rouché–Capelli

> **Theorem:** Let $A$ be a matrix of size $m \times n$ and $b$ a vector of size $m$. The system of equations $Ax = b$ has:
1. **No solution** if $rk(A) \neq rk(A|b)$.
2. **A unique solution** if $rk(A) = rk(A|b) = n$.
3. **Infinitely many solutions** if $rk(A) = rk(A|b) < n$.
---
## Application of the Rouché–Capelli Theorem

> Given a system of linear equations (SLE), we can determine the number of solutions by computing the rank of the matrix and the augmented matrix.

> If the system is consistent, then we have that $$rk(A)+\#(\mbox{solution parameters})=\#(\mbox{unknowns}).$$

