---
title: "Vector Spaces"
subtitle: "Session 01" 
author: 
  - "Alejandro Ucan"
date: '`r Sys.Date()`'
output:
  xaringan::moon_reader:
    css: xaringan-themer.css
    nature:
      slideNumberFormat: "%current%"
      highlightStyle: github
      highlightLines: true
      ratio: 16:9
      countIncrementalSlides: true
---
```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
knitr::opts_chunk$set(
  fig.width = 9, fig.height = 3.5, fig.retina = 3,
  out.width = "100%",
  cache = FALSE,
  echo = FALSE,
  message = FALSE,
  warning = FALSE,
  hiline = TRUE
)
```

```{r xaringan-themer, include=FALSE, warning=FALSE}
library(xaringanthemer)
style_duo_accent(
  primary_color = "#000080",
  secondary_color = "#f2f3f4",
  inverse_header_color = "#00147e"
)
```

```{r xaringan-editable, echo=FALSE}
xaringanExtra::use_editable(expires = 1)
xaringanExtra::use_scribble()
```

# Goals of the Session
* Define the structure of a vector space. <br/><br/>
* Introduce examples of vector spaces. <br/><br/>
* Define vector subspaces. <br/><br/>

---
# Motivation

Three factory workstations share materials to produce different products. This setting is called a production system. Each workstation outputs depend on the materials it receives from the other workstations and it is given by 
$$T=\left(\begin{array}{ccc} 2& -1 & 0 \\ 0 & 1 & 1 \\ 1 & 0 & -1 \end{array}\right)$$

where the columns represent the materials and the rows represent the workstations, negative numbers indicate a consumption of materials. 

---
# Motivation 

We want to understand the following from the system:

1. How to determine when the system does not produce anything? <br/><br/>
2. How we can describe the set of all possible outputs? <br/><br/>
3. Is there an arrangement of the material distribution that guaranties the optimal production? <br/><br/>

---
# Motivation

All these questions can be answered by the theory of vector spaces. <br/><br/>

The theory of vector spaces is a branch of mathematics that studies the properties and structures of vector spaces, which are mathematical objects that generalize the concept of vectors. <br/><br/>

---

# Deducing the structure of a vector space.

Consider the following two sets: <br/>
* $\mathbb{R}$ the set of all real numbers. <br/>
* $\mathbb{R}^2$ the set of all pairs of real numbers. <br/><br/>

What are the operations that we can define on these sets? <br/>

Is there any properties that these operations satisfy? <br/>



---

# What is a vector space?

> **Definition:**
> A set of objects $V$ with two operations, addition and scalar multiplication, is called a *vector space* if it satisfies certain properties. <br/><br/> The objects in $V$ are called *vectors* and the scalars are elements of a field $\mathbb{R}$<sup> 1 </sup> .footnote[In general cases, this could be any other field of numbers like complex, quaternions or rationals] <br/><br/>

**Note:** The propierties that a vector space must satisfy are called *axioms* and are listed below. <br/><br/>

---

## Axioms of a Vector space

1. **Closure under addition:** For all $u, v \in V,$ the sum $u + v \in V.$ <br/>
2. **Closure under scalar multiplication:** For all $u \in V$ and $\lambda \in \mathbb{R},$ the product $\lambda u \in V.$ <br/>
3. **Associativity of addition:** For all $u, v, w \in V,$ $(u + v) + w = u + (v + w).$ <br/>
4. **Commutativity of addition:** For all $u, v \in V,$ $u + v = v + u.$ <br/>
5. **Existence of additive identity:** There exists an element $0 \in V$ such that for all $u \in V,$ $u + 0 = u.$ <br/>
6. **Existence of additive inverses:** For every $u \in V,$ there exists an element $-u \in V$ such that $u + (-u) = 0.$ <br/>
7. **Distributive property of scalar multiplication with respect to vector addition:** For all $\lambda \in \mathbb{R}$ and $u, v \in V,$ $\lambda (u + v) = \lambda u + \lambda v.$ <br/>
8. **Distributive property of scalar multiplication with respect to scalar addition:** For all $\lambda, \mu \in \mathbb{R}$ and $u \in V,$ $(\lambda + \mu) u = \lambda u + \mu u.$ <br/>
9. **Associativity of scalar multiplication:** For all $\lambda, \mu \in \mathbb{R}$ and $u \in V,$ $(\lambda \mu) u = \lambda (\mu u).$ <br/>
10. **Existence of multiplicative identity:** For all $u \in V,$ $1 u = u.$ <br/>

---

## Activity

<div style='position: relative; padding-bottom: 56.25%; padding-top: 35px; height: 0; overflow: hidden;'><iframe sandbox='allow-scripts allow-same-origin allow-presentation' allowfullscreen='true' allowtransparency='true' frameborder='0' height='315' src='https://www.mentimeter.com/app/presentation/alno87qexhorh1cnxt96rmgaq7zjsvg7/embed' style='position: absolute; top: 0; left: 0; width: 100%; height: 100%;' width='420'></iframe></div>

---

## Generic elements of a vector space.

> **Convention:** Given a set $V$ a variable $x$ is used to denote a generic element of $V$ and must satisfy the constraints of belonging to $V.$ <br/><br/>
> For example, if $V=\mathbb{R}^2,$ then a generic element of $V$ is denoted by $x=(a,b)$ where $a,b\in\mathbb{R}.$ <br/><br/> If $V=\mathbb{R}^{n \times m},$ then a generic element of $V$ is denoted by $x=(a_{ij})_{i=1,\dots,n}^{j=1,\dots,m}$ where $a_{ij}\in\mathbb{R}.$ <br/><br/> If $V=P_n(\mathbb{R}),$ then a generic element of $V$ is denoted by $x=a_0+a_1x+\cdots+a_nx^n$ where $a_i\in\mathbb{R}.$

---

## Examples of Vector Spaces.

> Consider the space $V=\mathbb{R}^n$ of n-tuples of real numbers, with the following operations: <br/>
* Sum: $(a_1,a_2,\dots,a_n)+(b_1,b_2,\dots,b_n)=(a_1+b_1,a_2+b_2,\dots,a_n+b_n).$ <br/>
* Multiplication by a scalar: $\lambda(a_1,a_2,\dots,a_n)=(\lambda a_1,\lambda a_2,\dots,\lambda a_n).$ <br/>
Let's prove that it is a vector space.

---

## Examples of Vector Spaces.

> Consider the space $V=P_3(\mathbb{R})$ of polynomials of degree less than or equal to three in the variable $x,$ with real coefficients, with the usual operations of sum of polynomials and scalar multiplication. <br/><br/> 

--

> Consider the space $V=C([a,b],\mathbb{R})$ of continuous functions defined on the interval $[a,b]$ with real coefficients, with the usual operations of sum of functions and scalar multiplication. <br/><br/> Let's prove that it is a vector space.

---

## Counterexamples of Vector Spaces.

> Consider the set $V=\mathbb{R}^2$ with the vector addition and the scalar multiplication defined as follows: <br/>
* Sum: $(a_1,a_2)+(b_1,b_2)=(a_1+b_1,a_2+b_2).$ <br/>
* Multiplication by a scalar: $\lambda(a_1,a_2)=(0,0).$ <br/><br/> Let's prove that this is not a vector space. <br/><br/>

---

## Vector Subspaces.

> **Definition:** A vector subspace of a vector space $V$ is a subset $W$ of $V$ that is a vector space with the operations of $V.$ <br/><br/>

> **Note:** If $W$ is a subset of a vector space $V,$ it will be sufficient to check the closure under the operations of addition and scalar multiplication to prove that $W$ is a vector subspace of $V.$ <br/><br/>

---

### Examples of Vector Subspaces.

> For example, in the vector space $\mathbb{R}^2,$ the line that passes through the origin is a vector subspace. <br/><br/> Consider the line $W=\{(x,y)\in\mathbb{R}^2:y=mx\}$ where $m$ is a constant. <br/><br/> Let's see that $W$ is a vector subspace.

---

## Vector Subspaces.

> For example, in the vector space $\mathbb{R}^3,$ the lines that pass through the origin are vector subspaces. <br/><br/> Consider the line $W_1=\{(x,y,z)\in\mathbb{R}^3:ax=z\}$ where $a$ is a constant. <br/><br/> Let's see that $W_1$ is a vector subspace. This is also true for the planes that pass through the origin. <br/><br/> Consider the plane $W_2=\{(x,y,z)\in\mathbb{R}^3:ax+by=c\}$ where $a,b,c$ are constants and $c=0.$ <br/><br/> Let's see that $W_2$ is a vector subspace.

---

## Vector Subspaces.
> For example, in the vector space $P_3(\mathbb{R}),$ the set of polynomials of degree less than or equal to two is a vector subspace. <br/><br/> Let's see that it is a vector subspace.

---

## Applied Examples of Vector spaces.

> 1. Consider the set of all possible production outputs of the production system described in the motivation. <br/><br/> The set of all possible outputs is a vector subspace of $\mathbb{R}^3.$ <br/><br/>

> 2. Consider the set of all possible distributions of materials in the production system described in the motivation. <br/><br/> The set of all possible distributions is a vector subspace of $\mathbb{R}^3.$ <br/><br/>

---

## Applied Examples of Vector subspaces

> 1. Consider the set all possible production outputs of the production system described in the motivation. Consider all those where the workstation does not produce anything. *Could be this subset a vector subspace? Which one does describe it better?* <br/><br/>

> 2. Consider the set of all possible distributions of materials in the production system described in the motivation. Consider all those outputs where three workstations produce the same amount of output. *Could be this subset a vector subspace? Which one does describe it better?* <br/><br/>

---

### Activity

[Practice on Applied Vector Spaces](https://play.kahoot.it/v2/lobby?quizId=2d659160-583f-4df2-9480-94a0abba953a)