---
title: "Optimization and Modelling"
subtitle: "Session 08"
author: 
  - "Alejandro Ucan"
date: '`r Sys.Date()`'
output:
  xaringan::moon_reader:
    css: xaringan-themer.css
    nature:
      slideNumberFormat: "%current%"
      highlightStyle: github
      highlightLines: true
      ratio: 16:9
      countIncrementalSlides: true
---

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
knitr::opts_chunk$set(
  fig.width=9, fig.height=3.5, fig.retina=3,
  out.width = "100%",
  cache = FALSE,
  echo = FALSE,
  message = FALSE, 
  warning = FALSE,
  hiline = TRUE
)
```

```{r xaringan-themer, include=FALSE, warning=FALSE}
library(xaringanthemer)
style_duo_accent(
  primary_color = "#000080",
  secondary_color = "#f2f3f4",
  inverse_header_color = "#00147e"
)
```

```{r xaringan-editable, echo=FALSE}
xaringanExtra::use_editable(expires = 1)
xaringanExtra::use_scribble()
```

```{r xaringan-logo, echo=FALSE}
xaringanExtra::use_logo(
  image_url = "https://github.com/alxcn/TecLogoEIC/blob/9562a53875418e749a296c85808a19c85fc4be74/IngenieriaCiencias_Horizontal_RGB.png",
  position = xaringanExtra::css_position(top = "2em", right = "2em")
)
```
# Goals

* Propose objective functions to solve optimization problems. <br/><br/>
* Stablish constraints to solve optimization problems. <br/><br/>
* Use Extreme values to solve optimization problems. <br/><br/>


---
# Optimization and Modelling

> Optimization is the selection of a best element from some set of available alternatives. In Differential Calculus, optimization is the process of finding the extreme values of a function based on a contextualized problem.

--
## Process of Optimization

> The process of optimization involves the following steps: <br/><br/>
  1. **Modeling**: Define the problem and the objective function. <br/><br/>
  2. **Formulation**: Define the constraints and the feasible region. <br/><br/>
  3. **Solution**: Find the extreme values of the objective function. <br/><br/>

---
## Objective Function

> The objective function is a mathematical function that represents the quantity to be optimized. It is usually denoted by $f(x)$, where $x$ is the variable to be optimized.

--
#### Example 1:

> A company produces two types of products, $A$ and $B$. The profit per unit of product $A$ is $P_A(x) = 2x$ and the profit per unit of product $B$ is $P_B(y) = 3y$. The company has a budget of $1000 to produce both products. What is the objective function?

---
#### Example 2:

> The time in a manufactury process depends on the number of products of each type produced. For type $A$, each unit increment the time the same number of minuts, and for type $B$, each unit increment the time the double of minuts. What is the objective function?

--

#### Example 3:

> The cost of producing a product depends on the number of products demanded by the public. Usually, the demand for a product is a periodic function with period of 12 months, and the lower produced products is 1000 units, and the higher produced products is 2000 units. If each product cost $10 to produce, what is the objective function?

---
## Constraints 

> Constraints are restrictions that limit the feasible region of the optimization problem. They are usually inequalities that restrict the values of the variables.

--
#### Example 1:

> A company want to produce a cardboard box with a volume of 1000 cm$^3$. The box has a square base and the height is twice the side of the base. What are the constraints?

---
#### Example 2:

> The tube of an hydraulic system needs to be replaced. We know that the superficial area of the tube is 1000 cm$^2$. The tube has a circular base and the height is twice the radius of the base. What are the constraints?

--

#### Example 3:

> A company produces two types of products, and the cost of production is limited to 1000usd. If the cost of type $A$ is 2usd and the cost of type $B$ is 3usd, what are the constraints?

---
## Extreme Values

> Once you get the objective function and the constraints, you can find the extreme values of the function. The extreme values are the maximum and minimum values of the function.}

--

#### Example 1:

> A company wants to produce a cardboard box with a volume of 1000 cm$^3$. The box has a square base and the height is twice the side of the base. We want to minize the cost of production of the box, if the cost of the material is 0.1usd/cm$^2$. What are the extreme values?

---
#### Example 2:

> The tube of an hydraulic system needs to be replaced. We know that the superficial area of the tube is 1000 cm$^2$. The tube has a circular base and the height is twice the radius of the base. We want to maximize the volume of the tube. What are the dimensions of the tube to maximize the volume?

---
#### Example 3:

> A company produces two types of products, and the cost of production is limited to 1000usd. If the cost of type $A$ is 2usd and the cost of type $B$ is 3usd, we want to maximize the profit of the company. The profit of type $A$ is 0.1 usd and the profit of type $B$ is 0.3usd. We want to obtain the maximum profit based on the conditions.