---
title: "Polynomial Functions"
subtitle: "Session 06"
author: 
  - "Alejandro Ucan"
date: '`r Sys.Date()`'
output:
  xaringan::moon_reader:
    css: xaringan-themer.css
    nature:
      slideNumberFormat: "%current%"
      highlightStyle: github
      highlightLines: true
      ratio: 16:9
      countIncrementalSlides: true
---

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
knitr::opts_chunk$set(
  fig.width=9, fig.height=3.5, fig.retina=3,
  out.width = "80%",
  cache = FALSE,
  echo = FALSE,
  message = FALSE, 
  warning = FALSE,
  hiline = TRUE
)
```

```{r xaringan-themer, include=FALSE, warning=FALSE}
library(xaringanthemer)
style_duo_accent(
  primary_color = "#000080",
  secondary_color = "#f2f3f4",
  inverse_header_color = "#00147e"
)
```

```{r xaringan-editable, echo=FALSE}
xaringanExtra::use_editable(expires = 1)
xaringanExtra::use_scribble()
```

```{r xaringan-logo, echo=FALSE}
xaringanExtra::use_logo(
  image_url = "https://github.com/alxcn/TecLogoEIC/blob/9562a53875418e749a296c85808a19c85fc4be74/IngenieriaCiencias_Horizontal_RGB.png",
  position = xaringanExtra::css_position(top = "2em", right = "2em")
)
```
# Goals

* Describe the general form a the Polynomial function model. <br/><br/>
* Describe the general behavior of a polynomial function. <br/><br/>
* Apply the polynomial function to model business problems. <br/><br/>

---
# Polynomial Functions

> __Definition:__ A polynomial function is a function of the form $$f(x)=a_nx^n+a_{n-1}x^{n-1}+\cdots+a_1x+a_0$$ where $a_n\neq 0$ and $n$ is a nonnegative integer. The domain of a polynomial function is $(-\infty,\infty)$.

<br/><br/>

* $a_n$ are known as coefficientes. <br/><br/>
* $n$ is called the degree of the polynomial. <br/><br/>
* $a_0$ is called the constant term. <br/><br/>

---
# Criterium 

> Assume that we have a large collection of data (that depends on one variable). This data could fit a polynomial function if there are multiple sign change (in the mean changes of the data). 

<br/><br/>
* The degree of the polynomial is the number of sign changes minus one. <br/><br/>
* The polynomial functions take values in $(-\infty, \infty).$ <br/><br/>
* The numbers of intercepts of the polynomial function with the $x$-axis is lower than its degree. <br/><br/>

---
#### Example

Consider the following polynomial function $f(x)=x^3-3x^2-4x+12.$ 

```{r, echo=TRUE, eval=FALSE}
x <- seq(-5, 5, by = 0.1)
y <- x^3 - 3*x^2 - 4*x + 12
plot(x, y, type = "l", col = "blue", lwd = 2)
abline(h = 0, col = "red")
```
---

#### Example

Consider the following polynomial function $f(x)=x^4 + 2 x^3 - 19 x^2 - 8 x + 60.$ 

```{r, echo=TRUE, eval=FALSE}
x <- seq(-6, 6, by = 0.1)
y <- x^4 + 2*x^3 - 19*x^2 - 8*x + 60
plot(x, y, type = "l", col = "blue", lwd = 2)
abline(h = 0, col = "red")
```
---
# What happens when we work with real life data?

> Sometimes collected data does not fit a linear regression model. 


---
## How to solve that problem?

> There is a technique called __polynomial regression__ that allows us to fit a polynomial function to a set of data.

## How to decide if my data fits a polynomial function?

1. Plot the data in a scatter plot to see some tendencies. <br/><br/>
2. Use the __criterium__ to decide if the data fits a polynomial function. <br/><br/>
3. Use the __polynomial regression__ to fit a polynomial function to the data. <br/><br/>

---


