---
title: "Generalized Eigenvectors for repeated eigenvalues"
subtitle: "Session 08"
author: 
  - "Alejandro Ucan"
date: '`r Sys.Date()`'
output:
  xaringan::moon_reader:
    css: xaringan-themer.css
    nature:
      slideNumberFormat: "%current%"
      highlightStyle: github
      highlightLines: true
      ratio: 16:9
      countIncrementalSlides: true
---

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
knitr::opts_chunk$set(
  fig.width=9, fig.height=3.5, fig.retina=3,
  out.width = "100%",
  cache = FALSE,
  echo = FALSE,
  message = FALSE, 
  warning = FALSE,
  hiline = TRUE
)
```

```{r xaringan-themer, include=FALSE, warning=FALSE}
library(xaringanthemer)
style_duo_accent(
  primary_color = "#000080",
  secondary_color = "#f2f3f4",
  inverse_header_color = "#00147e"
)
```

```{r xaringan-editable, echo=FALSE}
xaringanExtra::use_editable(expires = 1)
xaringanExtra::use_scribble()
```

```{r xaringan-logo, echo=FALSE}
xaringanExtra::use_logo(
  image_url = "https://github.com/alxcn/TecLogoEIC/blob/9562a53875418e749a296c85808a19c85fc4be74/IngenieriaCiencias_Horizontal_RGB.png",
  position = xaringanExtra::css_position(top = "2em", right = "2em")
)
```
# Goals

* Understand the concept of generalized eigenvectors. <br/><br/>
* Understand the solution in the case of complex eigenvalues. <br/><br/>
* Understand the phase plane. <br/><br/>

---
# Generalized eigenvectors


> Consider the system $$X'= \left(\begin{array}{cc} 1 & 2 \\ -2 & -3 \end{array}\right) X.$$ We want to build a general solution. The characteristic polynomial is: $$\lambda^2-2\lambda +1=0,$$ so the only eigenvalue is $\lambda=1.$ Its eigenvector is $$(A+I)=\left(\begin{array}{cc} 2 & 2 \\ -2 & -2 \end{array}\right)\Rightarrow 2v_1+2v_2=0\Rightarrow v_2=-v_1.$$ <br/>
In this case, we say that the eigenvector is not enough to build a general solution. We need to find a second vector. In order to do this, we need to find $w$ such that $$(A-\lambda I)w=v.$$
---
## Continuation example

> Solve the system $$X'= \left(\begin{array}{cc} 1 & 2 \\ -2 & -3 \end{array}\right) X.$$ <br/><br/> We have that $$(A+I)w=v\Rightarrow \left(\begin{array}{cc} 2 & 2 \\ -2 & -2 \end{array}\right) \left(\begin{array}{c} w_1  \\ w_2 \end{array}\right) =\left(\begin{array}{c} 1  \\ -1 \end{array}\right),$$ which reduces to the equation $$2w_1+2w_2=1\Leftrightarrow 2w_1=1+2w_2.$$ We can take the value $w_2=0,$ and my other vector would be $$\left(\begin{array}{c} 1/2 \\ 0 \end{array}\right)$$ Thus my general solution is $$X=c_1e^{-t}\left(\begin{array}{c} -1 \\ 1 \end{array}\right)+c_2te^{-t}\left(\begin{array}{c} 1/2 \\ 0 \end{array}\right).$$

---
### Example 2:

> Find the general solution for $$X'=\left(\begin{array}{cc} 3 & -18 \\ 2 & -9 \end{array}\right)X.$$


---
# Complex Eigenvalues

> Consider the system $$X'= \left(\begin{array}{cc} 6 & -1 \\ 5 & 4 \end{array}\right) X.$$ <br/><br/> Its characteristic polynomial is $$\lambda^2-10\lambda+29=0\Rightarrow \lambda_1=5+2i\quad \lambda_2=5-2i.$$ <br/><br/> Its eigenvectors are $$v_1=\left(\begin{array}{c} 1 \\ 1-2i \end{array}\right)\quad v_2=\left(\begin{array}{c} 1 \\ 1+2i \end{array}\right)$$
---
## Rewriting the eigenvectors 

> As we want real functions, we need to rewrite the eigenvectors in order to get real functions. <br/><br/> Let $$w_1=\left(\begin{array}{c} 1 \\ 1 \end{array}\right) \quad w_2=\left(\begin{array}{c} 0 \\ -2 \end{array}\right)$$ Thus my general solution is $$X=c_1e^{(5+2i)t}\left(\begin{array}{c} 1 \\ 1-2i \end{array}\right)+c_2e^{(5-2i)t}\left(\begin{array}{c} 1 \\ 1+2i \end{array}\right).$$
---
### Example 3:

> Find the general solution for $$X'=\left(\begin{array}{cc} -1 & 2 \\ -1 & 1 \end{array}\right)X.$$

---
# Phase Plane

<iframe src="https://www.geogebra.org/calculator/xnzc4b6w?embed" width="1200" height="450" allowfullscreen style="border: 1px solid #e4e4e4;border-radius: 4px;" frameborder="0"></iframe>

