---
title: "1st Traslation Theorem"
subtitle: "Session 11"
author: 
  - "Alejandro Ucan"
date: '`r Sys.Date()`'
output:
  xaringan::moon_reader:
    css: xaringan-themer.css
    nature:
      slideNumberFormat: "%current%"
      highlightStyle: github
      highlightLines: true
      ratio: 16:9
      countIncrementalSlides: true
---

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
knitr::opts_chunk$set(
  fig.width=9, fig.height=3.5, fig.retina=3,
  out.width = "100%",
  cache = FALSE,
  echo = FALSE,
  message = FALSE, 
  warning = FALSE,
  hiline = TRUE
)
```

```{r xaringan-themer, include=FALSE, warning=FALSE}
library(xaringanthemer)
style_duo_accent(
  primary_color = "#000080",
  secondary_color = "#f2f3f4",
  inverse_header_color = "#00147e"
)
```

```{r xaringan-editable, echo=FALSE}
xaringanExtra::use_editable(expires = 1)
xaringanExtra::use_scribble()
```

```{r xaringan-logo, echo=FALSE}
xaringanExtra::use_logo(
  image_url = "https://github.com/alxcn/TecLogoEIC/blob/9562a53875418e749a296c85808a19c85fc4be74/IngenieriaCiencias_Horizontal_RGB.png",
  position = xaringanExtra::css_position(top = "2em", right = "2em")
)
```

# Goals

* State 1st Traslation Theorem. <br/><br/>
* Provide some examples on how to apply the theorem. <br/><br/>
* Apply the theorem to solve ODEs.

---
# Motivation:
<br/><br/>

> What is the Laplace transform of the function $f(t)=e^{5t}t.$

--
<br/><br/>
Following the definition of the Laplace transform, we have that: $$\mathcal{L}\{e^{5t}t\}=\int_0^\infty e^{-st}e^{5t}t\,dt=\int_0^\infty te^{-(s-5)t}\,dt$$ $$\lim_{a\to\infty} -\frac{te^{-(s-5)t}}{s-5}-\frac{e^{-(s-5)t}}{(s-5)^2}\bigg|_0^a=\frac{1}{(s-5)^2}.$$

---
# 1st Traslation Theorem

> __Theorem:__ Let $f(t)$ be a function such that $\mathcal{L}\{f(t)\}=F(s)$ exists. Then $$\mathcal{L}\{e^{at}f(t)\}=F(s-a).$$
<br/><br/>
> __Some Implications:__ From now, every time we see a known function "shifted" by an exponential factor, this translates as a shift in $s$ in the transform.

---
## Applying the Theorem:

### Example (Direct Calculation):

> Compute the Laplace Transform of $f(t)=e^{5t}t^3,$ $g(t)=e^{-2t}\cos(4t)$ y $h(t)=e^{-2t}$ if $t>1$ and $h(t)=0$ otherwise.

--
<br/><br/>
##### Solution:

1. The shifted function of $f$ is $t^3$ and its transform is $\frac{3!}{s^4}.$ Therefore the transform of $f$ is $\frac{3!}{(s-5)^4}.$ <br/>
2. The shifted function of $g$ is $\cos(4t)$ and its transform is $\frac{s}{s^2+4^2}.$ Therefore the transform of $g$ is $\frac{s+2}{(s+2)^2+4^2}.$ <br/>
3. The shifted function of $h$ is the unit step function and its transform is $\frac{e^s}{s}.$ Therefore the transform of $h$ is $e^{s+2}\frac{1}{s+2}.$ <br/>

---
### Example (Inverse Calculation):

> Find the inverse Laplace transform of: 
$\mathcal{L}^{-1}\left\{\frac{2s+5}{(s-3)^2}\right\}$
--
<br/>
##### Solution:
Let's find out the partial fractions: $\frac{2s+5}{(s-3)^2}=\frac{A}{(s-3)}+\frac{B}{(s-3)^2}=\frac{2}{s-3}+\frac{-1}{(s-3)^2}.$

Now we notice that those are shifted functions, therefore we look for the corresponding transform and we multiply it by an exponential factor: $$\mathcal{L}^{-1}\left\{\frac{2s+5}{(s-3)^2}\right\}=\mathcal{L}^{-1}\left\{\frac{2}{s-3}\right\}+\mathcal{L}^{-1}\left\{\frac{-1}{(s-3)^2}\right\}=2e^{3t}-te^{3t}.$$

---
### Example (Inverse Calculation):

> Find the inverse Laplace transform of:
$\mathcal{L}^{-1}\left\{\frac{s+5}{s^2+4s+6}\right\}$
--
<br/>
##### Solution:
First we check if the denominator has real roots: $s^2+4s+6=0\implies s=\frac{-4\pm\sqrt{16-4(1)(6)}}{2}=-2\pm i.$ Therefore the roots are complex and we need to complete the square: $$s^2+4s+6=(s+2)^2+4.$$ Therefore the fraction is $$\frac{s+5}{(s+2)^2+4}=\frac{(s+2)}{(s+2)^2+4}+\frac{3}{(s+2)^2+4}$$ We use the formulaes to find the inverse: $$\mathcal{L}^{-1}\left\{\frac{(s+2)}{(s+2)^2+4}\right\}=e^{-2t}\cos(2t)\quad \mathcal{L}^{-1}\left\{\frac{3}{(s+2)^2+4}\right\}=\frac{3}{2}e^{-2t}\sin(2t).$$

---
### Example (Solving IVP)

> Find the solution of the IVP: $y''-6y'+9y=t^2e^{3t},\quad y(0)=2,\quad y'(0)=17.$

--
<br/><br/>
First we apply Laplace transform in both sides of the ODE. $$s^2\mathcal{L}\{y\}-sy(0)-y'(0)-6s\mathcal{L}\{y\}+6y(0)+9\mathcal{L}\{y\}=\mathcal{L}\{t^2e^{3t}\}$$ $$\implies (s^2-6s+9)\mathcal{L}\{y\}=2s+5+\frac{2}{(s-3)^3}$$ 
We simplify and check if the quadratic polynomial is factorizable:
$$\mathcal{L}\{y\}=\frac{2s+17}{(s-3)^3}+\frac{2}{(s-3)^3(s-3)^2}$$ We find the inverse Laplace transform.

---
### Example (Solving IVP)

> Find the solution of the IVP: $$y''+4y'+6y=e^{-t},\quad y(0)=0,\quad y'(0)=0.$$

--
<br/><br/>
First we apply Laplace transform in both sides fo the ODE $$s^2\mathcal{L}\{y\}-sy(0)-y'(0)+4s\mathcal{L}\{y\}-4y(0)+6\mathcal{L}\{y\}=\mathcal{L}\{e^{-t}\}$$ $$\implies (s^2+4s+6)\mathcal{L}\{y\}=\frac{1}{s+1}$$ We use partial fractions: $$\mathcal{L}\{y\}=\frac{1}{(s+1)(s^2+4s+6)}=\frac{1/3}{s-1}+\frac{s-1/3}{s^2+4s+6}$$ We construct the function using the inverse Laplace transform.

