---
title: "Laplace Transform"
subtitle: "Session 09"
author: 
  - "Alejandro Ucan"
date: '`r Sys.Date()`'
output:
  xaringan::moon_reader:
    css: xaringan-themer.css
    nature:
      slideNumberFormat: "%current%"
      highlightStyle: github
      highlightLines: true
      ratio: 16:9
      countIncrementalSlides: true
---

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
knitr::opts_chunk$set(
  fig.width=9, fig.height=3.5, fig.retina=3,
  out.width = "100%",
  cache = FALSE,
  echo = FALSE,
  message = FALSE, 
  warning = FALSE,
  hiline = TRUE
)
```

```{r xaringan-themer, include=FALSE, warning=FALSE}
library(xaringanthemer)
style_duo_accent(
  primary_color = "#000080",
  secondary_color = "#f2f3f4",
  inverse_header_color = "#00147e"
)
```

```{r xaringan-editable, echo=FALSE}
xaringanExtra::use_editable(expires = 1)
xaringanExtra::use_scribble()
```

```{r xaringan-logo, echo=FALSE}
xaringanExtra::use_logo(
  image_url = "https://github.com/alxcn/TecLogoEIC/blob/9562a53875418e749a296c85808a19c85fc4be74/IngenieriaCiencias_Horizontal_RGB.png",
  position = xaringanExtra::css_position(top = "2em", right = "2em")
)
```
# Goals:

* Motivate the Laplace transform. <br/><br/>
* Define the Laplace transform. <br/><br/>
* Compute the Laplace transform of some functions.

---
# Integral transforms

> The derivative and anti-derivative are transformations of functions, and can be thought of as linear transformations.

<br/><br/><br/><br/>

> Remember that through anti-derivatives we can define functions.


---
### Motivation:

Some problems (stated in some domain) are difficult to solve in their original presentation, but transformed can be simpler. The _integral transform_ maps an equation in its original "domain" into another "domain" in such a way that the manipulation of the equation is simpler than originally and the solution can be reconverted to the original domain.

---
# Laplace Transform

> __Definition:__ the Laplace transform of a function $f(t)$ with $t\geq 0$ is the function $F(s)$ defined by $$F(s)=\int_0^\infty f(t)e^{-st}dt=\lim_{b\to\infty}\int_0^b f(t)e^{-st}$$ where $s$ is a parameter. <br/><br/> The usual notation for the Laplace transform is $\mathcal{L}\{f(t)\}=F(s).$
---
### Example:

> Compute the Laplace transforms of $f(t)=1$ and $f(t)=t.$

---
### Example:

> Compute the Laplace transform of: $f(t)=-3t^2+2t+1,$ $f(t)=\cos(3t)$ and $f(t)=e^{2t}.$
---
### The advantages

> The Laplace transform calculation is something that you do only once, because there are some formulaes that you can use directly.
<br/><br/>
> __Theorem:__ It follows that:
  * $\mathcal{L}\{1\}=\frac{1}{s}.$
  * $\mathcal{L}\{t^n\}=\frac{n!}{s^{n+1}}.$ for an integer $n>0$
  * $\mathcal{L}\{e^{at}\}=\frac{1}{s-a}.$
  * $\mathcal{L}\{\sin(kt)\}=\frac{k}{s^2+k^2}.$
  * $\mathcal{L}\{\cos(kt)\}=\frac{s}{s^2+k^2}.$

---

## Laplace transform Linearity


> __Linearity:__ Let $f$ and $g$ be two functions defined on $[0,\infty)$ such that their Laplace transforms exist, then $$\mathcal{L}\{c_1 f+c_2 g\}=c_1\mathcal{L}\{f\}+c_2 \mathcal{L}\{g\}.$$

---
### Example

> Using Linearity in the Laplace transform, transform the following functions:
  * $f(t)=-5+3t+4t^2$
  * $f(t)=5e^{t}-2e^{-3t}$
  * $f(t)=-3t^3+\cos(4t)+ 2\sin(4t).$
  * $f(t)=\frac{\sin(3t)-\cos(4t)}{5}.$

---
# Laplace Transforms for Piecewise Functions:

> Using the integral definition, compute the Laplace transform of the following functions: <br/>
  * $f(t)=-1$ if $0\leq t \leq 1$ and $f(t)=1$ if $t>1.$
  * $f(t)=0$ if $0\leq t\leq a$ and $f(t)=t$ if $t>a.$
  * $f(t)=1$ if $a\leq t\leq b$ and $f(t)=0$ otherwise.
