---
title: "Integration Techniques and Applications"
subtitle: "Session 03"  
author: 
  - "Alejandro Ucan"
date: '`r Sys.Date()`'
output:
  xaringan::moon_reader:
    css: xaringan-themer.css
    nature:
      slideNumberFormat: "%current%"
      highlightStyle: github
      highlightLines: true
      ratio: 16:9
      countIncrementalSlides: true
---

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
knitr::opts_chunk$set(
  fig.width=9, fig.height=3.5, fig.retina=3,
  out.width = "100%",
  cache = FALSE,
  echo = FALSE,
  message = FALSE, 
  warning = FALSE,
  hiline = TRUE
)
```

```{r xaringan-themer, include=FALSE, warning=FALSE}
library(xaringanthemer)
style_duo_accent(
  primary_color = "#000080",
  secondary_color = "#f2f3f4",
  inverse_header_color = "#00147e"
)
```

```{r xaringan-editable, echo=FALSE}
xaringanExtra::use_editable(expires = 1)
xaringanExtra::use_scribble()
```

# Goals:

* Substitution. <br/><br/>
* By parts. <br/><br/>
* Center of Mass. <br/><br/>

---
# Substitution

> __Method:__ Substitution is a method of integration in which we use a change of variable to transform the integrand into a form that is easier to integrate (and use direct integration). <br/><br/>

$$\int_{a}^{b} f(g(x))g'(x)dx = \int_{g(a)}^{g(b)} f(u)du$$

--
#### Example

1. $\int_{0}^{1} \frac{2x+2}{(x^2 + 2x + 1)} dx$ <br/><br/>
2. $\int_{-1}^{1} xe^{3x^2} dx$ <br/><br/>
3. $\int_{0}^{\pi/4} \tan(x)dx$ <br/><br/>

---
# By Parts

> __Method:__ By parts is a method of integration in which we use the product rule of differentiation to transform the integrand into a form that is easier to integrate (and use direct integration). <br/><br/>

$$\int_{a}^{b} f(x)g'(x)dx = f(x)g(x) \Big|_{a}^{b} - \int_{a}^{b} f'(x)g(x)dx$$

__Remarks:__ The choice of $f$ and $g'$ in the integral depend in the following criteria: the choice of $f$ must be the function that simplify when we derive it, and the choice of $g'$ should be the one that is easier to integrate. <br/><br/>

---
#### Example

1. $\int_{0}^{1} x\ln(x)dx$ <br/><br/>
2. $\int_{0}^{1} xe^x dx$ <br/><br/>
3. $\int_{0}^{1} x\cos(x) dx$ <br/><br/>

---
# Center of Mass

> __Definition:__ Given a wire with endpoint $(a,0)$ and $(b,0)$, that has density $\rho(x)$, the center of mass of the wire is given by the following formula: <br/><br/> $$\overline{x}=\displaystyle{\frac{\int_{a}^{b} \rho(x)x dx}{\int_a^b \rho(x)dx}}$$ and the point where is the center of mass is given by $(\overline{x},0).$

---
#### Example

> Consider a wire from $(0,0)$ to $(10,0)$ with density function given by $x.$ Find the center of mass.

--
<br/><br/>
#### Solution:

1. We compute the corresponding integrals: $$\int_{0}^{10} x dx = 50$$ and $$\int_{0}^{10} x^2 dx = \frac{1000}{3}$$ <br/><br/>
2. We compute the center of mass: $$\overline{x} = \frac{\frac{1000}{3}}{50} = \frac{20}{3}$$ <br/><br/>
3. The center of mass is given by $(\frac{20}{3},0).$

---
#### Example

> Consider a wire from $(0,0)$ to $(\pi,0)$ with density function given by $\sin(x).$ Find the center of mass.

--
<br/><br/>
#### Solution:

1. We compute the corresponding integrals: $$\int_{0}^{\pi} \sin(x) dx = 2$$ and $$\int_{0}^{\pi} x\sin(x) dx = \pi$$ <br/><br/>
2. We compute the center of mass: $$\overline{x} = \frac{\pi}{2}$$ <br/><br/>
3. The center of mass is given by $(\frac{\pi}{2},0).$

