---
title: "2nd Traslation Theorem"
subtitle: "Session 12"
author: 
  - "Alejandro Ucan"
date: '`r Sys.Date()`'
output:
  xaringan::moon_reader:
    css: xaringan-themer.css
    nature:
      slideNumberFormat: "%current%"
      highlightStyle: github
      highlightLines: true
      ratio: 16:9
      countIncrementalSlides: true
---

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
knitr::opts_chunk$set(
  fig.width=9, fig.height=3.5, fig.retina=3,
  out.width = "100%",
  cache = FALSE,
  echo = FALSE,
  message = FALSE, 
  warning = FALSE,
  hiline = TRUE
)
```

```{r xaringan-themer, include=FALSE, warning=FALSE}
library(xaringanthemer)
style_duo_accent(
  primary_color = "#000080",
  secondary_color = "#f2f3f4",
  inverse_header_color = "#00147e"
)
```

```{r xaringan-editable, echo=FALSE}
xaringanExtra::use_editable(expires = 1)
xaringanExtra::use_scribble()
```

```{r xaringan-logo, echo=FALSE}
xaringanExtra::use_logo(
  image_url = "https://github.com/alxcn/TecLogoEIC/blob/9562a53875418e749a296c85808a19c85fc4be74/IngenieriaCiencias_Horizontal_RGB.png",
  position = xaringanExtra::css_position(top = "2em", right = "2em")
)
```

# Goals

* Introduce the 2nd Traslation Theorem. <br/><br/>
* Apply the 2nd Traslation Theorem in the calculation of Laplace Transforms. <br/><br/>
* Apply the 2nd Traslation Theorem in the solution of ODEs. <br/><br/>

---
# Requirements:
<br/><br/>

> The unit step function in $a$ is denoted by $u_a(t)$ and it is defined as $$u_a(t)=\left\{\begin{array}{cc} 0, & 0\leq t < a \\ 1, t\geq a \end{array} \right.$$

<br/><br/>
How does the unit step function affect our functions? <br/>
Picture the graph of the functions $f(t)=e^{t-2} u_2(t),\, g(t)=e^{t-3} u_3(t),\, h(t)=e^{t-4} u_4(t).$ <br/><br/>

---
# 2nd Traslation Theorem

> __Theorem:__ Let $f(t)$ be a function such that $\mathcal{L}\{f(t)\}=F(s)$ exists. Then $$\mathcal{L}\{f(t-a)u_a(t)\}=e^{-as}F(s).$$

<br/><br/>
> _Theorem Implications:_ Now every time we calculate the Laplace Transform of scalonated function, we can use the 2nd Traslation Theorem to obtain the result. <br/><br/>

---
## Applying the Theorem

#### Example (Direct Computation):

> Compute the Laplace Transform of $f(t)=e^{t-2}u_2(t)$, $\cos(t-\pi)u_{\pi}(t)$ y $(-3(t-3)^4+(t-3)^3)u_{3}(t).$

--
<br/><br/>
#### Solution

> $$\mathcal{L}\{e^{t-2}u_2(t)\}=\mathcal{L}\{e^{t-2}u_2(t-2)\}=e^{-2s}\mathcal{L}\{e^t\}=e^{-2s}\frac{1}{s-1}.$$ <br/><br/>

---
#### Example (Direct Computation):

> Compute the Inverse Laplace transform of $F(s)=\frac{1}{s-4}e^{-2s},$ $F(s)=\frac{s}{s^2+9}e^{-\pi s/2}.$

--
<br/><br/>
#### Solution

The Inverse laplace transform of the function without the exponential factor is $$\mathcal{L}^{-1}\left\{\frac{1}{s-4}\right\}=e^{4t}$$ since it has an exponential factor, we need to modify it with the unit step function: $$\mathcal{L}^{-1}\left\{\frac{1}{s-4}e^{-2s}\right\}=e^{4(t-2)}u_2(t-2).$$ <br/><br/>

---
### Example (Computation with  Partial Fractions)

> Compute the inverse Laplace transform of $F(s)=\frac{e^{-s}}{s(s^2-4)}.$

--
<br/>
#### Solution

1. We need to do partial fraction in the fraction without the exponential factor. $$\frac{1}{s(s-2)(s+2)}=\frac{A}{s}+\frac{B}{s-2}+\frac{C}{s+2}$$ $$=\frac{-1}{4 s} + \frac{1}{8 (s + 2)} + \frac{1}{8 (s - 2)}$$ <br/> Now, we use the 2nd traslation theorem to use express the inverse $$f(t)=\frac{-1}{4}u_1(t)+\frac{1}{8}e^{2(t+1)}u_{1}(t)+\frac{1}{8}e^{-2(t+1)}u_1(t).$$ <br/><br/> 


---
### Example (Solving ODEs):

> Find the solution for $y'+y=f(t),\quad y(0)=0,\quad f(t)=\left\{\begin{array}{cc} 0, & 0\leq t < 1 \\ 5, & t\geq 1 \end{array}\right.$

--
<br/>
#### Solution:

1. First we need to express the function $f$ as a unit step function: $f(t)=5u_1(t).$ <br/><br/>
1. We need to find the Laplace Transform of the ODE. $$\mathcal{L}\{y'\}+\mathcal{L}\{y\}=\mathcal{L}\{f(t)\}$$ $$sY(s)-y(0)+Y(s)=\frac{5}{s}$$ $$Y(s)=\frac{5e^{-s}}{s(s+1)}.$$ <br/>
1. Reduce, use partial fractions and take Inverse Laplace transform.

---
### Example (Solving ODEs):

> Find the solution for $y'+y=f(t),\quad y(0)=5,\quad f(t)=\left\{\begin{array}{cc} 0, & 0\leq t <\pi \\ 3\cos(t-\pi), & t\geq \pi \end{array}\right.$
+--
<br/>
#### Solution:

1. First we need to express the function $f$ as a unit step function: $f(t)=3\cos(t-\pi)u_\pi(t).$ <br/><br/>
1. We need to find the Laplace Transform of the ODE. $$\mathcal{L}\{y'\}+\mathcal{L}\{y\}=\mathcal{L}\{f(t)\}$$ $$sY(s)-y(0)+Y(s)=\frac{3se^{-\pi s}}{s^2+1}$$ $$Y(s)=\frac{3e^{-s}}{s(s+1)}+\frac{5}{s^2+1}.$$ <br/>
1. Reduce, use partial fractions and take Inverse Laplace transform.

