---
title: "Qualitative Description of a Function via its derivative"
subtitle: "Session 02"  
author: 
  - "Alejandro Ucan"
date: '`r Sys.Date()`'
output:
  xaringan::moon_reader:
    css: xaringan-themer.css
    nature:
      slideNumberFormat: "%current%"
      highlightStyle: github
      highlightLines: true
      ratio: 16:9
      countIncrementalSlides: true
---

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
knitr::opts_chunk$set(
  fig.width=9, fig.height=3.5, fig.retina=3,
  out.width = "100%",
  cache = FALSE,
  echo = FALSE,
  message = FALSE, 
  warning = FALSE,
  hiline = TRUE
)
```

```{r xaringan-themer, include=FALSE, warning=FALSE}
library(xaringanthemer)
style_duo_accent(
  primary_color = "#000080",
  secondary_color = "#f2f3f4",
  inverse_header_color = "#00147e"
)
```

```{r xaringan-editable, echo=FALSE}
xaringanExtra::use_editable(expires = 1)
xaringanExtra::use_scribble()
```

# Goals:

 * Describe the function via its derivative. <br/><br/>
 * Compute: extreme values, monotony intervals, and concavity intervals. <br/><br/>
 
---
# How the Derivative is related to the function?

<iframe src="https://www.desmos.com/calculator/2ew0jangf3?embed" width="1200" height="450" style="border: 1px solid #ccc" frameborder=0></iframe>

---
# Extreme Values:

> To find the extreme values of a function, we need to search for those values of $x$ where the derivative is cero. These points are called __critical__ points and satisfy $$f'(x)=0.$$

<br/><br/>

> __First derivative criterium:__ A critical point $x_0$ is said to be:
  * a _minimum_ of the function if $f'(x)<0$ for $x<x_0$ and $f'(x)>0$ if $x>x_0.$ <br/><br/>
  * a _maximum_ of the function if $f'(x)>0$ for $x<x_0$ and $f'(x)<0$ if $x>x_0.$

---
## Example:

> The function $f(x)=-x^2$ has a maximum in $x=0.$

<iframe src="https://www.desmos.com/calculator/hlq2dsx9yj?embed" width="1200" height="300" style="border: 1px solid #ccc" frameborder=0></iframe>
---
## Example:

> $f'(x)=-2x$ <br/><br/>.

Which are the critical points? $$-2x=0\Rightarrow x=0.$$

---
# Monotony.

> We say that a function $f$ is:
  * _increasing_ in an interval $(a,b)$ if $f'(x)>0$ for each point of $(a,b).$ <br/>
  * _decreasing_ in an interval $(a,b)$ if $f'(x)<0$ for each point of $(a,b).$

---
## Example:

Find the intervals where $f(x)=x^2-6x+5$ is monotonus.

> Solution:

<br/>
* Find the critical points (solve for $f'(x)=0$): $$f'(x)=2x-6=0\Rightarrow x=3.$$

* We made a test: we check the sign of the derivative at one point in $x<3$ and in other point in $x>3.$ <br/><br/>

* Increasing at $(3,+\infty)$ and decreasing at $(-\infty,3).$

---
# Excercise:
Find the intervals where $f(x)=2x^3-3x^2-12x+7$ is decresing.

---
# Excercise:
Find the intervals where $f(x)=2x^3-3x^2-12x+7$ is decresing.

> Solution:

<br/>

* We compute the derivative $f'(x)=6x^2-6x-12.$ <br/>

* Solve for $f'(x)<0.$ $$6x^2-6x-12<0 \Rightarrow (x-2)(x+1)<0.$$ <br/><br/><br/>


* Is decresing at $(-1,2).$

---
# Concavity:

> We are interested to know where a function looks like a mointain, like a valler and where these types changes.

---
# Concavity and derivaties.

> For a function $f(x)$: <br/><br/><br/>
* $f(x)$ es __concave up__ at $(a,b),$ if $f''(x)>0$ at $(a,b).$ <br/><br/>
* $f(x)$ es __concave down__ at $(a,b),$ if $f''(x)<0$ at $(a,b).$ <br/><br/>
* If $f''(x)$ changes it sign at $x=c,$ then $f$ has an __inflection point__ at $x=c.$ 

---
# Example:
Find the concavity intervals and inflection points of $f(x)=2x^3-3x^2-12x+7.$

> Solution:

<br/>

* Solve $f''(x)=0.$ $$ f''(x)=12x-6=0\Rightarrow x=\frac{1}{2}.$$

* We made a test for $f''$ for $x<1/2$ y $x>1/2.$ <br/><br/><br/>


* It is concave up at $(1/2,\infty),$ concave down at $(-\infty, 1/2)$ and has an inflection point at $x=1/2.$

---
# Practical excercise:

> Determine the monotony intervals, the concavity intervals and inflections points of the following points. Made a sketch of the graph of the function using this information. <br/><br/>
1. $f(x)=x^2-12x+1.$ <br/><br/>
2. $f(x)=5-3x^2+x^3.$ <br/><br/>
3. $f(x)=-3x^2+2x-3.$