---
title: "Newton's Cooling Law"
subtitle: "Session 02"  
author: 
  - "Alejandro Ucan"
date: '`r Sys.Date()`'
output:
  xaringan::moon_reader:
    css: xaringan-themer.css
    nature:
      slideNumberFormat: "%current%"
      highlightStyle: github
      highlightLines: true
      ratio: 16:9
      countIncrementalSlides: true
---

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
knitr::opts_chunk$set(
  fig.width=9, fig.height=3.5, fig.retina=3,
  out.width = "100%",
  cache = FALSE,
  echo = FALSE,
  message = FALSE, 
  warning = FALSE,
  hiline = TRUE
)
```

```{r xaringan-themer, include=FALSE, warning=FALSE}
library(xaringanthemer)
style_duo_accent(
  primary_color = "#000080",
  secondary_color = "#f2f3f4",
  inverse_header_color = "#00147e"
)
```

```{r xaringan-editable, echo=FALSE}
xaringanExtra::use_editable(expires = 1)
xaringanExtra::use_scribble()
```

# Goals:

 * State Newton's Cooling Law. <br/><br/>
 * Find the Function that models this law. <br/><br/>
 * Understand the qualitative behavior of the function. <br/><br/>
 * Solve some modelling problems related to Newton's cooling law.
---
# Recall Riemann sums


---
# Definite Integral

> __Definition:__ Let $f$ be a function defined on the interval $[a,b]$. The definite integral of $f$ from $a$ to $b$ is defined as the limit of a Riemann sum when we increase the number of subintervals in a partition. In symbols $$\int_a^b f(x) dx = \lim_{n \to \infty} \sum_{i=1}^n f(x_i^*) \Delta x_i$$ provided that this limit exists and gives the same value for all possible choices of sample points $x_i^*$ in the subintervals $[x_{i-1}, x_i]$.

---
## Some conventions:

* We will only use regular partitions but the definition is valid for any kind of partition. <br/><br/>
* We will only consider left Riemann sums, but this computation is also valid for right and middle sums. <br/><br/>
* We will use the notation $\Delta x = \frac{b-a}{n}$ and $x_i = a + i \Delta x$.

---
# Main Hypothesis

> **Newton's Cooling Law:** <br/><br/> The *rate of change* of the Temperature of an object, at any time, is *proportional* to the *difference* of object's temperature and medium's temperature.

--

#### How can we represent that statement into a mathematical object?

---
## The Differential Equation

\[\frac{d T(t)}{dt}= k(T(t)-T_m)\]

--

* $T(t)$ is the function that measure object's temperature at any time. (Type: Unknown) <br/><br/>

* $k$ is the cooling ratio. (Type: Parameter usually known) <br/><br/>

* $T_m$ is the medium temperature. (Type: Parameter usually known, could be a function)

---
## How to determine a function that models this phenomena?

