---
title: "Newton's Cooling Law"
subtitle: "Session 02"  
author: 
  - "Alejandro Ucan"
date: '`r Sys.Date()`'
output:
  xaringan::moon_reader:
    css: xaringan-themer.css
    nature:
      slideNumberFormat: "%current%"
      highlightStyle: github
      highlightLines: true
      ratio: 16:9
      countIncrementalSlides: true
---

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
knitr::opts_chunk$set(
  fig.width=9, fig.height=3.5, fig.retina=3,
  out.width = "100%",
  cache = FALSE,
  echo = FALSE,
  message = FALSE, 
  warning = FALSE,
  hiline = TRUE
)
```

```{r xaringan-themer, include=FALSE, warning=FALSE}
library(xaringanthemer)
style_duo_accent(
  primary_color = "#000080",
  secondary_color = "#f2f3f4",
  inverse_header_color = "#00147e"
)
```

```{r xaringan-editable, echo=FALSE}
xaringanExtra::use_editable(expires = 1)
xaringanExtra::use_scribble()
```

# Goals:

 * State Newton's Cooling Law. <br/><br/>
 * Find the Function that models this law. <br/><br/>
 * Understand the qualitative behavior of the function. <br/><br/>
 * Solve some modelling problems related to Newton's cooling law.
---
# Newton's Cooling Law

> **Hypothesis:** The rate of change of the temperature of an object is proportional to the difference between the temperature of the object and the temperature of the surrounding medium.

--

> **Mathematically:** If $T(t)$ is the temperature of the object at time $t$, and $T_s$ is the temperature of the surrounding medium, then the rate of change of the temperature of the object is given by $$\frac{dT}{dt} = k(T - T_s)$$ where $k$ is the cooling constant.

---
# Parts of the Mathematical Model:

In the equation $$\frac{dT}{dt} = k(T - T_s) \quad T(x_0)=T_0$$

 * $T(t)$ is the temperature of the object at time $t$. <br/><br/>
 * $T_s$ is the temperature of the surrounding medium. <br/><br/>
 * $k$ is the cooling constant. <br/><br/>
 * $\frac{dT}{dt}$ is the rate of change of the temperature of the object at time $t$. <br/><br/>
 * $T(x_0)=T_0$ is the initial condition.

---
# Finding the Unknown?

This is a first order linear differential equation. We can solve it by separation of variables.


--
The general solution is given by $$T(t) = T_s + Ce^{kt}$$ where the constant $C$ depends on the initial condition. 

---
# What can we say about the function?

 * The temperature of the object will approach the temperature of the surrounding medium as time goes to infinity. <br/><br/>
 * The temperature of the object increase or decrease depending on the difference between the temperature of the object and the temperature of the surrounding medium. <br/><br/>
 * The temperature of the object will decrease if the temperature of the object is greater than the temperature of the surrounding medium. <br/><br/>
  * The temperature of the object will increase if the temperature of the object is less than the temperature of the surrounding medium. <br/><br/>
  * The cooling constant $k$ determines how fast the temperature of the object approaches the temperature of the surrounding medium.
---
# Example 1:

A cup of coffee has a temperature of $90^\circ C$ in a room with a temperature of $20^\circ C$. After 5 minutes, the temperature of the coffee is $70^\circ C$. Find the temperature of the coffee after 10 minutes.

--
**Solution:**

 * $T_s = 20^\circ C$, $T_0 = 90^\circ C$, $T(5) = 70^\circ C$.
 * We need to find the cooling constant $k$.
 * We can use the general solution to find the temperature of the coffee after 10 minutes.

