---
title: "Orthogonality"
subtitle: "Session 07"  
author: 
  - "Alejandro Ucan"
date: '`r Sys.Date()`'
output:
  xaringan::moon_reader:
    css: xaringan-themer.css
    nature:
      slideNumberFormat: "%current%"
      highlightStyle: github
      highlightLines: true
      ratio: 16:9
      countIncrementalSlides: true
---

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
knitr::opts_chunk$set(
  fig.width=9, fig.height=3.5, fig.retina=3,
  out.width = "80%",
  cache = FALSE,
  echo = FALSE,
  message = FALSE, 
  warning = FALSE,
  hiline = TRUE
)
```

```{r xaringan-themer, include=FALSE, warning=FALSE}
library(xaringanthemer)
style_duo_accent(
  primary_color = "#000080",
  secondary_color = "#f2f3f4",
  inverse_header_color = "#00147e"
)
```

```{r xaringan-editable, echo=FALSE}
xaringanExtra::use_editable(expires = 1)
xaringanExtra::use_scribble()
```


# Goals of the Session

* Introduce the concept of inner product. <br/><br/>
* Define the angle between two vectors. <br/><br/>
* Introduce the concept of orthogonality. <br/><br/>
* Define the projection of a vector onto another vector. <br/><br/>

---
# Motivation

A factory produces three product variants (P1, P2, P3) with quality metrics measured along three dimensions (Precision, Durability, Finish). The quality control department proposes as quality directions the vectors: $b_1 = (1,1,1)^t$ (balanced quality), $b_2 = (1,-1,0)^t$ (precision-durability tradeoff), and $b_3 = (1,1,-2)^t$ (finish emphasis). The quality control team wants to ensure that the production process is efficient and balanced across these dimensions, meaning that this proposal is an orthonormal basis. In a given production cycle, the quality metrics of a batch is given by $q=(2.3, 1.7, 0.8)^t$ and the team wants to determine how much of each quality direction is present in the batch.

---
# Inner Product

> __Definition:__ Let $V$ be a vector space over $\mathbb{R}$. An inner product on $V$ is a function $\langle \cdot, \cdot \rangle: V \times V \rightarrow \mathbb{R}$ that satisfies the following properties for all $u, v, w \in V$ and $c \in \mathbb{R}$: <br/><br/>
  1. Commutative: $\langle u, v \rangle = \langle v, u \rangle$ <br/>
  1. Linear in the first argument: $\langle cu + w, v \rangle = c \langle u, v \rangle + \langle w, v \rangle$ <br/>
  1. Positive definite: $\langle u, u \rangle \geq 0$ and $\langle u, u \rangle = 0$ if and only if $u = 0$ <br/><br/>
---

## Examples of Inner Product

Let $V=\mathbb{R}^n$ and $u, v \in \mathbb{R}^n$, then the standard inner product on $\mathbb{R}^n$ is given by: <br/><br/> $$\langle u, v\rangle = u_1v_1 + u_2v_2 + \cdots + u_nv_n = \sum_{i=1}^n u_iv_i$$
--

Let $V=\mathbb{R}^2$ and $u, v \in \mathbb{R}^2$, then the standard inner product on $\mathbb{R}^2$ is given by: <br/><br/> $$\langle u, v\rangle = u_1v_1 + 2u_2v_2$$ 

---
## Counterexample

Let $V=\mathbb{R}^2$ and $u, v \in \mathbb{R}^2$, then the function $\langle u, v\rangle = u_1v_1 - u_2v_2$ is not an inner product on $\mathbb{R}^2$ because it is not positive definite. <br/><br/> For example, $\langle (1,2), (1,2)\rangle < 0$.

---
# Projections and Orthogonality

> **Definition:** Let $V$ be a vector space with an inner product, and $u, v \in V$ two vectors. We define the projection of $u$ onto $v,$ denoted by $\text{proj}_v(u),$ as the vector $$\text{proj}_v(u)=\frac{\langle u,v\rangle}{\langle v,v\rangle} v.$$
--
<br/><br/>
##### Example
Let $u$ and $v$ be the vectors $(2,3)$ and $(4,1)$ in $\mathbb{R}^2$ with the standard inner product. Then, $$\text{proj}_{(4,1)}(2,3)=\frac{\langle (2,3),(4,1)\rangle}{\langle (4,1),(4,1)\rangle} (4,1)=\frac{11}{17}(4,1).$$

---
## Angle between Two Vectors

> **Definition:** Let $V$ be a vector space with an inner product, we define the angle between two vectors $u,v\in V$ as the angle $\theta$ such that $$\cos(\theta)=\frac{\langle u,v\rangle}{\sqrt{\langle u,u\rangle \langle v,v\rangle}}.$$

--
##### Example

Find the angle between $(2,3)$ and $(4,1)$ in $\mathbb{R}^2$ with the standard inner product.


---
## Orthogonality

> **Definition:** We say that two vectors $u,v\in V$ are orthogonal if $\langle u,v\rangle=0.$

<br/><br/>

Why this defintion implies orthogonality? <br/><br/>
--
<br/>
##### Remark:

If $\langle u,v\rangle=0,$ then $\cos(\theta)=0,$ therefore, $\theta=\pm k \frac{\pi}{2}$ with $k$ an odd integer.
---
##### Example: 

> In $\mathbb{R}^2,$ these vectors <br/><br/>
 * $(0,1),(1,0),$ <br/><br/>
 * $(1,1),(-1,1),$ <br/><br/>
 * $(3,1), (1,-3)$ <br/><br/>
 are orthogonals to each other.


---
## Applied Orthogonality and Projections

A factory produces three product variants $(P1, P2, P3)$ with quality metrics measured along three dimensions (Precision, Durability, Finish). The quality control department proposes as quality directions the vectors: $b_1 = (1, 1, 1)^t$ (balanced quality), $b_2 = (1, -1, 0)^t$ (precision-durability tradeoff), and $b_3 = (1, 1, -2)^t$ (finish emphasis). The quality control team wants to ensure that the production process is efficient and balanced across these dimensions, meaning that this proposal is an orthonormal basis. In a given production cycle, the quality metrics of a batch is given by $q=(2.3, 1.7, 0.8)^t$ and the team wants to determine how much of each quality direction is present in the batch.

---
## Applied Orthogonality and Projections

A robotic arm uses three inertial sensors $(S1, S2, S3)$ measuring vibrations in different directions. Sensor $S1$ measures vibrations along the x-axis, $S2$ along the y-axis, and $S3$ along the z-axis. The team wants to change the places of the sensors in new based on the industrial needs. The new sensors are placed in the directions $b_1 = (1, 1, 0)^t$ (x-axis), $b_2 = (-1, 1, 0)^t$ (y-axis), and $b_3 = (0, 0, 1)^t$ (z-axis). Where $b_2$ represent a dangerous direction for the robot implying twisting. Signal readings from the sensors are given by $s=(2.1, 0.3, 1.5)$ and the teams wants to determine how much of the vibrations are present in the twisting direction.