---
title: "Minima and Maxima for One Variable Functions"
subtitle: "Session 07"
author: 
  - "Alejandro Ucan"
date: '`r Sys.Date()`'
output:
  xaringan::moon_reader:
    css: xaringan-themer.css
    nature:
      slideNumberFormat: "%current%"
      highlightStyle: github
      highlightLines: true
      ratio: 16:9
      countIncrementalSlides: true
---

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
knitr::opts_chunk$set(
  fig.width=9, fig.height=3.5, fig.retina=3,
  out.width = "100%",
  cache = FALSE,
  echo = FALSE,
  message = FALSE, 
  warning = FALSE,
  hiline = TRUE
)
```

```{r xaringan-themer, include=FALSE, warning=FALSE}
library(xaringanthemer)
style_duo_accent(
  primary_color = "#000080",
  secondary_color = "#f2f3f4",
  inverse_header_color = "#00147e"
)
```

```{r xaringan-editable, echo=FALSE}
xaringanExtra::use_editable(expires = 1)
xaringanExtra::use_scribble()
```

```{r xaringan-logo, echo=FALSE}
xaringanExtra::use_logo(
  image_url = "https://github.com/alxcn/TecLogoEIC/blob/9562a53875418e749a296c85808a19c85fc4be74/IngenieriaCiencias_Horizontal_RGB.png",
  position = xaringanExtra::css_position(top = "2em", right = "2em")
)
```
# Goals

* Introduce the concept of local minima and maxima. <br/><br/>
* Introduce the concept of global minima and maxima. <br/><br/>
* Introduce the concept of critical points. <br/><br/>
* Introduce the First Derivative Test. <br/><br/>

---
# Motivation

Imagine that my function $f$ models: 
1. The cost of producing $x$ units of a product.
2. The distance from a given point. 
3. The temperature of a room.

It is possible to know the minimum cost of producing $x$ units of a product, the maximum distance from a given point, and the minimum temperature of a room, whitout the need to graph the function.
---
# Local Minima and Maxima

> **Definition:** Let $f$ be a function defined on an interval $I$ and let $c$ be a point in $I$. We say that $f$ has a local minimum at $c$ if there exists an interval $(a,b)$ containing $c$ such that $f(c) \leq f(x)$ for all $x$ in $(a,b)$. We say that $f$ has a local maximum at $c$ if there exists an interval $(a,b)$ containing $c$ such that $f(c) \geq f(x)$ for all $x$ in $(a,b)$.

---
## Derivatives and Local Minima and Maxima

> **Theorem:** Let $f$ be a function defined on an interval $I$ and let $c$ be a point in $I$. If $f$ has a local minimum or maximum at $c$ and $f$ is differentiable at $c$, then $f'(c) = 0$.

--
#### 1st Derivative Test

0. If $f'(c)=0,$ then $x=c$ is called a critical point.
1. If $f'(c) = 0$ and $f'(x)<0$ for $x<c$ and $f'(x)>0$ for $x>c$, then $f$ has a local minimum at $c$.
2. If $f'(c) = 0$ and $f'(x)>0$ for $x<c$ and $f'(x)<0$ for $x>c$, then $f$ has a local maximum at $c$.

---
#### Example 1

Describe the Local Extrema of the function $f(x) = x^3 - 3x^2 + 2.$ <br/><br/>

--
#### Solution

1. what are the possible values where the extrema are? $f'(x) = 3x^2 - 6x = 3x(x-2).$ Then, $f'(x) = 0$ for $x=0$ and $x=2.$ <br/><br/>

2. What is the sign of $f'(x)$ in the intervals $(-\infty,0),$ $(0,2),$ and $(2,\infty)$? <br/> Chosing some test points, we have that $f'(-1) = 3(-1)(-1-2) = 9 > 0,$ $f'(1) = 3(1)(1-2) = -3 < 0,$ and $f'(3) = 3(3)(3-2) = 9 > 0.$ <br/><br/>}

3. Therefore, $f$ has a local minimum at $x=2$ and a local maximum at $x=0.$

---
#### Example 2.

Describe the local extrema points for the function $f(x)=x^2e^{2x}.$ <br/><br/>

--
#### Solution

1. What are the possible values where the extrema are? $f'(x) = 2xe^{2x} + x^2(2e^{2x}) = 2xe^{2x}(1+x).$ Then, $f'(x) = 0$ for $x=0$ and $x=-1.$ <br/><br/>

2. What is the sign of $f'(x)$ in the intervals $(-\infty,-1),$ $(-1,0),$ and $(0,\infty)$? <br/> Chosing some test points, we have that $f'(-2) = 2(-2)e^{-4}(1-2) = 4e^{-4} > 0,$ $f'(-1/2) = 2(-1/2)e^{-1}(1-1/2) = -e^{-1} < 0,$ and $f'(1) = 2(1)e^{2}(1+1) = 4e^{2} > 0.$ <br/><br/>

3- Therefore, $f$ has a local minimum at $x=0$ and a local maximum at $x=-1.$

---
#### Example 3

Describe the local extrema point for the function $f(x)=\sqrt{r^2-x^2}.$ <br/><br/>

--
#### Solution

1. What are the possible values where the extrema are? $f'(x) = -\frac{x}{\sqrt{r^2-x^2}}.$ Then, $f'(x) = 0$ for $x=0.$ <br/><br/>

2. What is the sign of $f'(x)$ in the intervals $(-r,0)$ and $(0,r)$? <br/> Chosing some test points, we have that $f'(-r-1) = \frac{r+1}{\sqrt{r^2-1}} > 0,$ and $f'(r+1) = -\frac{r+1}{\sqrt{r^2-1}} < 0.$ <br/><br/>

3. Therefore, $f$ has a local maximum at $x=0.$


