---
title: "Solids of Revolution"
subtitle: "Session 14"
author: 
  - "Alejandro Ucan"
date: '`r Sys.Date()`'
output:
  xaringan::moon_reader:
    css: xaringan-themer.css
    nature:
      slideNumberFormat: "%current%"
      highlightStyle: github
      highlightLines: true
      ratio: 16:9
      countIncrementalSlides: true
---

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
knitr::opts_chunk$set(
  fig.width=9, fig.height=3.5, fig.retina=3,
  out.width = "100%",
  cache = FALSE,
  echo = FALSE,
  message = FALSE, 
  warning = FALSE,
  hiline = TRUE
)
```

```{r xaringan-themer, include=FALSE, warning=FALSE}
library(xaringanthemer)
style_duo_accent(
  primary_color = "#000080",
  secondary_color = "#f2f3f4",
  inverse_header_color = "#00147e"
)
```

```{r xaringan-editable, echo=FALSE}
xaringanExtra::use_editable(expires = 1)
xaringanExtra::use_scribble()
```

```{r xaringan-logo, echo=FALSE}
xaringanExtra::use_logo(
  image_url = "https://github.com/alxcn/TecLogoEIC/blob/9562a53875418e749a296c85808a19c85fc4be74/IngenieriaCiencias_Horizontal_RGB.png",
  position = xaringanExtra::css_position(top = "2em", right = "2em")
)
```

# Goals

* Understand the concept of solid of revolution: disk method and washer method.
* Calculate the volume of a solid of revolution using the disk method.
* Calculate the volume of a solid of revolution using the washer method.

---
# Solids of Revolution


> **Definition:** A *solid of revolution* is a solid (in 3D) obtained by rotating a 2D region around an axis (usually the x-axis).

--
<br/><br/>

> **Example:** The solid obtained by rotating the region between the curve $y = x^2$ and the x-axis from $x = 0$ to $x = 1$ around the x-axis.


---
# Idea of the Method

We will use the same idea of rectangles, we can approximate the value of the volume of the solid using *disks* or *washers* depending on the type of figure we are rotating. <br/><br/>

Recall that a disk is a cyllinder with circular base, whose volume is given by $V=\pi r^2 h$,$ r$ is the radius of the base and $h$ is the height of the cylinder. <br/><br/>

The radious of the disk is the distance from the axis of rotation of the curve and the curve, but the curve is given by $f(x),$ so the radius is $r=f(x):$ $$V_{\Delta x}=\pi f(x)^2 \Delta x$$

---
## Disk Method:

The volume of the solid of revolution obtained by rotating the region between the curve $y=f(x)$ and the x-axis from $x=a$ to $x=b$ around the x-axis is given by: $$V=\lim_{n\to\infty}\sum_{i=1}^n \pi f(x_i)^2 \Delta x=\pi\int_a^b (f(x))^2 dx$$

Where $x_i$ is a point in the interval $[a,b]$ and $\Delta x=\frac{b-a}{n}$.

---
## Washer Method:

The volume of the solid of revolution obtained by rotating the region between the curves $y=f(x)$ and $y=g(x)$ from $x=a$ to $x=b$ around the x-axis is given by: $$V=\lim_{n\to\infty}\sum_{i=1}^n \pi (f(x_i)^2-g(x_i)^2) \Delta x=\pi\int_a^b (f(x))^2-(g(x))^2 dx$$

Where $x_i$ is a point in the interval $[a,b]$ and $\Delta x=\frac{b-a}{n}$.



