---
title: "U-Substitution and By Parts"
subtitle: "Session 10"
author: 
  - "Alejandro Ucan"
date: '`r Sys.Date()`'
output:
  xaringan::moon_reader:
    css: xaringan-themer.css
    nature:
      slideNumberFormat: "%current%"
      highlightStyle: github
      highlightLines: true
      ratio: 16:9
      countIncrementalSlides: true
---

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
knitr::opts_chunk$set(
  fig.width=9, fig.height=3.5, fig.retina=3,
  out.width = "100%",
  cache = FALSE,
  echo = FALSE,
  message = FALSE, 
  warning = FALSE,
  hiline = TRUE
)
```

```{r xaringan-themer, include=FALSE, warning=FALSE}
library(xaringanthemer)
style_duo_accent(
  primary_color = "#000080",
  secondary_color = "#f2f3f4",
  inverse_header_color = "#00147e"
)
```

```{r xaringan-editable, echo=FALSE}
xaringanExtra::use_editable(expires = 1)
xaringanExtra::use_scribble()
```

```{r xaringan-logo, echo=FALSE}
xaringanExtra::use_logo(
  image_url = "https://github.com/alxcn/TecLogoEIC/blob/9562a53875418e749a296c85808a19c85fc4be74/IngenieriaCiencias_Horizontal_RGB.png",
  position = xaringanExtra::css_position(top = "2em", right = "2em")
)
```
# Goals


* Apply the Method of Substitution to calculate integrals. <br/><br/>
* Apply the Method of Integration by Parts to calculate integrals. <br/><br/>

---
---
# Substitution

> __Method:__ Substitution is a method of integration in which we use a change of variable to transform the integrand into a form that is easier to integrate (and use direct integration). <br/><br/>

$$\int f(g(x))g'(x)dx = \int f(u)du= F(u)+C$$

__Remarks:__ The choice of $u$ in the integral depend in the following criteria: the choice of $u$ must be an internal function on a composite function. <br/><br/>

---
#### Example

1. $\int \frac{2x+2}{(x^2 + 2x + 1)} dx$ <br/><br/>
2. $\int xe^{3x^2} dx$ <br/><br/>
3. $\int \tan(x)dx$ <br/><br/>
4. $\int \frac{1}{x\ln(x)}dx$ <br/><br/>
5. $\int \frac{3x}{\sqrt{x^2+2}e^{\sqrt{x^2+2}}dx$ <br/><br/>

---
# By Parts

> __Method:__ By parts is a method of integration in which we use the product rule of differentiation to transform the integrand into a form that is easier to integrate (and use direct integration). <br/><br/>

$$\int f(x)g'(x)dx = f(x)g(x) - \int f'(x)g(x)dx$$
$$\int udv= uv - \int vdu$$

__Remarks:__ The choice of $f$ and $g'$ in the integral depend in the following criteria: the choice of $f$ must be the function that simplify when we derive it, and the choice of $g'$ should be the one that is easier to integrate. <br/><br/>

---
## ILATE

The ILATE rule is a mnemonic to help you remember which function to differentiate and which to integrate when using integration by parts. <br/><br/> If the term is a product is upper in the list, then it should be the function to differentiate. <br/><br/>

1. Inverse functions. <br/><br/>
2. Logarithmic functions. <br/><br/>
3. Algebraic functions. <br/><br/>
4. Trigonometric functions. <br/><br/>
5. Exponential functions. <br/><br/>

---
#### Example

1. $\int x\ln(3x)dx$ <br/><br/>
2. $\int 4x e^{x+3} dx$ <br/><br/>

---
#### Multiple Times By Parts

1. $\int x^2e^{4x-3}dx$ <br/><br/>
1. $\int x^2\sin(3x+2)dx$ <br/><br/>

#### Cyclic Integration by Parts

1. $\int \cos(x)e^x dx$ <br/><br/>
1. $\int \sin(x)e^{4x+1} dx$ <br/><br/>


