---
title: "Physics and Differential Equations"
subtitle: "Session 04"  
author: 
  - "Alejandro Ucan"
date: '`r Sys.Date()`'
output:
  xaringan::moon_reader:
    css: xaringan-themer.css
    nature:
      slideNumberFormat: "%current%"
      highlightStyle: github
      highlightLines: true
      ratio: 16:9
      countIncrementalSlides: true
---

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
knitr::opts_chunk$set(
  fig.width=9, fig.height=3.5, fig.retina=3,
  out.width = "100%",
  cache = FALSE,
  echo = FALSE,
  message = FALSE, 
  warning = FALSE,
  hiline = TRUE
)
```

```{r xaringan-themer, include=FALSE, warning=FALSE}
library(xaringanthemer)
style_duo_accent(
  primary_color = "#000080",
  secondary_color = "#f2f3f4",
  inverse_header_color = "#00147e"
)
```

```{r xaringan-editable, echo=FALSE}
xaringanExtra::use_editable(expires = 1)
xaringanExtra::use_scribble()
```

# Goals:

 * Broaden the type of Differential Equations. <br/><br/>
 * State physical systems via differential equations. <br/><br/>
 * Look for solutions.
 
---
# A key example.

> Imagine that we have a contaminated slice of bread with some bacteria. After some observations, we can assure that the growing rate of the population is 3 times each second. <br/><br/>
If $P(t)$ denotes the measure of the bacteria population (by thousands), then the __growing rate__ corresponds to $$\frac{dP}{dt}.$$ <br/><br/>
Following our observations, we can state that $$\frac{dP}{dt}=3P.$$

---
## Information of our problem. 

> Recall that (from Differential Calculus) there is one function whose derivative its almost the same as the function (up to a constant). If $f(t)=e^{3t},$ then $$\frac{df}{dt}=3e^{3t}=3f(t).$$ <br/><br/> But this also works for the function $f_1(t)=2e^{3t},$ $f_1(t)=10e^{3t},$ etc. 

---
## A picture of this. 

<iframe src="https://www.geogebra.org/calculator/eqekpbc2?embed" width="1200" height="450" allowfullscreen style="border: 1px solid #e4e4e4;border-radius: 4px;" frameborder="0"></iframe>

---
# A physical model.

> Assume that we have an object of mass $m$ attached at the end of a spring, and the other end of the spring is fixed. <br/><br/> Hooke's law says that:<br/> If a spring is stretched (compressed) $x$ units from its natural length, then it exerts a force that is proportional to $x:$ $$\mbox{restoring force}=-kx.$$
By Newton's second law (ignoring all other external forces), says that _force is equal mass times acceleration_ $$m\frac{d^2x}{dt^2}=-kx.$$

---
## The mass-spring model

> Previous differential equations is a model of the physical phenomena. This differential equation doesn't look like solvable. To find its solution we need more refinated techniques for differential equations theory. <br/><br/>
We can picture what should happend. 

---
## A picture of the phenomenon

<iframe src="https://www.geogebra.org/calculator/j7vydp2d?embed" width="1200" height="450" allowfullscreen style="border: 1px solid #e4e4e4;border-radius: 4px;" frameborder="0"></iframe>

---
# A less-ideal model

> Let's assume that in our last model we consider the presence of a new external force, for example the existence of resistance. <br/><br/> This could be think as that our spring-mass is inmersed in a viscous medium. <br/><br/> In mechanics, the force induced by the viscous medium acts proportionally to the instant velocity. Our model transforms to $$m\frac{d^2x}{dt^2}=-kx-\beta \frac{dx}{dt}.$$

---
## Lets see some pictures.


