---
title: "Partial Fractions (1st and 2nd Case)"
subtitle: "Session 11"
author: 
  - "Alejandro Ucan"
date: '`r Sys.Date()`'
output:
  xaringan::moon_reader:
    css: xaringan-themer.css
    nature:
      slideNumberFormat: "%current%"
      highlightStyle: github
      highlightLines: true
      ratio: 16:9
      countIncrementalSlides: true
---

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
knitr::opts_chunk$set(
  fig.width=9, fig.height=3.5, fig.retina=3,
  out.width = "100%",
  cache = FALSE,
  echo = FALSE,
  message = FALSE, 
  warning = FALSE,
  hiline = TRUE
)
```

```{r xaringan-themer, include=FALSE, warning=FALSE}
library(xaringanthemer)
style_duo_accent(
  primary_color = "#000080",
  secondary_color = "#f2f3f4",
  inverse_header_color = "#00147e"
)
```

```{r xaringan-editable, echo=FALSE}
xaringanExtra::use_editable(expires = 1)
xaringanExtra::use_scribble()
```

```{r xaringan-logo, echo=FALSE}
xaringanExtra::use_logo(
  image_url = "https://github.com/alxcn/TecLogoEIC/blob/9562a53875418e749a296c85808a19c85fc4be74/IngenieriaCiencias_Horizontal_RGB.png",
  position = xaringanExtra::css_position(top = "2em", right = "2em")
)
```

# Goals

* Decompose a rational function into partial fractions. <br/><br/>
* Apply the method of undetermined coefficients to calculate the coefficients of the partial fractions. <br/><br/>

---
# Integration by Partial Fractions 

### Main assumptions

> Given the a reduced rational function $$R(x)=\frac{p(x)}{q(x)}$$ where $p(x)$ and $q(x)$ are polynomials and the degree of $p(x)$ is less than the degree of $q(x)$, then $R(x)$ can be expressed as a sum of "smaller" fractions called partial fractions. The *number* of partial fractions depends on the number of factors of the denominator $q(x)$.

---
## Case 1: Distinct Linear Factors

> If the denominator $q(x)$ has distinct linear factors, then the partial fractions are of the form: $$\frac{A_1}{x-a_1} + \frac{A_2}{x-a_2} + \cdots + \frac{A_n}{x-a_n}$$ for each linear factor in the denominator (that does not repeat).

### Example

1. Compute $$\int \frac{2x+2}{(x^2 + 3x + 2)}dx$$
2. Compute $$\int \frac{1}{x^2 +3 x -4}dx$$
3. Compute $$\int \frac{4}{x^2-x}dx$$

---
## Case 2: Repeated Linear Factors

> If the denominator $q(x)$ has repeated linear factors, then the partial fractions are of the form: $$\frac{A_1}{x-a_1} + \frac{A_2}{(x-a_1)^2} + \cdots + \frac{A_n}{(x-a_1)^n}$$ for each repeated linear factor in the denominator and we add the same number of repetitions.

### Example

1. Compute $$\int \frac{1}{x^2-2x+1}dx$$
2. Compute $$\int \frac{1}{x^3-2x^2}dx$$
3. Compute $$\int \frac{1}{(x-1)^3}dx$$

---
## Example

We can mix case 1 and case 2.

1. Compute $$\int \frac{1}{{x^3-2x^2+x}}dx$$
1. Compute $$\int \frac{2x-3}{x^4-4x^2}dx$$
