---
title: "Derivatives as Limit"
subtitle: "Session 04"  
author: 
  - "Alejandro Ucan"
date: '`r Sys.Date()`'
output:
  xaringan::moon_reader:
    css: xaringan-themer.css
    nature:
      slideNumberFormat: "%current%"
      highlightStyle: github
      highlightLines: true
      ratio: 16:9
      countIncrementalSlides: true
---

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
knitr::opts_chunk$set(
  fig.width=9, fig.height=3.5, fig.retina=3,
  out.width = "100%",
  cache = FALSE,
  echo = FALSE,
  message = FALSE, 
  warning = FALSE,
  hiline = TRUE
)
```

```{r xaringan-themer, include=FALSE, warning=FALSE}
library(xaringanthemer)
style_duo_accent(
  primary_color = "#000080",
  secondary_color = "#f2f3f4",
  inverse_header_color = "#00147e"
)
```

```{r xaringan-editable, echo=FALSE}
xaringanExtra::use_editable(expires = 1)
xaringanExtra::use_scribble()
```

```{r xaringan-logo, echo=FALSE}
xaringanExtra::use_logo(
  image_url = "https://github.com/alxcn/TecLogoEIC/blob/9562a53875418e749a296c85808a19c85fc4be74/IngenieriaCiencias_Horizontal_RGB.png",
  position = xaringanExtra::css_position(top = "2em", right = "2em")
)
```
# Goals

* Introduce the concept of mean rate of change of a function. <br/><br/>
* Introduce the concept of derivative of a function. <br/><br/>
* Present the principal theorems of derivatives. <br/><br/>
* Introduce the principal rules of differentiation. <br/><br/>

---
# Motivation 

Previous session we present the instant velocity as a limit of the average velocity. This reasoning is not particular to velocity, but it can be applied to any quantity that changes. For example, we have that the function $f(x)$ models a given quantity (e.g., the population of a city, the temperature of a body, the position of a particle, etc.). <br/><br/>

From a fixed observation point $x_0,$ we can calculate the *average rate of change* of the quantity in the interval $[x_0, x_0 + h].$ This average rate of change is called the **mean rate of change** of the function $f(x)$ at $x_0$ and it is given by $$\frac{f(x_0 + h) - f(x_0)}{h}.$$ <br/><br/>

If we want to measure the instantaneous rate of change, we need to consider limits.

---
# Derivative and differentiation

> **Definition.** Let $f(x)$ be a function defined in an interval $I$ that contains the point $x_0.$ The **derivative** of $f(x)$ at $x_0$ is defined as $$f'(x_0) = \lim_{h \to 0} \frac{f(x_0 + h) - f(x_0)}{h}.$$ If this limit exists, we say that $f(x)$ is **differentiable** at $x_0.$

--
#### Example: 

Let $f(x)= x + 1,$ what is the instantenous rate of change of $f$ at $4.$ <br/><br/>

We have that $f'(4) = \lim_{h \to 0} \frac{f(4 + h) - f(4)}{h} = \lim_{h \to 0} \frac{4 + h + 1 - 4 - 1}{h} = \lim_{h \to 0} \frac{h}{h} = 1.$ <br/><br/>

Therefore, the instantaneous rate of change of $f$ at $4$ is $1.$

---
#### Example: 

Let $f(x)= mx+b$ be any linear function, what is the instantaneous rate of change of $f$ at $x_0?$ <br/><br/>

--
We have that $f'(x_0) = \lim_{h \to 0} \frac{f(x_0 + h) - f(x_0)}{h} = \lim_{h \to 0} \frac{m(x_0 + h) + b - mx_0 - b}{h} = \lim_{h \to 0} \frac{mh}{h} = m.$ <br/><br/>

---
#### Example: 

Let $f(x)= 3x^2$ be function, what is the instantaneous rate of change of $f$ at $x_0?$ <br/><br/>

--
We have that $f'(x_0) = \lim_{h \to 0} \frac{f(x_0 + h) - f(x_0)}{h} = \lim_{h \to 0} \frac{m(x_0 + h) + b - mx_0 - b}{h} = \lim_{h \to 0} \frac{mh}{h} = m.$ <br/><br/>

---
# Theorems of Derivatives

1. **Derivative of a Constant:** $(k)' = 0.$ <br/>
2. **Derivative of a Power:** $(x^n)' = nx^{n-1}.$ <br/>
3. **Derivative of a Sum:** $(f(x) + g(x))' = f'(x) + g'(x).$ <br/>
4. **Derivative of a Product:** $(f(x)g(x))' = f'(x)g(x) + f(x)g'(x).$ <br/>
5. **Derivative of a Quotient:** $\left(\frac{f(x)}{g(x)}\right)' = \frac{f'(x)g(x) - f(x)g'(x)}{[g(x)]^2}.$ <br/>

---

# Rules of differentiation

1. **Derivative of Exponential** $(a^x)' = a^x \ln(a).$ <br/>
2. **Derivative of Logarithm** $(\log_a(x))' = \frac{1}{x \ln(a)}.$ <br/>
3. **Derivative of Trigonometric Functions** $(\sin(x))' = \cos(x),$ $(\cos(x))' = -\sin(x),$ $(\tan(x))' = \sec^2(x).$ <br/>
4. **Derivative of Inverse Trigonometric Functions** $(\arcsin(x))' = \frac{1}{\sqrt{1-x^2}},$ $(\arccos(x))' = -\frac{1}{\sqrt{1-x^2}},$ $(\arctan(x))' = \frac{1}{1+x^2}.$ <br/>

---
#### Examples 1

1. Find the derivative of the function $f(x) = 3x^2 + 2x - 1.$ <br/><br/>

--
We have that $f'(x) = (3x^2)' + (2x)' - 1' = 3(2)x^{2-1} + 2(1)x^{1-1} - 0 = 6x + 2.$ <br/><br/>

---
#### Examples 2

2. Find the derivative of the function $f(x) = (x+1)\sqrt{x}.$ <br/><br/>

--
We have that $f'(x) = [(x+1)\sqrt{x}]' = (x+1)' \sqrt{x} + (x+1)'\frac{1}{2\sqrt{x}} = 1\sqrt{x} + 1\frac{1}{2\sqrt{x}} = \sqrt{x} + \frac{1}{2\sqrt{x}}.$ <br/><br/>

---
#### Example 3

3. Find the derivative of the function $f(x) = \frac{x^2}{x+1}.$ <br/><br/>

--
We have that $f'(x) = \left(\frac{x^2}{x+1}\right)' = \frac{x^2(x+1)' - x^2(x+1)}{(x+1)^2} = \frac{x^2 - x^2}{(x+1)^2} = \frac{1}{(x+1)^2}.$ <br/><br/>

---
#### Example 4

4. Find the derivative of the function $f(x) = x^2(\sin(x) + \cos(x)).$ <br/><br/>

--
We have that $f'(x) = [x^2(\sin(x) + \cos(x))]' =$
$$(x^2)'(\sin(x) + \cos(x)) + x^2(\sin(x) + \cos(x))' = 2x(\sin(x) + \cos(x)) + x^2(\cos(x) - \sin(x)).$$ 

---
#### Example 5

5. Find the derivative of the function $$f(x) = \frac{\sin(x)-e^x+3x}{\sqrt{x}+3\cos(x)-\ln(x)}.$$ <br/><br/>

