---
title: "Introduction to Business Variables"
subtitle: "Session 01"  
author: 
  - "Alejandro Ucan"
date: '`r Sys.Date()`'
output:
  xaringan::moon_reader:
    css: xaringan-themer.css
    nature:
      slideNumberFormat: "%current%"
      highlightStyle: github
      highlightLines: true
      ratio: 16:9
      countIncrementalSlides: true
---

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
knitr::opts_chunk$set(
  fig.width=9, fig.height=3.5, fig.retina=3,
  out.width = "100%",
  cache = FALSE,
  echo = FALSE,
  message = FALSE, 
  warning = FALSE,
  hiline = TRUE
)
```

```{r xaringan-themer, include=FALSE, warning=FALSE}
library(xaringanthemer)
style_duo_accent(
  primary_color = "#000080",
  secondary_color = "#f2f3f4",
  inverse_header_color = "#00147e"
)
```

```{r xaringan-editable, echo=FALSE}
xaringanExtra::use_editable(expires = 1)
xaringanExtra::use_scribble()
```

```{r xaringan-logo, echo=FALSE}
xaringanExtra::use_logo(
  image_url = "https://github.com/alxcn/TecLogoEIC/blob/9562a53875418e749a296c85808a19c85fc4be74/IngenieriaCiencias_Horizontal_RGB.png",
  position = xaringanExtra::css_position(top = "2em", right = "2em")
)
```
# Session's Goals

 * Understand the concept of variable dependencies. <br/><br/>
 * Understand the concept of function and its difference with relations. <br/><br/>
 * Applications of functions in business. <br/><br/>

---
# Motivation

  __Q1:__ When you go to the gym, on what depends the amount of burned calories? <br/><br/>
--
 __Q2:__ When you are uploading a document, on what depends the time it takes to upload? <br/><br/>
--
 __Q3:__ When you pay taxes, on what depends the amount of taxes you pay? <br/><br/>
 
---
# Variable Dependency

> __Definition:__ Given two variables $x$ and $y,$ we say that $y$ _depends_ on $x$ if the value of $y$ is determined by the value of $x.$ <br/><br/> This dependency implies a _relation_ between the possible values of $x$ and $y.$

---
## Functions

> __Definition:__ A _function_ is a relation between two variables $x$ and $y$ such that for each value of $x$ there is a unique value of $y.$ <br/><br/> We write $y=f(x)$ to denote that $y$ is a function of $x.$

---
### Parts of a function:
$$f:D\to R$$ $$x\mapsto y=f(x)$$
* __Domain:__ the set of all possible values of $x$ that can be inputs of the function. <br/><br/>
* __Codomain:__ the set of all possible values of $y$ that can be outputs of the function. <br/><br/>
* __Relation Rule:__ the expression that relates $x$ and $y.$ <br/><br/>
---
## Example 1

> The price of a minute call in AT&T cost $0.79$mxn, then the cost of a call of $x$ minutes is given by the function $f(x)=0.79x.$ <br/><br/>
 * Domain: the set of all positive numbers. <br/><br/> 
 * Codomain: the set of all positive real numbers. <br/><br/> 
 * Relation rule: is given by the expression $f(x)=0.79x.$

---
## Example 2

> The profit of an enterprise depends on time, where its total revenue is $1000t$ and its cost at time $t$ is $300t^2$. <br/><br/>
* Domain: the set of all positive numbers. <br/><br/>
* Codomain: the set of all real numbers. <br/><br/>
* Relation rule: is given by the expression $f(x)=1000t-300t^2.$

---
## Graph of a function

> __Definition:__ the _graph_ of a function is a visual representation of the function $f,$ where the domain is represented in the horizontal axis and the codomain in the vertical axis, and a point $p$ is drawn at the coordinate $(x,f(x)).$<br/><br/>

---
## Example 3

> The graph of the function $f(x)=0.79x$ is given by:

```{r, echo=FALSE, fig.cap="Graph of the function $f(x)=0.79x$"}
library(ggplot2)
ggplot(data.frame(x=c(0,10)), aes(x)) + 
  stat_function(fun = function(x) 0.79*x, geom = "line") +
  labs(x = "Minutes", y = "Cost (mxn)")
```
---
## Example 4


```{r, echo=FALSE, fig.cap="Graph of the Temperature as time function"}
library(ggplot2)
library(dplyr)
library(patchwork) # To display 2 charts together
library(hrbrthemes)

# Build dummy data
data <- data.frame(
  day = as.Date("2019-01-01") + 0:99,
  temperature = runif(100) + seq(1,100)^2.5 / 10000,
  price = runif(100) + seq(100,1)^1.5 / 10
)

# Most basic line chart
p1 <- ggplot(data, aes(x=day, y=temperature)) +
  geom_line(color="#69b3a2", size=2) +
  ggtitle("Temperature: range 1-10") +
  theme_ipsum()
p1
```
---
## The Supply function

> __Definition:__ the _supply function_ is a function that relates the price of a product with the quantity of the product that is supplied to the market. <br/><br/> $$p=f(q)$$ <br/><br/> where $p$ is the price of the product and $q$ is the quantity of the product.

--
<br/><br/>
Notes on Supply function: <br/>
* The supply function is an increasing function. <br/><br/>

---
## The Demand function

> __Definition:__ the _demand function_ is a function that relates the price of a product with the quantity of the product that is demanded to the market. <br/><br/> $$p=f(q)$$ <br/><br/> where $p$ is the price of the product and $q$ is the quantity of the product.

--
<br/><br/>
Notes on Demand function: <br/>
* The demand function is a decreasing function. <br/><br/>

---
### The equilibrium state

> __Definition:__ the _equilibrium state_ is the state where the supply and demand functions intersect.


```{r, echo=FALSE, fig.cap="Example of an Equilibrium State"}
library(ggplot2)
library(dplyr)
library(patchwork) # To display 2 charts together
library(hrbrthemes)
data <- data.frame(
  day = as.Date("2019-01-01") + 0:99,
  temperature = runif(100) + seq(1,100)^2.5/ 10000,
  price = runif(100) + seq(100,1)^1.5 / 10
)

coeff <- 10

ggplot(data, aes(x=day)) +
  
  geom_line( aes(y=temperature)) + 
  geom_line( aes(y=price / coeff)) + # Divide by 10 to get the same range than the temperature
  
  scale_y_continuous(
    
    # Features of the first axis
    name = "First Axis",
    
    # Add a second axis and specify its features
    sec.axis = sec_axis(~.*coeff, name="Second Axis")
  )
```
---
## The Total Cost function

> __Definition:__ the _total cost function_ is a function that relates the quantity of a product with the total cost of producing that quantity. <br/><br/> $$C=f(q)$$ <br/><br/> where $C$ is the total cost and $q$ is the quantity of the product.

---
## The Income function

> __Definition:__ the _income function_ represent the total amount of money obtained by selling a product. <br/><br/> $$I=f(q)$$ <br/><br/> where $I$ is the income and $q$ is the quantity of the product.

---
## The Utility function

> __Definition:__ the _utility function_ is defined as the total income minus the cost functions. <br/><br/> $$U=f(q)$$ <br/><br/> where $U$ is the utility and $q$ is the quantity of the product. <br/><br/> $$U=I-C$$