---
title: "Graphs and Special Functions"
subtitle: "Session 02"  
author: 
  - "Alejandro Ucan"
date: '`r Sys.Date()`'
output:
  xaringan::moon_reader:
    css: xaringan-themer.css
    nature:
      slideNumberFormat: "%current%"
      highlightStyle: github
      highlightLines: true
      ratio: 16:9
      countIncrementalSlides: true
---

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
knitr::opts_chunk$set(
  fig.width=9, fig.height=3.5, fig.retina=3,
  out.width = "100%",
  cache = FALSE,
  echo = FALSE,
  message = FALSE, 
  warning = FALSE,
  hiline = TRUE
)
```

```{r xaringan-themer, include=FALSE, warning=FALSE}
library(xaringanthemer)
style_duo_accent(
  primary_color = "#000080",
  secondary_color = "#f2f3f4",
  inverse_header_color = "#00147e"
)
```

```{r xaringan-editable, echo=FALSE}
xaringanExtra::use_editable(expires = 1)
xaringanExtra::use_scribble()
```

```{r xaringan-logo, echo=FALSE}
xaringanExtra::use_logo(
  image_url = "https://github.com/alxcn/TecLogoEIC/blob/9562a53875418e749a296c85808a19c85fc4be74/IngenieriaCiencias_Horizontal_RGB.png",
  position = xaringanExtra::css_position(top = "2em", right = "2em")
)
```
# Session's Goals

* Introduce the concept of a graph. <br/><br/>
* Modifications of graphs via translations, reflections, and dilations. <br/><br/>
* Introduce the Special Functions and their graphs. <br/><br/>

---
# Motivation

Provide information about some phenomena is it important, but the function notation could not suit for everyone. For this reason, we need a better way to represent the information.

```{r}
library(tidyverse)
library(ggplot2)
gdp <- readr::read_csv('https://raw.githubusercontent.com/rsquaredacademy/datasets/master/gdp.csv')
ggplot(gdp, aes(year, india)) +
  geom_line(color = 'blue')
```


---
# Graphs 

> **Definition:** A graph is a visual representation of data that displays the relationship between two variables vía a function. It is defined as a the pairs of points $(x, f(x))$ in the $\mathbb{R}^2$ plane.

--

##### Steps to graph a function
1. Find the domain of the function.
2. Choose some sample points in the domain.
3. Evaluate the function at the sample points.
4. Plot the points in the Cartesian plane.
5. Connect the points with a smooth curve.

---
# Examples:

Find the graph of the following functions:
1. $f(x) = x^2+3x-4$
2. $g(x) = \sqrt{x+4}$
3. $h(x) = \frac{1}{x-2}$

---
# Remarks on graphs

1. The graph of a polynomial function depends on the degree of the polynomial and the leading coefficient. 

2. The graph of a root function depends on the oddity of the root and the sign of the radicand.

3. The graph of a rational function depends on the degree of the numerator and the denominator, and their assymptotes.

---
### Assymptotes

> **Definition:** A vertical line $x=a$ is a vertical assymptote of a function $f(x)=\frac{p(x)}{q(x)}$ if $q(a)=0$ and $p(a)\neq 0.$ <br/><br/>

--
##### Remarks:

For a rational function, the vertical assymptotes could implies a chaotic behavior of the function. 

---
# Common Graphs:

<iframe src="https://www.wolframcloud.com/obj/alejandroucan-puc/Published/FunctionsCatalog.nb?_embed=iframe" width="600" height="800"></iframe>
---
## Transformation of Graphs

1. **Traslations** 

Given a function $f(x),$ the traslation of $f(x)$ by $a$ units to the right is the function $f(x-a).$ The traslation of $f(x)$ by $a$ units to the left is the function $f(x+a)$ for any $a>0.$ <br/><br/>

Given a function $f(x),$ the traslation of $f(x)$ by $b$ units up is the function $f(x)+b.$ The traslation of $f(x)$ by $b$ units down is the function $f(x)-b$ for any $b>0.$

---
2. **Reflections**

Given a function $f(x),$ the reflection of $f(x)$ about the $x$-axis is the function $-f(x).$ The reflection of $f(x)$ about the $y$-axis is the function $f(-x).$ <br/><br/>

3. **Dilations**

Given a function $f(x),$ the dilation of $f(x)$ by a factor of $k>1$ is the function $k\cdot f(x).$ The dilation of $f(x)$ by a factor of $k<1$ is the function $\frac{1}{k}\cdot f(x).$

---
# Examples:

<iframe src="https://www.wolframcloud.com/obj/alejandroucan-puc/Published/GraphModifications.nb?_embed=iframe" width="600" height="800"></iframe>
---
## Special Functions

1. **Exponential Function:** $f(x)=a^x,$ where $a>0$ and $a\neq 1.$ <br/><br/>

Domain: $\mathbb{R},$ Codomain: $\mathbb{R}^+.$

Applications: Growth and decay of populations, radioactive decay, etc.

---
## Special Functions

2. **Logarithmic Function:** $f(x)=\log_a(x),$ where $a>0$ and $a\neq 1.$ <br/><br/>

Domain: $\mathbb{R}^+,$ Codomain: $\mathbb{R}.$

Applications: Measures of sound intensity, measures of earthquake intensity, etc.

---
## Special Functions

3. **Trigonometric Functions:** $f(x)=\sin(x),$ $f(x)=\cos(x),$ $f(x)=\tan(x),$ etc. <br/><br/>

Domain: $\mathbb{R},$ Codomain: $\mathbb{R}.$ 

Applications: Sound waves, light waves, etc.

---
## Special Functions

4. **Inverse Trigonometric Functions:** $f(x)=\arcsin(x),$ $f(x)=\arccos(x),$ $f(x)=\arctan(x),$ etc.

Domain: it depends on the function, Codomain: $\mathbb{R}.$

Applications: Measures of angles, etc.



