---
title: "Eigenfunctions for Systems of LODEs"
subtitle: "Session 07"
author: 
  - "Alejandro Ucan"
date: '`r Sys.Date()`'
output:
  xaringan::moon_reader:
    css: xaringan-themer.css
    nature:
      slideNumberFormat: "%current%"
      highlightStyle: github
      highlightLines: true
      ratio: 16:9
      countIncrementalSlides: true
---

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
knitr::opts_chunk$set(
  fig.width=9, fig.height=3.5, fig.retina=3,
  out.width = "100%",
  cache = FALSE,
  echo = FALSE,
  message = FALSE, 
  warning = FALSE,
  hiline = TRUE
)
```

```{r xaringan-themer, include=FALSE, warning=FALSE}
library(xaringanthemer)
style_duo_accent(
  primary_color = "#000080",
  secondary_color = "#f2f3f4",
  inverse_header_color = "#00147e"
)
```

```{r xaringan-editable, echo=FALSE}
xaringanExtra::use_editable(expires = 1)
xaringanExtra::use_scribble()
```

```{r xaringan-logo, echo=FALSE}
xaringanExtra::use_logo(
  image_url = "https://github.com/alxcn/TecLogoEIC/blob/9562a53875418e749a296c85808a19c85fc4be74/IngenieriaCiencias_Horizontal_RGB.png",
  position = xaringanExtra::css_position(top = "2em", right = "2em")
)
```
# Goals

* Use eigenvalues and eigenvectors to solve systems of ODEs. <br/><br/>
* Use the Wronskian to test for linear independence. <br/><br/>

---
# Conventions for Systems of ODEs

> In what follows we will deal with systems of two equations with two unknown functions whose ODEs are linear with constant coefficients. This is described in the general form $$\begin{array}{c} x'= a_{11}x+a_{22}y+f_1(t) \\ y'=a_{21}x+a_{22}y+f_2(t) \end{array}$$ where $a_{ij}$ are constants and $f_1(t)$ and $f_2(t)$ are functions of $t.$

---
# Solving for Homogeneous systems

> Given the system $$X'=AX$$ consider the matrix $A.$ Suppose that $(\lambda,v)$ is an eigenvalue-eigenvector pair of $A$ and assume that $\lambda$ is real. Let $F=e^{\lambda t}v$ be a vector of functions, then the vector $F$ is a solution vector of the system.


---
#### Example: 

> Consider the system $$X'=AX$$ where $$A=\left(\begin{array}{cc} 3 & 1 \\ 0 & -3 \end{array}\right).$$ Find the eigenvalues and eigenvectors of $A$ and construct the general solution.

---
#### Example: 

> Consider the system $$X'=\left(\begin{array}{cc} 1 & 2 \\ -2 & -3 \end{array}\right) X.$$ Find the eigenvalues and eigenvectors of $A$ and construct the general solution.

---
## Test for linear independence

> __Definition:__ Consider the solution function vectors $X_1$ and $X_2$ for a system of ODEs. The __Wronskian__ of the solutions is the determinant of the matrix $$\left(\begin{array}{cc} x_1 & x_2 \\ y_1  & y_2\end{array}\right)$$ and is denoted by $W(X_1,X_2).$
<br/><br/>
> __Theorem [Abel's]:__ The solutions $X_1$ y $X_2$ are __linearly independents__ if $$W(X_1,X_2)\neq 0.$$

---
## Solution for Systems of ODEs

> __Method of Eigenfunctions:__ For the system $X'=AX.$ 
  1. Find the eigenvalues and eigenvectors of $A.$ 
      * If the eigenvalues are real and distinct, then the general solution $$X_1=e^{\lambda_1}v_1\mbox{ and } X_2=e^{\lambda_2}v_2.$$
      * If the eigenvalues are real and repeated, then the general solution $$X_1=e^{\lambda}v_1\mbox{ and } X_2=te^{\lambda}v_2.$$
      * If the eigenvalues are complex conjugated, then the general solution $$X_1=e^{\alpha t}\left(\cos(\beta t)v_1+\sin(\beta t)v_2\right)\mbox{ and } X_2=e^{\alpha t}\left(\cos(\beta t)v_1-\sin(\beta t)v_2\right).$$
      
  2. Express the general solution as $$X=c_1 X_1+c_2 X_2.$$
  3. If there are initial conditions, use them to find the constants $c_1$ and $c_2.$
  

---
#### Example 1:

Find the general solution for $X'=AX$ if $A=\left(\begin{array}{cc} 1 & 2 \\ 4 & 3 \end{array}\right).$ <br/><br/>
  
  * $A=\left(\begin{array}{cc} -4 & 2 \\ -\frac{5}{2} & 2 \end{array}\right).$ <br/><br/>

---
#### Example 2:

Find the general solution for $X'=AX$ if $A=\left(\begin{array}{cc} 10 & -5 \\ 8 & -12 \end{array}\right).$ <br/><br/>

---
#### Example 3:

Find the general solution for $A=\left(\begin{array}{cc} 3 & -1 \\ 9 & -3 \end{array}\right).$ <br/><br/>

---