---
title: "Matrix Inverse"
subtitle: "Session 04"  
author: 
  - "Alejandro Ucan"
date: '`r Sys.Date()`'
output:
  xaringan::moon_reader:
    css: xaringan-themer.css
    nature:
      slideNumberFormat: "%current%"
      highlightStyle: github
      highlightLines: true
      ratio: 16:9
      countIncrementalSlides: true
---

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
knitr::opts_chunk$set(
  fig.width=9, fig.height=3.5, fig.retina=3,
  out.width = "100%",
  cache = FALSE,
  echo = FALSE,
  message = FALSE, 
  warning = FALSE,
  hiline = TRUE
)
```

```{r xaringan-themer, include=FALSE, warning=FALSE}
library(xaringanthemer)
style_duo_accent(
  primary_color = "#000080",
  secondary_color = "#f2f3f4",
  inverse_header_color = "#00147e"
)
```

```{r xaringan-editable, echo=FALSE}
xaringanExtra::use_editable(expires = 1)
xaringanExtra::use_scribble()
```

```{r xaringan-logo, echo=FALSE}
xaringanExtra::use_logo(
  image_url = "https://github.com/alxcn/TecLogoEIC/blob/9562a53875418e749a296c85808a19c85fc4be74/IngenieriaCiencias_Horizontal_RGB.png",
  position = xaringanExtra::css_position(top = "2em", right = "2em")
)
```
# Goals

* Zero Divisors. <br/><br/>
* Matrix Inverse algorithm. <br/><br/>
* Criteria for Matrix Inverse Existence. <br/><br/>


---
# Is there multiplicative inverses?


> Recall that in $\mathbb{R}$ for each $x\neq 0$ there exists a unique number $y$ such that $xy=1,$ this number is known as the multiplicative inverse. <br/><br/>
> Also, in $\mathbb{R}$ we have the property that $xy=0$ implies $x=0$ or $y=0$. <br/><br/>

--
> Is it possible to define the multiplicative inverse for matrices? <br/><br/>

Compute the following product: $$\begin{bmatrix} 1 & 1 \\ 1 & 1 \end{bmatrix} \begin{bmatrix} -1 & 1 \\ 1 & -1 \end{bmatrix}$$

---
# Is there multiplicative inverses?

Consider the matrix $$A=\begin{bmatrix} 1 & 0 \\ 2 & 0 \end{bmatrix},$$ if there is a matrix $B$ such that $AB=I_{2\times 2}$, then we can write $$\begin{bmatrix} 1 & 0 \\ 2 & 0 \end{bmatrix} \begin{bmatrix} a & b \\ c & d \end{bmatrix} = \begin{bmatrix} 1 & 0 \\ 2 & 0 \end{bmatrix}$$ where $I_{2\times 2}$ is the identity matrix of size $2\times 2$. <br/><br/> 

---
## Non-Singular Matrices

> **Definition:** A square matrix $A$ of size $n\times n$ is said to be non-singular if there exists a matrix $B$ of size $n\times n$ such that $$AB=BA=I_{n\times n}$$ where $I_{n\times n}$ is the identity matrix of size $n\times n$. <br/><br/>

--

##### Example 1: 

Verify that the following matrices are inverses: $A=\begin{bmatrix} 1 & 0 \\ 2 & 1 \end{bmatrix}$ Addition $B=\begin{bmatrix} 1 & 0 \\ -2 & 1 \end{bmatrix}.$

---

##### Example 2:

Verify that the following matrices are inverses: $A=\begin{bmatrix} 1 & 0 & 1 \\ 0 & 1 & 0 \\ 1 & 0 & -1\end{bmatrix}$ and $B=\begin{bmatrix} 1/2 & 0 & 1/2 \\ 0 & 1 & 0 \\ 1/2 & 0 & -1/2 \end{bmatrix}$

---
##### Example 3:

Find the nverse matrix of $$A= \begin{bmatrix} 3 & 2 \\ -2 & 1 \end{bmatrix}$$

---
## Gauss Elimination for Finding the Inverse

> As the process of finding the inverse is similar to the process of solving a system of linear equations, we can use the same method. <br/><br/>


--

##### Example 4:

Compute the inverse of the matrix $$A=\begin{bmatrix} 5 & 1 \\ 4 & 1 \end{bmatrix}$$ using Gauss elimination. <br/><br/>


---
# Algorithm


1. Use Gaussian Elimination to reduce the augmented matrix $[A|I]$ to a upper triangular form. <br/><br/>

2. Continue the reduction but now making zero the entries above the pivots (make the left matrix the identity), you will obtain a matrix of the form $[I|B]$. <br/><br/>
the form $[I|B]$, where $I$ is the identity matrix and $B$ is the inverse of $A$. <br/><br/>



---
##### Practice:

Find the inverse of the following matrices using the Gauss elimination method:

1. $A=\begin{bmatrix} 1 & 0 & -1 \\ 1 & 0 & 2 \\ 0 & 1 & 0 \end{bmatrix}$

2. $A=\begin{bmatrix} 1 & 1 & 0 \\ 1 & 0 & 2 \\ 0 & 1 & 0 \end{bmatrix}$

3. $A=\begin{bmatrix} 1 & 0 & -1 \\ 1 & 1 & 0 \\ 0 & 1 & 0 \end{bmatrix}$

---
# Why we can use the inverse?

Assume we have a system of linear equations (SLE) $$Ax=b$$ and that $A$ is an invertible matrix. Due to the equality, we can multiply both sides (from the right) by the inverse and calculate the solution "directly": $$x=Ix= (A^{-1} A)x= A^{-1} (Ax)=A^{-1}b.$$ The fact that my matrix is invertible restricts my SLE to have only one solution.

---
#### Practice:

With the previous matrices, compute the solution of the following systems of linear equations $Ax=b$ where $b$ are:

1. $b=\begin{bmatrix} 1 \\ 0 \\ 1 \end{bmatrix}$

2. $b=\begin{bmatrix} 10 \\ -10 \\ 10 \end{bmatrix}$

3. $b=\begin{bmatrix} 500 \\ 420 \\ 120 \end{bmatrix}$

---
# How to know if a matrix is invertible?

Until now we do not know how to decide if my matrix is invertible other than performing the method to calculate the inverse. However, what would happen if I have a $10000 \times 10000$ system? I cannot perform the process and at the end it turns out that it is not invertible.

--

> **Definition:** The *determinant* of a square matrix $A$ is a number associated with the matrix that allows us to determine if the matrix is invertible or not. The criterion is: if the determinant is different from zero, then my matrix is invertible (non-singular), and if it is zero, then my matrix is not invertible (singular).



---
## Determinant of a $2\times 2$ Matrix


> **Definition:** Let $A=\begin{bmatrix} a & b \\ c & d \end{bmatrix}$ be a matrix of size $2\times 2$, then the determinant of $A$ is defined as $$\det(A)=ad-bc.$$


--

##### Example 5: 

Compute the determinant of the following matrices: 
1. $\begin{bmatrix} 1 & 2 \\ 3 & 4 \end{bmatrix}$
2. $\begin{bmatrix} 2 & 1 \\ -4 & -2 \end{bmatrix}$
3. $\begin{bmatrix} 1 & 0 \\ 0 & 1 \end{bmatrix}$
4. $\begin{bmatrix} 1 & 0 \\ 0 & 0 \end{bmatrix}$

---
## Determinant of a $3\times 3$ Matrix


> **Definition:** Let $A=\begin{bmatrix} a & b & c \\ d & e & f \\ g & h & i \end{bmatrix}$ be a matrix of size $3\times 3$, then the determinant of $A$ is defined as $$\det(A)=a(ei-fh)-b(di-fg)+c(dh-eg).$$


--
1. $\begin{bmatrix} 1 & 2 & 3 \\ 4 & 5 & 6 \\ 7 & 8 & 9 \end{bmatrix}$
2. $\begin{bmatrix} 1 & 0 & 0 \\ 0 & 1 & 0 \\ 0 & 0 & 1 \end{bmatrix}$
3. $\begin{bmatrix} 1 & 0 & 0 \\ 0 & 1 & 0 \\ 0 & 0 & 0 \end{bmatrix}$

---
## Cofactor de una Matriz de $n\times n$

> **Definición:** Sea $A$ una matriz de $n\times n$ y $a_{ij}$ la entrada en la fila $i$ y columna $j$ entonces el cofactor de $a_{ij}$ es $$C_{ij}=(-1)^{i+j}M_{ij}$$ donde $M_{ij}$ es el determinante de la matriz obtenida al eliminar la fila $i$ y columna $j$ de $A$.

---
## Matriz de Cofactores e Inversa por Cofactores

> **Definición:** Sea $A$ una matriz de $n\times n$ entonces la matriz de cofactores de $A$ es la matriz $C$ de $n\times n$ cuya entrada $c_{ij}$ es el cofactor de $a_{ij}.$

--

> **Definición:** Sea $A$ una matriz de $n\times n$ y $\det(A)\neq 0$ entonces la inversa de $A$ es $$A^{-1}=\frac{1}{\det(A)}C^T$$ donde $C$ es la matriz de cofactores de $A$ y $C^T$ es la transpuesta de $C.$

---
## Ejemplo de Inversa por Cofactores

Calculemos la inversa de la matriz: $$A=\begin{bmatrix} 1 & 2 & 3 \\ 0 & 1 & 4 \\ 5 & 6 & 0 \end{bmatrix}$$





