---
title: "Diagonalization"
subtitle: "Session 07"
author: 
  - "Alejandro Ucan"
date: '`r Sys.Date()`'
output:
  xaringan::moon_reader:
    css: xaringan-themer.css
    nature:
      slideNumberFormat: "%current%"
      highlightStyle: github
      highlightLines: true
      ratio: 16:9
      countIncrementalSlides: true
---

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
knitr::opts_chunk$set(
  fig.width=9, fig.height=3.5, fig.retina=3,
  out.width = "100%",
  cache = FALSE,
  echo = FALSE,
  message = FALSE, 
  warning = FALSE,
  hiline = TRUE
)
```

```{r xaringan-themer, include=FALSE, warning=FALSE}
library(xaringanthemer)
style_duo_accent(
  primary_color = "#000080",
  secondary_color = "#f2f3f4",
  inverse_header_color = "#00147e"
)
```

```{r xaringan-editable, echo=FALSE}
xaringanExtra::use_editable(expires = 1)
xaringanExtra::use_scribble()
```

```{r xaringan-logo, echo=FALSE}
xaringanExtra::use_logo(
  image_url = "https://github.com/alxcn/TecLogoEIC/blob/9562a53875418e749a296c85808a19c85fc4be74/IngenieriaCiencias_Horizontal_RGB.png",
  position = xaringanExtra::css_position(top = "2em", right = "2em")
)
```
# Goals of the Session

* Basis change matrices. <br/><br/>
* Linear transformations diagonalization. <br/><br/>

---
# Basis Change Matrices

> Assume we have a vector space $V$ and two bases for $V,$ say $B_1$ and $B_2.$ The basis change matrix from $B_1$ to $B_2$ is the matrix $P$ such that $[v]_{B_2}=P[v]_{B_1}$ for all $v\in V.$

--
##### Example

> Compute the basis change matrix of $\mathbb{R}^2$ with bases $\{(1,0),(0,1)\}$ and $\{(-3,2),(2,3)\}.$
--

##### Solution:
The basis change matrix is $$P=\left(\begin{array}{cc} -3 & 2 \\ 2 & 3 \end{array} \right).$$

---
##### Example

> Compute the basis change matrix of $\mathbb{R}^2$ with bases $\{(1,2),(3,4)\}$ and $\{(1,1),(1,-1)\}.$ <br/><br/>

--
<br/<br/>
Write the basis $\{(1,2),(3,4)\}$ in linear combinations of $\{(1,1),(1,-1)\}:$

$$\left(\begin{array}{cc|cc} 1 & 1 & 1 & 3 \\ 1 & -1 & 2 & 4 \end{array}\right)\sim \left(\begin{array}{cc|cc} 1 & 1 & 1 & 3 \\ 0 & 1 & \frac{-1}{2} & \frac{-1}{2} \end{array}\right)\sim \left(\begin{array}{cc|cc} 1 & 0 & \frac{3}{2} & \frac{7}{2} \\ 0 & 1 & \frac{-1}{2} & \frac{-1}{2} \end{array}\right)$$

---
# Diaganolization of Linear Transformations

> **Definition:** Let $T:V\to V$ be a linear transformation. We say that $T$ is diagonalizable if there exists a basis $B$ of $V$ such that $[T]_B$ is a diagonal matrix. <br/><br/>

--
<br/><br/> 

In the case of matrices, a matrix $A$ is diagonalizable if there exists an invertible matrix $P$ such that $P^{-1}AP$ is diagonal. Then, if the eigenvectors of $A$ form a basis, the basis change matrix from the canonical to the eigenvectors is the matrix that works for diagonalization.

---
##### Example

Find the diagonalization of the transformation $T(x,y)=(5x-12y,-12x-5y)$ and the basis change matrix for the diagonalization.

--
<br/><br/>
##### Solution:

The characteristic polynomial of $T$ is $\lambda^2-169$ whose roots are $13$ and $-13.$ The associated eigenvectors are $(-2,3)$ and $(3,2).$ The basis change matrix from the standard basis to the eigenvectors is $$P=\left(\begin{array}{cc} -2 & 3 \\ 3 & 2 \end{array}\right).$$ So the diagonalization is $$T=\left\{\left(\begin{array}{cc} 13 & 0 \\ 0 & -13 \end{array}\right),\left(\begin{array}{cc} -2 & 3 \\ 3 & 2 \end{array}\right)\right\}.$$

---
##### Example

Find the diagonalization of the transformation $T(x,y)=(7x-2y,2x+2y)$ and the basis change matrix for the diagonalization.
--
<br/><br/>
##### Solution:

The characteristic polynomial of $T$ is $\lambda^2+9\lambda-18$ and its roots are $3$ and $6.$ The eigenvectors of this transformation are $(1,2)$ and $(2,1).$ The basis change matrix from the standard basis to the eigenvectors is $$P=\left(\begin{array}{cc} 1 & 2 \\ 2 & 1 \end{array}\right).$$ So the diagonalization is $$T=\left\{\left(\begin{array}{cc} 3 & 0 \\ 0 & 6 \end{array}\right),\left(\begin{array}{cc} 1 & 2 \\ 2 & 1 \end{array}\right)\right\}.$$
