---
title: "Linear Transformations"
subtitle: "Session 03" 
author: 
  - "Alejandro Ucan"
date: '`r Sys.Date()`'
output:
  xaringan::moon_reader:
    css: xaringan-themer.css
    nature:
      slideNumberFormat: "%current%"
      highlightStyle: github
      highlightLines: true
      ratio: 16:9
      countIncrementalSlides: true
---

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
knitr::opts_chunk$set(
  fig.width=9, fig.height=3.5, fig.retina=3,
  out.width = "100%",
  cache = FALSE,
  echo = FALSE,
  message = FALSE, 
  warning = FALSE,
  hiline = TRUE
)
```

```{r xaringan-themer, include=FALSE, warning=FALSE}
library(xaringanthemer)
style_duo_accent(
  primary_color = "#000080",
  secondary_color = "#f2f3f4",
  inverse_header_color = "#00147e"
)
```

```{r xaringan-editable, echo=FALSE}
xaringanExtra::use_editable(expires = 1)
xaringanExtra::use_scribble()
```

# Goals for the Session

* Introduce the concept of Linear Transformation. <br/><br/>
* Demonstrate that a function is a linear transformation. <br/><br/>
* Calculate the preferred subspaces: Kernel and Image. <br/><br/>
* Dimension Theorem.

---
# Motivation
A factory has a production process that transforms three raw materials into two finished product types, the products needs specific quantities of raw material. The transformation can be represented by the matrix $$T=\begin{pmatrix} 1 & 2 & 0 \\ 3 & 0 & 1 \end{pmatrix}.$$ Negative numbers indicates the needs of raw materials or lack of production. Determine the all possible outputs of the factory based on the raw materials available. Is there a way that the factory can produce a specific amount of products? <br/><br/>

---
# Linear Transformations

> **Definition:** Let $V$ and $W$ be vector spaces. A function $T:V\rightarrow W$ is called a __linear transformation__ if for all $u,v\in V$ and for all scalar $\alpha\in \mathbb{R}$ it holds that: $$T(u+v)=T(u)+T(v)$$ $$T(\alpha u)=\alpha T(u).$$

--

If a function *does not* at least one of these properties, it is not a linear transformation.

---

##### Example 1:

The function $T:\mathbb{R}^2\rightarrow \mathbb{R}^2$ defined by $T(x,y)=(x+y,2x-y)$ is a linear transformation, because $$T(x_1+x_2,y_1+y_2)=(x_1+x_2+y_1+y_2,2(x_1+x_2)-(y_1+y_2))$$ $$=(x_1+y_1,2x_1-y_1)+(x_2+y_2,2x_2-y_2)=T(x_1,y_1)+T(x_2,y_2)$$ and $$T(\alpha x_1,\alpha y_1)=(\alpha x_1+\alpha y_1,2\alpha x_1-\alpha y_1)=\alpha(x_1+y_1,2x_1-y_1)=\alpha T(x_1,y_1).$$

---
##### Example 2:

The function $T:\mathbb{R}^3\rightarrow P_2$ defined by $T(x,y,z)=x+yt+zt^2$ is a linear transformation, because $$T(x_1+x_2,y_1+y_2,z_1+z_2)=x_1+x_2+(y_1+y_2)t+(z_1+z_2)t^2=(x_1+y_1t+z_2t^2)+(x_2+y_2t+z_2t^2)$$ $$=T(x_1,y_1,z_1)+T(x_2,y_2,z_2)$$ and $$T(\alpha x_1,\alpha y_1,\alpha z_1)=\alpha x_1+\alpha y_1t+\alpha z_1t^2= \alpha(x_1+y_1t+z_1t^2)=\alpha T(x_1,y_1,z_1).$$
--
<br/>

##### Counterexample 1:

The function $T:\mathbb{R}\rightarrow \mathbb{R}^2$ defined by $T(x)=(x,1)$ is not a linear transformation, because $T(1+2)=T(3)=(3,1)$ meanwhile $T(1)+T(2)=(1,1)+(2,1)=(3,2).$ Also, $T(2\cdot 2)=(2\cdot 2,1)=(4,1)$ meanwhile $2\cdot T(2)=(2\cdot 2,2)=(4,2).$

---
##### Ejemplo 3

Let's prove that the function $T:P_3\to \mathbb{R}$  given by $T(p)=p(0)$ is a linear transformation.

--
<br/><br/>
**Solution:** Let $p,q\in P_3$ and $\alpha\in \mathbb{R}$, then $$T(\lambda p+q)=(\lambda p+q)(0)=(\lambda a_0+b_0+(\lambda a_1+b_1)(0)+(\lambda a_2+b_2)(0)^2=\lambda a_0+b_0=\lambda T(p)+T(q).$$

---
## Kernel and Range

> **Definition:** Let $T:V\to W$ be a linear transformation: <br/><br/>
  * __The kernel:__ is the subspace of $V$ defined by all vectors that are mapped to zero under $T$ $$N(T)=\{v\in V:T(v)=0\}.$$  Its dimension is called *nullity*<br/><br/>
  * __The range:__ is the subspace of $W$ defined by all vectors of the form $T(v)$ for $v\in V$ $$R(T)=\{T(v):v\in V\}.$$ Its dimension is called *rank*.

---
##### Example 4:

Consider the following linear transformation $T: \mathbb{R}^2\to \mathbb{R}^2$ defined by $T(x,y)=(x+y,2x-y)$. <br/><br/>

--
* The kernel of $T$ is the set of all vectors $(x,y)$ such that $T(x,y)=(0,0)$, that is, $x+y=0$ and $2x-y=0$ $$\left(\begin{array}{cc|c} 1 & 1 & 0\\ 2 & -1 & 0\end{array}\right)\sim \left(\begin{array}{cc|c} 1 & 1 & 0\\ 0 & -3 & 0\end{array}\right)\sim \left(\begin{array}{cc|c} 1 & 0 & 0\\ 0 & 1 & 0\end{array}\right)$$ therefore, $N(T)=\{(0,0)\}$ and $\dim N(T)=0$.

In order to compute the rank of $T$ we take a generic vector $(u,v)$ from $\mathbb{R}^2$ and solve the system $$\left(\begin{array}{cc|c} 1 & 1 & u\\ 2 & -1 & v\end{array}\right)\sim \left(\begin{array}{cc|c} 1 & 1 & u\\ 0 & -3 & v-2u\end{array}\right)\sim \left(\begin{array}{cc|c} 1 & 0 & \frac{v+2u}{3}\\ 0 & 1 & \frac{2u-v}{3}\end{array}\right)$$ therefore, $R(T)=\{(u,v)\in \mathbb{R}^2:u=\frac{y+2x}{3},v=\frac{2x-y}{3}\}$ and $\dim R(T)=2$.

---
##### Example 5:

Consider the linear transformation $D:P_2\to P_1$ defined by $D(p)=p'$. <br/><br/>
--
* $D(a_0+a_1x+a_2x^2)=a_1+2a_2x=0+0x$ if and only if $a_1=0$ and $a_2=0$, therefore, $N(D)=\{p\in P_2:a_1=a_2=0\}=\{a_1\}$ and $\dim N(D)=1$.

* To find the range of $D$ we take a generic vector $p=a_0+a_1x$ from $P_1$ and calculate $$D(b_0+b_1x+b_2x^2)=b_1+2b_2x=a_0+a_1x$$, therefore: $a_0=b_1$ and $b_2=\frac{a_1}{2}$, therefore, $R(D)=P_1$ and $\dim R(D)=2$.

---
##### Example 6

Conider the following linear transformation $T:\mathbb{R}^2\to \mathbb{R}^3$ defined by $T(x,y)=A(x,y)$ where $$A=\left(\begin{array}{cc} 0 & 1 \\ 2 & 1 \\ 1 & 3\end{array}\right).$$ Find the kernel and range of $T$.

---
## Dimension Theorem

> **Theorem:** Let $T:V\to W$ be a linear transformation, then $$\dim V=\dim N(T)+\dim R(T).$$

---
### Applied Linear Transformations

> A factory has a production process that transforms three raw materials into two finished product types, the products needs specific quantities of raw material. The transformation can be represented by the matrix $$T=\begin{pmatrix} 1 & 2 & 0 \\ 3 & 0 & 1 \end{pmatrix}.$$ Negative numbers indicates the needs of raw materials or lack of production. Determine the all possible outputs of the factory based on the raw materials available. Is there a way that the factory can produce a specific amount of products?

---
### Applied Linear Transformations

> A DC motor's behavior is changes Voltage (V) and Current (I) to produce Angular moment ( $\omega$ ), Torque (T) and Temperature (S). This change is via the matrix $$T=\begin{pmatrix} 2 & 3 \\ 1 & -1 \\ 0 & 4 \end{pmatrix}.$$ Negative numbers indicates the needs of Voltage or Current. Determine all possible inputs such that we have a zero output. Interpret the results in terms of the motor's behavior. 