---
title: "Cylindrical Coordinates"
subtitle: "Session 13"
author: 
  - "Alejandro Ucan"
date: '`r Sys.Date()`'
output:
  xaringan::moon_reader:
    css: xaringan-themer.css
    nature:
      slideNumberFormat: "%current%"
      highlightStyle: github
      highlightLines: true
      ratio: 16:9
      countIncrementalSlides: true
---

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
knitr::opts_chunk$set(
  fig.show =('asis'),
  dev='png',
	message = FALSE,
	warning = FALSE,
	cache = FALSE,
	hiline = TRUE,
	out.width = "90%"
)
```

```{r xaringan-themer, include=FALSE, warning=FALSE}
library(xaringanthemer)
style_duo_accent(
  primary_color = "#000080",
  secondary_color = "#f2f3f4",
  inverse_header_color = "#00147e"
)
```

```{r xaringan-editable, echo=FALSE}
xaringanExtra::use_editable(expires = 1)
xaringanExtra::use_scribble()
```
# Goals for the session:

 * Define the Cylindrical coordinate system. <br/><br/>
 * Introduce the volume differential in cylindrical coordinates. <br/><br/>
 * Compute triple integrals in cylindrical coordinates.


---
# Cylindrical Coordinates $(r,\theta,z).$

<iframe src="https://www.geogebra.org/calculator/za424zgv?embed" width="1200" height="450" allowfullscreen style="border: 1px solid #e4e4e4;border-radius: 4px;" frameborder="0"></iframe>

---
# How to translate between coordinates?

There are expressions that allow us to convert between different types of coordinates. For example:

> Given a point $(x,y,z)$ (rectangular coordinates) and I want it in _cylindrical coordinates_, then it satisfies: <br/>
- $r^2= x^2+y^2,$ <br/>
- $\tan(\theta)=\frac{y}{x}$ <br/>
- $z=z.$

<br/>

> Given a point $(r,\theta,z)$ and I want it in _rectangular coordinates_, then it satisfies: <br/>
- $x=r\cdot \cos(\theta)$ <br/>
- $y=r\cdot \sin(\theta)$  
- $z=z.$

---
# How my Volume differential changes?

> The volume differential changes to $$dV=r dr d\theta dz.$$

---
## Lets express 3D regions in cylindrical coordinates.

> Describe the region bounded by the plane $z=0$ and the paraboloid $z=4-x^2-y^2$ in cylindrical coordinates, and compute its volumen.
---

## Lets express 3D regions in cylindrical coordinates.

> Describe the region enclosed by the sphere $x^2+y^2+z^2=4$ and the cylinder $x^2+y^2=2.$
---
# Example

> Compute the forth-dimensional volume of the region whose height is given by the function $f(x,y,z)=x^2+y^2$ and base the 3-dimensional region bounded by $x^2+y^2=2z$ and $z=2.$ <br/><br/>

--
Recall that for the paraboloid, the level curves are given by circles $x^2+y^2=k^2$ which is a circle. The shadow of the biggest circle that appears is the one at level $k^2=2.$ Therefore, my radious takes values from $[0,\sqrt{2}]$ and $\theta$ from $[0,2\pi].$
Changing my integral to cylindrical coordinates, I have: $$\int_0^{2\pi}\int_0^{\sqrt{2}}\int_{r^2}^2 r^2 r dz dr d\theta$$
$$=\int_0^{2\pi} \int_0^2 r^3\left(2-\frac{r^2}{2}\right) dr d\theta = \int_0^{2\pi}\frac{r^4}{2}-\frac{r^6}{12}\big{|}_{0}^2 d\theta$$
$$=\frac{8}{3}\int_0^{2\pi}d\theta = \frac{16\pi}{3}.$$

---
# Exercise:

> Compute the volume of the region enclosed by the paraboloids $x^2+y^2=4z$ and $x^2+y^2=-2z+6$

---
# Solution:

> The described region is: $0\leq r\leq 2,$ $0\leq \theta \leq 2\pi$ and $\frac{r^2}{4}\leq r \leq \frac{6-r^2}{2}.$ <br/><br/> 
Therefore my integral is given by $$\int_0^{2\pi}\int_0^2\int_{\frac{r^2}{4}}^{\frac{6-r^2}{2}}dzdrd\theta=\int_0^{2\pi}\int_0^2 \frac{12-r^2}{4} dr d\theta$$
$$\int_0^{2\pi}3r-\frac{r^3}{12}\big{|}_0^2 d\theta =2\pi \left(6-\frac{8}{12}\right)=\frac{32\pi}{3}.$$

---
# Activity:

> Compute the following integrals in cylindrical coordinates: <br/><br/>
 * $f(x,y,z)=x^2+y^2$ region bounded by the cone $z^2=x^2+y^2$ and $z=2.$ <br/><br/>
  * $f(x,y,z)=\sqrt{x^2+y^2}$ in the region bounded by the cylinder $x^2+y^2=16$ between $z=-5$ y $z=4.$ <br/><br/>
  * $f(x,y,z)=e^z$ in the region bounded $z=1+x^2+y^2,$ the cylinder $x^2+y^2=5$ and the plane $z=0.$
  
