---
title: "Vector Spaces"
subtitle: "Session 01" 
author: 
  - "Alejandro Ucan"
date: '`r Sys.Date()`'
output:
  xaringan::moon_reader:
    css: xaringan-themer.css
    nature:
      slideNumberFormat: "%current%"
      highlightStyle: github
      highlightLines: true
      ratio: 16:9
      countIncrementalSlides: true
---

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
knitr::opts_chunk$set(
  fig.width = 9, fig.height = 3.5, fig.retina = 3,
  out.width = "100%",
  cache = FALSE,
  echo = FALSE,
  message = FALSE,
  warning = FALSE,
  hiline = TRUE
)
```

```{r xaringan-themer, include=FALSE, warning=FALSE}
library(xaringanthemer)
style_duo_accent(
  primary_color = "#000080",
  secondary_color = "#f2f3f4",
  inverse_header_color = "#00147e"
)
```

```{r xaringan-editable, echo=FALSE}
xaringanExtra::use_editable(expires = 1)
xaringanExtra::use_scribble()
```

```{r xaringan-logo, echo=FALSE}
xaringanExtra::use_logo(image_url = "https://github.com/alxcn/TecLogoEIC/blob/9562a53875418e749a296c85808a19c85fc4be74/IngenieriaCiencias_Horizontal_RGB.png",
  position = xaringanExtra::css_position(top = "2em", right = "2em")
)
```
# Goals of the Session
* Define the structure of a vector space. <br/><br/>
* Introduce examples of vector spaces. <br/><br/>
* Define vector subspaces. <br/><br/>

---
# What is a vector and what can you do with them?



---
# What is a vector space?

> **Definition:** A vector space is an algebraic structure that consists of a set $V$ of objects called vectors, and two operations, addition and scalar multiplication, that satisfy certain properties. <br/>
The axioms are:
  * The sum is commutative. <br/>
  * The sum is associative. <br/>
  * There is a zero vector. <br/>
  * There is an additive inverse. <br/>
  * The scalar multiplication is associative. <br/>
  * The scalar multiplication distributes over the sum of vectors. <br/>
  * The scalar 1 is the multiplicative identity. <br/>
  * Scalar multiplication distributes over the sum of scalars. <br/>
  
---
## Examples of Vector Spaces.

> Consider the space of real numbers $\mathbb{R}$ as a set of vectors and scalars, with the usual operations. Let's see that it satisfies the properties of a vector space.

--
<br/><br/>
 
> Consider the space of pairs of real numbers $V=\mathbb{R}^2,$ with the following operations: <br/>
* Sum: $(a,b)+(c,d)=(a+c,b+d).$ <br/>
* Multiplication by a scalar: $\lambda(a,b)=(\lambda a,\lambda b).$ <br/><br/>
Let's prove that it is a vector space.

---
## Examples of Vector Spaces.


> Consider the space $V=\mathbb{R}^n$ of n-tuples of real numbers, with the following operations: <br/>
* Sum: $(a_1,a_2,\dots,a_n)+(b_1,b_2,\dots,b_n)=(a_1+b_1,a_2+b_2,\dots,a_n+b_n).$ <br/>
* Multiplication by a scalar: $\lambda(a_1,a_2,\dots,a_n)=(\lambda a_1,\lambda a_2,\dots,\lambda a_n).$ <br/>
Let's prove that it is a vector space.

---
## Examples of Vector Spaces.

> Consider the space $V=M(n\times m, \mathbb{R})$ of $n\times m$ matrices with real coefficients, with the operations of matrices. <br/><br/> Let's prove that it is a vector space.
--
<br/><br/>

> Consider the space $V=P_3(\mathbb{R})$ of polynomials of degree less than or equal to three in the variable $x,$ with real coefficients, with the usual operations of sum of polynomials and scalar multiplication. <br/><br/> Let's prove that it is a vector space.

---
## Counterexamples of Vector Spaces.

> Consider the space of polynomials of degree two in the variable $x,$ with real coefficients. Let's prove that this is not a vector space.

---
## Counterexamples of Vector Spaces.

> Consider the set $W=\{(x,y)\in\mathbb{R}^2:x^2+y^2=1\}.$ <br/><br/> Let's prove that it is not a vector space with the sum and scalar product of vectors in $\mathbb{R}^2.$


---
## Vector Subspaces.

> **Definition:** A vector subspace of a vector space $V$ is a subset $W$ of $V$ that is a vector space with the operations of $V.$ <br/><br/
--
> For example, in the vector space $\mathbb{R}^2,$ the line that passes through the origin is a vector subspace. <br/><br/> Consider the line $W=\{(x,y)\in\mathbb{R}^2:y=mx\}$ where $m$ is a constant. <br/><br/> Let's see that $W$ is a vector subspace.


---
## Vector Subspaces.

> For example, in the vector space $\mathbb{R}^3,$ the lines that pass through the origin are vector subspaces. <br/><br/> Consider the line $W_1=\{(x,y,z)\in\mathbb{R}^3:ax=z\}$ where $a$ is a constant. <br/><br/> Let's see that $W_1$ is a vector subspace.

---
## Vector Subspaces.
> For example, in the vector space $P_3(\mathbb{R}),$ the set of polynomials of degree less than or equal to two is a vector subspace. <br/><br/> Let's see that it is a vector subspace.

---
## Activity

1. Consider $V=\mathbb{R}^2$ with the usual operations. ¿Puede ser $W=\{(3,x):x\in\mathbb{R}\}$ un subespacio vectorial? <br/><br/>
1. Prueba que $\mathbb{Z}^2$ es un espacio vectorial con la suma y producto por escalar de $\mathbb{R}^2$ pero con conjunto de escalares igual a $\mathbb{Z}.$ <br/><br/>
1. Let $V=M(2\times 2,\mathbb{R})$ be the set of $2\times 2$ matrices with coefficients in $\mathbb{R}.$ Prove that $W=\{A\text{ is a diagonal matrix}\}$ is a vector subspace of $V.$ <br/><br/>
1. Let $V=P_2(\mathbb{R})$ be the set of polynomials of degree less than or equal to two with coefficients in $\mathbb{R}.$ Prove that $W=\{p(x)\text{ has degree equal or less than 1}\}$ is a vector subspace of $V.$ <br/><br/>
