---
title: "Areas between curves, and Applications of Definite Integral"
subtitle: "Session 13"
author: 
  - "Alejandro Ucan"
date: '`r Sys.Date()`'
output:
  xaringan::moon_reader:
    css: xaringan-themer.css
    nature:
      slideNumberFormat: "%current%"
      highlightStyle: github
      highlightLines: true
      ratio: 16:9
      countIncrementalSlides: true
---

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
knitr::opts_chunk$set(
  fig.width=9, fig.height=3.5, fig.retina=3,
  out.width = "100%",
  cache = FALSE,
  echo = FALSE,
  message = FALSE, 
  warning = FALSE,
  hiline = TRUE
)
```

```{r xaringan-themer, include=FALSE, warning=FALSE}
library(xaringanthemer)
style_duo_accent(
  primary_color = "#000080",
  secondary_color = "#f2f3f4",
  inverse_header_color = "#00147e"
)
```

```{r xaringan-editable, echo=FALSE}
xaringanExtra::use_editable(expires = 1)
xaringanExtra::use_scribble()
```

```{r xaringan-logo, echo=FALSE}
xaringanExtra::use_logo(
  image_url = "https://github.com/alxcn/TecLogoEIC/blob/9562a53875418e749a296c85808a19c85fc4be74/IngenieriaCiencias_Horizontal_RGB.png",
  position = xaringanExtra::css_position(top = "2em", right = "2em")
)
```

# Goals

* The problem of area bounded between curves. <br/><br/>
* Define type I and type II regions. <br/><br/>
* Find the area between curves. <br/><br/>

---
# Motivation:

Given two functions $f(x)$ and $g(x)$, we want to find the area between the curves $y=f(x)$ and $y=g(x)$ over a determined interval $[a,b]$.


---
# Using the definite integral

> The area between the curves $y=f(x)$ and $y=g(x)$ over the interval $[a,b]$ is given by the definite integral:

$$\int_{a}^{b} [f(x)-g(x)]dx$$

#### Example:

Find the area between the curves $y=x^2$ and $y=2x-x^2$ over the interval $[0,2]$:

---
#### Solution:

The area is given by the definite integral:

$$\int_{0}^{2} [(2x-x^2)-x^2]dx=\int_0^2 (2x-2x^2)dx = x^2-\frac{2}{3}x^3|_{0}^3$$


#### Example 2:

Find the area bounded by the curves $y=x^2-2$ and $y=2-x^2$. 

---
#### Solution:

Since we don't have information about the interval, we need to find the maximal interval for the bounding. Let's find out the intersection points of the curves: $$x^2-2=2-x^2 \Rightarrow 2x^2=4 \Rightarrow x^2=2 \Rightarrow x=\pm \sqrt{2}.$$

Then the area is given by the definite integral: $$\int_{-\sqrt{2}}^{\sqrt{2}} [(2-x^2)-(x^2-2)]dx$$

---
##### Example 3:

Find the area of the triangle with vertices $(0,0), \, (4,0)$ and $(1,4)$ using integrals.

--

<br/><br>

#### Solution

For this case, we cannot use a single function to represent the triangle, but we can use two functions to represent each side of the triangle. This implies that we need to use two different integral.

---
# Type I and Type II regions

* A region is called a type I region if it is bounded by the $x$-axis and two curves $y=f(x)$ and $y=g(x)$.
* A region is called a type II region if it is bounded by the $y$-axis and two curves $x=f(y)$ and $x=g(y)$.

--
<br/><br/> 

**Remark:** To determine the tipe of region, we need to do the rectangle test: if a vertical rectangle intersects the region in two points, then it is a type I region, otherwise it is a type II region.

---

##### Example 4:

Find the area of the triangle with vertices $(0,0), \, (4,0)$ and $(1,4)$ using integrals of type II.

---

#### Example 5:

Find the area bounded by the curves $y=x^3$ and $x=y^3$ over the interval $[0,1]$ as type I and type II region.


---
# Center of Mass

> __Definition:__ Given a wire with endpoint $(a,0)$ and $(b,0)$, that has density $\rho(x)$, the center of mass of the wire is given by the following formula: <br/><br/> $$\overline{x}=\displaystyle{\frac{\int_{a}^{b} \rho(x)x dx}{\int_a^b \rho(x)dx}}$$ and the point where is the center of mass is given by $(\overline{x},0).$

---
#### Example

> Consider a wire from $(0,0)$ to $(10,0)$ with density function given by $x.$ Find the center of mass.

--
<br/><br/>
#### Solution:

1. We compute the corresponding integrals: $$\int_{0}^{10} x dx = 50$$ and $$\int_{0}^{10} x^2 dx = \frac{1000}{3}$$ <br/><br/>
2. We compute the center of mass: $$\overline{x} = \frac{\frac{1000}{3}}{50} = \frac{20}{3}$$ <br/><br/>
3. The center of mass is given by $(\frac{20}{3},0).$

---
#### Example

> Consider a wire from $(0,0)$ to $(\pi,0)$ with density function given by $\sin(x).$ Find the center of mass.

--
<br/><br/>
#### Solution:

1. We compute the corresponding integrals: $$\int_{0}^{\pi} \sin(x) dx = 2$$ and $$\int_{0}^{\pi} x\sin(x) dx = \pi$$ <br/><br/>
2. We compute the center of mass: $$\overline{x} = \frac{\pi}{2}$$ <br/><br/>
3. The center of mass is given by $(\frac{\pi}{2},0).$

---
# Mean value:

> __Definition:__ Given a function $f(x)$ continuous on $[a,b]$, the mean value of the function is given by the following formula: <br/><br/> $$\overline{f}=\displaystyle{\frac{1}{b-a}\int_{a}^{b} f(x) dx}$$

<br/><br/>

**Remark:** The previous definition can be applied to any function, but in particular in statistics and probability, the mean value is the average of a set of values for a given random variable following a probability distribution.

