---
title: "Applied Linear Algebra"
subtitle: "Review"
author: 
  - "Alejandro Ucan"
date: '`r Sys.Date()`'
output:
  xaringan::moon_reader:
    css: xaringan-themer.css
    nature:
      slideNumberFormat: "%current%"
      highlightStyle: github
      highlightLines: true
      ratio: 16:9
      countIncrementalSlides: true
---

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
knitr::opts_chunk$set(
  fig.width=9, fig.height=3.5, fig.retina=3,
  out.width = "100%",
  cache = FALSE,
  echo = FALSE,
  message = FALSE, 
  warning = FALSE,
  hiline = TRUE
)
```

```{r xaringan-themer, include=FALSE, warning=FALSE}
library(xaringanthemer)
style_duo_accent(
  primary_color = "#000080",
  secondary_color = "#f2f3f4",
  inverse_header_color = "#00147e"
)
```

```{r xaringan-editable, echo=FALSE}
xaringanExtra::use_editable(expires = 1)
xaringanExtra::use_scribble()
```

```{r xaringan-logo, echo=FALSE}
xaringanExtra::use_logo(
  image_url = "https://github.com/alxcn/TecLogoEIC/blob/9562a53875418e749a296c85808a19c85fc4be74/IngenieriaCiencias_Horizontal_RGB.png",
  position = xaringanExtra::css_position(top = "2em", right = "2em")
)
```

# Goals for the Session

* Revisitate the concepts of Linear Algebra in engineering contexts.

---
# Determine the appropriate Vector Space

> *Which vector space models the following data?*
<br/>

**Example 1:** A given enterprise stores the information of its employees in a table with the following columns: `Age`, `Hours worked`, `Salary`. What vector space could be useful to model this information? <br/><br/>

--
**Solution** <br/><br/>
Remember that both Age, hours worked, and salary are real quantities, and for each employee we have three data. Therefore, the space that models this data is $\mathbb{R}^3$.

---
# Determine the appropriate Vector Space

**Example 2:** An enterprise dedicated to the sale of electronic products has a table with the information of the products it sells. Each row of the table contains the following information: `Price`, `Quantity sold`, `Profit`. What vector space could be useful to model this information? <br/><br/>

**Example 3:** A company dedicated to the research of drugs, has a table with the information of the drugs it has developed. Each row of the table contains the following information: `Production cost`, `Selling price`, `Profit`, `volume produced`, `quantity of compound A`, `quantity of compound B`, `quantity of compound C`. What vector space could be useful to model this information? <br/><br/>

---
# Determine the appropriate Vector Space

**Example 4:** An automotive company has different tables with the information of each car it produces in different factories during different months. Each row of the table contains the following information: `Production cost`, `Selling price`, `Profit`, `volume produced`, `Production time`, and each row represents a different factory. If the company has 5 factories, what vector space could be useful to model this information? <br/><br/>

**Example 5:** An food company, has a list of tables with the information of each product produced in different processors on different days. Each row of the table contains the following information: `volume`, `sodium content`, `sugar content`, `trans fat content`, `calories` and each row represents a different processor. If the company has 10 processors, what vector space could be useful to model this information? <br/><br/>

---
# Linear Combinations

> *How to use linear combinations to rewrite data information?*

**Example 1:** In a company we have the information of the employees as `Age`, `Hours worked`, `Salary`. The standard is that the age is $(21,0,0),$ the hours worked $(0,8,0)$ and the salary $(0,0,1000)$. How does the information of an employee who is 25 years old, works 10 hours and earns 2000 look like? <br/><br/>

--
**Solution:** <br/><br/>

Using the standardization vectors we can rewrite the information if we solve $$(25,10,2000)=a(21,0,0)+b(0,8,0)+c(0,0,1000)$$ This is when $a=21/25$, $b=10/8$ and $c=2000/1000$.

---
# Linear Combination

**Example 2:** A prescribed drug is composed with three compounds `A`, `B` and `C`. We are interested in properties `1`, `2`, `3`and `4`. If each compound's properties is given by the vectors $A=(1,0.5,0.3,0.1),\, B=(0.5,0.8,0.2,0.4),\, C=(0.2,0.1,0.5,0.7)$, What quantities of compound should I use if I want to obtain a mix such that the properties are $(4,2.5,1.5,0.5)$? <br/><br/>

**Example 3:** Automovile parts need to go through four processes with 3 distinct stages where times are measured. If each process is given by the vectors $P=(1,0.5,0.3),\, Q=(0.5,0.8,0.2),\, R=(0.2,0.1,0.5),\, S=(0.7,0.4,0.6)$, what quantities of processes I need to use if I want to obtain a car such that the times are $(10,12,15)$? <br/><br/>

---
# Linear Combinations

**Example 4:** A food company wants to investigate the amount of vitamins $A,$ $B$ and $C$ found in its products. If each product has three ingredients $I_1=(0.1,0.2,0.3),\, I_2=(0.2,0.3,0.1),\, I_3=(0.3,0.1,0.2)$ per gram. What quantities of ingredients should I use if I want to obtain a product such that the vitamins are $(2,3,2)$? <br/><br/>

**Example 5:** An streaming company knows that by playing a song, the user consumes $0.1$ MB if using `low` quality, $0.5$ MB if using `medium` quality and $1$ MB if using `high` quality. If the proposed streaming algorithm adapts to the user's bandwidth and the song. How many MB in each quality will a user consume if listening to the following songs $A=(0.2,0.2,0.3),\, B=(0.3,0.1,0.2),\, C=(0.1,0.3,0.2)$? <br/><br/>

---
# Linear Transformations.

> *There is a way to modify my vector space to obtain relevant information?*

**Example 1:** We have the information of the working hours of the employees in a company during 5 days. At the end of the week we are interested in knowing how many hours each employee worked. Is there a linear transformation that allows us to obtain this information? <br/><br/>
--
**Solution:** <br/><br/>
If we define the vector space $V=\mathbb{R}^5$ where each vector represents the hours worked by an employee in the week, then the linear transformation $T:V\to \mathbb{R}$ given by $T(v)=v_1+v_2+v_3+v_4+v_5$ allows us to obtain the information we need. It is enough to prove that this map is a linear transformation.

---
# Linear Transformations. 

**Example 2:** With the previous information. If each day of the week has a different hourly cost: $D_1=10$, $D_2=15$, $D_3=20$, $D_4=25$, $D_5=30$. Is there a linear transformation that allows us to obtain the total cost of the week for each employee? <br/><br/>

**Ejemplo 3:** It is possible that the average of hours worked is a linear transformation? In case it is, how would the transformation look like? <br/><br/>

---
# Linear Transformations.

**Example 4:** The Body Mass Index (BMI) is a measure that is calculated with the formula $BMI=\frac{Weight}{Height^2}$. If we have the information of the height and weight of a person, is there a linear transformation that allows us to obtain the BMI? <br/><br/>


**Ejemplo 5:** An enterprise is planning to apply for certain certifications. For this, it needs the processes to meet certain standards (`A`,`B`,`C`). The Logistics team determined that the standards are related in one `D` such that $D=0.3A+0.5B-0.2C$. For which values of standards do we get a rating of $D$ equal to 0? <br/><br/>