---
title: "Quadratic Functions"
subtitle: "Session 05"  
author: 
  - "Alejandro Ucan"
date: '`r Sys.Date()`'
output:
  xaringan::moon_reader:
    css: xaringan-themer.css
    nature:
      slideNumberFormat: "%current%"
      highlightStyle: github
      highlightLines: true
      ratio: 16:9
      countIncrementalSlides: true
---

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
knitr::opts_chunk$set(
  fig.width=9, fig.height=3.5, fig.retina=3,
  out.width = "100%",
  cache = FALSE,
  echo = FALSE,
  message = FALSE, 
  warning = FALSE,
  hiline = TRUE
)
```

```{r xaringan-themer, include=FALSE, warning=FALSE}
library(xaringanthemer)
style_duo_accent(
  primary_color = "#000080",
  secondary_color = "#f2f3f4",
  inverse_header_color = "#00147e"
)
```

```{r xaringan-editable, echo=FALSE}
xaringanExtra::use_editable(expires = 1)
xaringanExtra::use_scribble()
```

```{r xaringan-logo, echo=FALSE}
xaringanExtra::use_logo(
  image_url = "https://github.com/alxcn/TecLogoEIC/blob/9562a53875418e749a296c85808a19c85fc4be74/IngenieriaCiencias_Horizontal_RGB.png",
  position = xaringanExtra::css_position(top = "2em", right = "2em")
)
```
# Goals

* Describe the general form a the Quadratic function model. <br/><br/>
* Describe intercepts and vertices of the quadratic function. <br/><br/>
* Apply the quadratic function to model business problems. <br/><br/>


---
# General Form of the Quadratic Function

> __Definition:__ The quadratic function is a polynomial function of degree 2. The general form of the quadratic function is given by: $$f(x)=ax^2+bx+c$$ where $a$, $b$ and $c$ are real numbers and $a\neq 0$. The domain of the quadratic function is the set of all real numbers, but the range differs from one function to another.

---
## Its graph

> __Definition:__ The graph associated to a quadratic function is called a parabola. The parabola is symmetric with respect to the vertical line that passes through the vertex. The vertex is the point where the parabola changes direction. The parabola opens upward if $a>0$ and downward if $a<0$.

---
## Its graph

```{r, echo=FALSE, fig.align='center', fig.cap="Parabola"}
library(ggplot2)
ggplot(data.frame(x=c(-5,5)), aes(x)) + 
  stat_function(fun = function(x) -3*x^2, geom = "line") +
  labs(x = "x", y = "f(x)")
```

---
## The intercepts with the x-axis.

> __Definition:__ The intercepts with the x-axis are the points where the parabola intersects the x-axis. The x-intercepts are the solutions of the equation $f(x)=0$. The x-intercepts are also called the zeros of the function.

--
In order to find our roots, we need to solve the equation $f(x)=0$. This equation is called the __quadratic equation__ and it has the form: $$ax^2+bx+c=0,$$ we can use the quadratic general formula to solve that equation: $$x=\frac{-b\pm\sqrt{b^2-4ac}}{2a}$$ and dependending on the number $\Delta=b^2-4ac$ we can have three different cases:
* $\Delta>0$ two real roots.
* $\Delta=0$ one real root.
* $\Delta<0$ no real roots.

---
## The intercepts with the x-axis.

> __Definition:__ the number $\Delta=b^2-4ac$ is called the __discriminant__ of the quadratic equation. 

<br/><br/>
In order to interpretate the discriminant, we need to think that:
* If $\Delta>0$ then the parabola intersects the x-axis in two different points.
* If $\Delta=0$ then the parabola intersects the x-axis in one point.
* If $\Delta<0$ then the parabola does not intersect the x-axis.

---
#### Examples:

> Determine the nature of the roots for each quadratic function: <br/><br/>
  * $f(x)=x^2-4x+3$ <br/><br/>
  * $f(x)=x^2-4x+4$ <br/><br/>
  * $f(x)=x^2+2x+1$ <br/><br/>
  * $f(x)=x^2+1$ <br/><br/>

---
## The vertex of the parabola.

> __Definition:__ The vertex of the parabola is the point where the parabola changes direction. The vertex is the point $(h,k)$ where $h$ and $k$ are the coordinates of the vertex. Usually, this point is where the parabola attains its minimum or maximum value.

--
To easily find my vertex, we need to write our parabola equation in the form: $$f(x)=p(x-h)^2+k$$ where $h$ and $k$ are the coordinates of the vertex.

---
### The Square Completing process

Assume that we have the following quadratic function: $$f(x)=ax^2+bx+c$$ and we want to write it vertex form. We can use the following process:<br/><br/>
1. Take the two first terms and factor out the coefficient of the squared term. $$f(x)=a(x^2+\frac{b}{a}x)+c$$ <br/><br/>
1. Add and subtract the square of half the coefficient of the linear term. $$f(x)=a(x^2+\frac{b}{a}x+\frac{b^2}{4a^2}-\frac{b^2}{4a^2})+c$$ <br/><br/>
1. Factor the first three terms and simplify. $$f(x)=a(x+\frac{b}{2a})^2-\frac{b^2}{4a}+c$$ <br/><br/>
1. Simplify the last two terms. $$f(x)=a(x+\frac{b}{2a})^2-\frac{b^2-4ac}{4a}$$ <br/><br/>

---
#### Example 1:

> Write the quadratic function $f(x)=x^2-4x+3$ in vertex form.
--
<br/><br/>

##### Solution:

$$f(x)=x^2-4x+3=x^2-4x+4-4+3=(x-2)^2-1$$

Note that here, $a=1$, $b=-4$ and $c=3,$ and $\Delta=b^2-4ac=16-12=4>0$ and $$h=\frac{-b}{2a}=\frac{4}{2}=2$$ and $$k=\frac{-b^2+4ac}{4a}=\frac{16-12}{4}=1$$

---
#### Example 2:

> Write the quadratic function $f(x)=x^2+1$ in vertex form.
--
<br/><br/>

##### Solution:

$$f(x)=x^2+1=x^2+1+0=(x+0)^2+1$$

Note that here, $a=1$, $b=0$ and $c=1,$ and $\Delta=b^2-4ac=0-4=4<0$ and $$h=\frac{-b}{2a}=\frac{0}{2}=0$$ and $$k=\frac{-b^2+4ac}{4a}=\frac{0+4}{4}=1$$

---
# Applications

#### Example 3:

> The financial department in the company that produces a digital camere arrived at the following price-demand function and the following revenue function: $$p(x)=94.8-5x$$ $$R(x)=xp(x)=x(94.8.5x)$$ where $p(x)$ measures the wholesale price per camera at which $x$ million cameras can be sold, and $R(x)$ is the corresponding revenue (in million of dollars). We are interested in the case where $1\leq x\leq 15.$ <br/><br/>
  * Find the price that maximizes the revenue. <br/><br/>
  * Find the maximum revenue. <br/><br/>

---
#### Example 4:

> The marketing research department for a company that manufactures and sells notebook computers stablished the following price-demand and revenue functions: $$p(x)=2000-60x\quad R(x)=xp(x)=x(2000-60x)$$ where $p(x)$ measures the wholesale price per computer at which $x$ thousand computers can be sold, and $R(x)$ is the corresponding revenue (in thousand of dollars). We are interested in the case where $0\leq x\leq 25.$ <br/><br/>
  * Find the price that maximizes the revenue. <br/><br/>
  * Find the maximum revenue. <br/><br/>

