---
title: "Orthogonality"
subtitle: "Session 08"  
author: 
  - "Alejandro Ucan"
date: '`r Sys.Date()`'
output:
  xaringan::moon_reader:
    css: xaringan-themer.css
    nature:
      slideNumberFormat: "%current%"
      highlightStyle: github
      highlightLines: true
      ratio: 16:9
      countIncrementalSlides: true
---

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
knitr::opts_chunk$set(
  fig.width=9, fig.height=3.5, fig.retina=3,
  out.width = "80%",
  cache = FALSE,
  echo = FALSE,
  message = FALSE, 
  warning = FALSE,
  hiline = TRUE
)
```

```{r xaringan-themer, include=FALSE, warning=FALSE}
library(xaringanthemer)
style_duo_accent(
  primary_color = "#000080",
  secondary_color = "#f2f3f4",
  inverse_header_color = "#00147e"
)
```

```{r xaringan-editable, echo=FALSE}
xaringanExtra::use_editable(expires = 1)
xaringanExtra::use_scribble()
```

```{r xaringan-logo, echo=FALSE}
xaringanExtra::use_logo(
  image_url = "https://github.com/alxcn/TecLogoEIC/blob/9562a53875418e749a296c85808a19c85fc4be74/IngenieriaCiencias_Horizontal_RGB.png",
  position = xaringanExtra::css_position(top = "2em", right = "2em")
)
```
# Goals of the Session

* Introduce the concept of inner product. <br/><br/>
* Define the angle between two vectors. <br/><br/>
* Introduce the concept of orthogonality. <br/><br/>
* Define the projection of a vector onto another vector. <br/><br/>

---
# Inner Product

> __Definition:__ Let $V$ be a vector space over $\mathbb{R}$. An inner product on $V$ is a function $\langle \cdot, \cdot \rangle: V \times V \rightarrow \mathbb{R}$ that satisfies the following properties for all $u, v, w \in V$ and $c \in \mathbb{R}$: <br/><br/>
  1. Commutative: $\langle u, v \rangle = \langle v, u \rangle$ <br/>
  1. Linear in the first argument: $\langle cu + w, v \rangle = c \langle u, v \rangle + \langle w, v \rangle$ <br/>
  1. Positive definite: $\langle u, u \rangle \geq 0$ and $\langle u, u \rangle = 0$ if and only if $u = 0$ <br/><br/>
---

## Examples of Inner Product

Let $V=\mathbb{R}^n$ and $u, v \in \mathbb{R}^n$, then the standard inner product on $\mathbb{R}^n$ is given by: <br/><br/> $$\langle u, v\rangle = u_1v_1 + u_2v_2 + \cdots + u_nv_n = \sum_{i=1}^n u_iv_i$$
--

Let $V=\mathbb{R}^2$ and $u, v \in \mathbb{R}^2$, then the standard inner product on $\mathbb{R}^2$ is given by: <br/><br/> $$\langle u, v\rangle = u_1v_1 + 2u_2v_2$$ 

---
## Counterexample

Let $V=\mathbb{R}^2$ and $u, v \in \mathbb{R}^2$, then the function $\langle u, v\rangle = u_1v_1 - u_2v_2$ is not an inner product on $\mathbb{R}^2$ because it is not positive definite. <br/><br/> For example, $\langle (1,2), (1,2)\rangle < 0$.

---
# Projections and Orthogonality

> **Definition:** Let $V$ be a vector space with an inner product, and $u, v \in V$ two vectors. We define the projection of $u$ onto $v,$ denoted by $\text{proj}_v(u),$ as the vector $$\text{proj}_v(u)=\frac{\langle u,v\rangle}{\langle v,v\rangle} v.$$
--
##### Example
Let $u$ and $v$ be the vectors $(2,3)$ and $(4,1)$ in $\mathbb{R}^2$ with the standard inner product. Then, $$\text{proj}_{(4,1)}(2,3)=\frac{\langle (2,3),(4,1)\rangle}{\langle (4,1),(4,1)\rangle} (4,1)=\frac{11}{17}(4,1).$$
---
## Angle between Two Vectors

> **Definition:** Let $V$ be a vector space with an inner product, we define the angle between two vectors $u,v\in V$ as the angle $\theta$ such that $$\cos(\theta)=\frac{\langle u,v\rangle}{\sqrt{\langle u,u\rangle \langle v,v\rangle}}.$$
--
##### Example

Find the angle between $(2,3)$ and $(4,1)$ in $\mathbb{R}^2$ with the standard inner product.
--
$$\cos(\theta)=\frac{\langle (2,3),(4,1)\rangle}{\sqrt{\langle(2,3),(2,3)\rangle\langle(4,1),(4,1)\rangle}}=\frac{11}{\sqrt{(13)(17)}}.$$ Por lo tanto, $\theta=\cos^{-1}\left(\frac{11}{\sqrt{(13)(17)}}\right).$

---
## Orthogonality

> **Definition:** We say that two vectors $u,v\in V$ are orthogonal if $\langle u,v\rangle=0.$

--
##### Remark:

If $\langle u,v\rangle=0,$ then $\cos(\theta)=0,$ therefore, $\theta=\pm k \frac{\pi}{2}$ with $k$ an odd integer.
---
##### Example: 

> In $\mathbb{R}^2,$ these vectors <br/><br/>
 * $(0,1),(1,0),$ <br/><br/>
 * $(1,1),(-1,1),$ <br/><br/>
 * $(3,1), (1,-3)$ <br/><br/>
 are orthogonals to each other.

 ---
## Orthogonal and Orthonormal Bases:

> **Definition:** A subset $S$ of $V$ is orthogonal if for all $u,v\in S$ with $u\neq v,$ we have $\langle u,v\rangle=0.$ <br/><br/> We say that $S$ is orthonormal if in addition to being orthogonal, $\langle u,u\rangle=1$ for all $u\in S.$
--
##### Example

The set $S=\left\{\left(\frac{1}{\sqrt{2}},\frac{1}{\sqrt{2}},0\right),\left(-\frac{\sqrt{2}}{6},\frac{\sqrt{2}}{6},\frac{2\sqrt{2}}{3}\right),\left(\frac{2}{3},-\frac{2}{3},\frac{1}{3}\right)\right\}$ is orthonormal.

---
## Orthogonal and Orthonormal Bases:

> **Theorem:** Let $V$ be a vector space with an inner product, and $S=\{v_1,v_2,\ldots,v_n\}$ an orthogonal set of non-zero vectors. Then $S$ is linearly independent.
--

> **Corollary:** If $V$ has dimension $n,$ then any orthogonal set of $n$ vectors is a basis.
---

# Gram-Schmidt Process

1. Let $S={v_1,v_2,\ldots,v_n}$ be a set of linearly independent vectors of $V.$ <br/><br/>
2. Define $u_1=v_1.$ <br/><br/>
3. For $k=2,3,\ldots,n,$ define $u_k=v_k-\sum_{j=1}^{k-1}\frac{\langle v_k,u_j\rangle}{\langle u_j,u_j\rangle}u_j.$ <br/><br/>
4. Define $w_1=\frac{u_1}{\sqrt{\langle u_1,u_1\rangle}},$ and for $k=2,3,\ldots,n,$ define $w_k=\frac{u_k}{\sqrt{\langle u_k,u_k\rangle}}.$ <br/><br/>

---
##### Example: 

Let $S=\{(1,1),(0,1)\}.$ Construct an orthonormal basis from $S.$ <br/><br/>

--
##### Solution:
$v_1=(1,1)$ y $v_2=(0,1).$ Following the Gram-Schmidt process, we have $u_1=v_1$ y $$u_2=v_2-\frac{\langle v_2,u_1\rangle}{\langle u_1,u_1\rangle}u_1=v_2-\frac{1}{2}v_1=(-\frac{1}{2},\frac{1}{2}).$$ <br/><br/> Therefore, $w_1=\frac{u_1}{\sqrt{\langle u_1,u_1\rangle}}=\frac{1}{\sqrt{2}}(1,1)$ y $w_2=\frac{u_2}{\sqrt{\langle u_2,u_2\rangle}}=\left(-\frac{\sqrt{2}}{2},\frac{\sqrt{2}}{2}\right).$ <br/><br/>

---
##### Example:

Let $S=\{(1,1,1),(1,0,1),(1,1,0)\}.$ Construct an orthonormal basis based on $S.$ <br/><br/>

---
##### Ejemplo:

Let $S=\{(1,1,0),(1,2,0),(0,1,2)\}.$ Construct an orthonormal basis bases on $S.$